### 2018-09-17 article/article.tex
Source and supporting data for "Impact of switching bug trackers"

\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
\usepackage{amsmath}
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\usepackage{hyperref,xcolor}
\renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Impact of switching bug trackers: a case study}
%
%\titlerunning{}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Th\'eo Zimmermann\inst{1,2} \and
Annal\'i Casanueva Art\'is\inst{3}\thanks{This work was partly funded by El Conacyt of the Mexican government.}}
%
\authorrunning{T. Zimmermann and A. Casanueva Art\'is}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{$\Pi R^2$, Inria, Paris, France \and
IRIF, Paris-Diderot University, Paris, France\\
\email{theo@irif.fr} \and
Paris School of Economics, Paris, France}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Any significant software project requires a bug tracker. In open source development, this plays an even bigger role. The impact of the bug reporting environment on the bug tracking activity is difficult to assess. In this paper, we take advantage of the switch from Bugzilla to GitHub of the bug tracker of Coq, a medium-sized open source project, to evaluate the impact that such a change can have. 
We first report on the switch itself, in particular the migration of 4900 preexisting bug reports using the GitHub issue import API; then we analyze data from before and after the migration. We show that the switch induces 
an increase in bug reporting, particularly from principal developers themselves, and more generally an increased engagement with the bug tracking platform, with more comments by developers and also more external commentators.

%The abstract should briefly summarize the contents of the paper in 15--250 words.

\keywords{bug tracker \and switch \and migration  \and bug report \and issue \and GitHub \and Bugzilla \and open source.}
\end{abstract}
%
%
%
\section{Introduction}

Bug reporting is an essential part of software development. However, it involves
a non-negligible human component, particularly in the context of an open source project where a larger part of bugs are reported by independent users and therefore depend a lot on the good will of individuals. Creating an environment that makes bug reporting easier and more appealing may lead to an increase of bug reporting (from both developers and independent users) which should ultimately benefit  software development \cite{Bettenburg2008} (provided that developers are equipped to handle the incoming reports \cite{Anvika,Kanwal2012}). However, the impact of such change has rarely been studied. Indeed, comparing the number of bug reports in different environments is generally not compelling because it is likely to depend  not just on the reporting environment but also on each particular software and its development dynamics. To do a valid comparison one would ideally need to compare two bug reporting environments for the same software, in the same period. But, or course, we can't observe this ideal. A context approaching this ideal scenario would be to compare the number of bug reports of a particular software before and after the switch to a new environment. Yet, substantial changes in the environment of bug reporting (e.g. changes of bug tracker platform) are rare because once a bug system is in place, any change comes with a significant cost.

We use the switch, from Bugzilla to GitHub, of the bug tracker system of the Coq proof assistant, a medium-sized open source software, to analyze the  effects of a change in the bug reporting environment on the number of bugs, reporters, bug comments and commentators. We describe the switch event and the associated migration process for preexisting bugs. Thanks to the way this migration was conducted, getting data on bug reporting from before and after the switch is easily done using the GitHub GraphQL API. Using a Regression in Discontinuity analysis we find strong evidence that the switch of platform increased the number of bug reports; started an increase of bug reporters per period (i.e. in a certain period there are more distinct reporters); increased the number of comments by developers and the number of distinct commentators per period among non-developers. Taken together our results suggest that switching from Bugzilla to GitHub increased the number of bug reports and bug reporters (probably by making the bug reporting process easier) but also increased the openness of software development and aroused discussion around bugs (notably between developers and non-developers).

\section{Context}

\subsection{Coq and GitHub}
Coq is a (free and open source) proof assistant, that is a software to write and automatically verify mathematical proofs (and programs). It originated as a research project in 1984 and the development team is still composed, for the most part, of researchers. It gained in popularity along the years, and is now at the center of many verification projects (some very large academic projects, and some industrial projects). Its initial developers were awarded  the ACM SIGPLAN Programming Languages Software 2013 Award and the ACM Software System 2013 award.

The development is currently centered around the GitHub platform. The current repository of Coq the current repository counts more than 25000 commits. Some of these commits date back from 1999, when the code received major architectural changes. The GitHub repository was initially created as a mirror of the official repository (which itself was migrated from SVN to git in 2013) and really started attracting the interest of the development team as a way to open up the development to external contributions, especially starting with the first Coq coding sprint in 2015 (later renamed into the Coq Implementors Workshop and held every year in France). From then on, core developers started using pull requests for their own changes more and more often, and it became the norm around the beginning of 2017 when systematic continuous integration tests were introduced on every pull requests. In the last two years, more than 2,200 pull requests have been opened.

\subsection{The Coq bug tracker}

From 2007 to 2017, the Coq bug tracker platform was a self-hosted Bugzilla instance (before this, from 2001 to 2007 it was using JitterBug, a now discontinued bug tracking system, and before 2001 a simple mailing list). In 2017, developers had become quite accustomed to the look and feel of discussion threads on GitHub pull requests, which offered a much nicer experience than the Bugzilla bug tracker. Younger developers (in particular, one of the author of this article, who eventually pushed for the switch and conducted the migration) were frustrated by what they perceived as an older and heavier platform. The rest of the developers were convinced to migrate towards a non-self-hosted solution in August 2017, when the whole Coq website was shut down for more than a week  because an IT team was frightened by a scam bug report.\footnote{\url{https://sympa.inria.fr/sympa/arc/coq-club/2017-08/msg00040.html}}

The expected advantages of the GitHub issue tracker over Bugzilla were:
\begin{itemize}
\item Better formatting, ability to edit comments, and more generally a more pleasant tool to use. We felt that Bugzilla was unpleasant enough to discourage developers to use it to its full potential.
\item Browsing between code, pull requests and bug reports is easier when everything is in the same place. Furthermore, it means a common notification, mention, and assignment system.
\item Better indexing by search engines.
\item Shared management of permissions, milestones, categories (labels), etc.
\item Easier to get started for newcomers (especially as many may already know GitHub and already have a GitHub account).
\end{itemize}

The possible drawbacks were mostly that GitHub provides a less-advanced system with very few customization options. In particular:
\begin{itemize}
\item The permission system is much less fine-grained. Giving triaging rights to users requires giving them ``write access'', which includes commit rights on non-protected branches. That's why the Coq GitHub repository now only has protected branches (and accidentally-pushed topic branches are promptly deleted).
\item Creating a new bug report doesn't require filling a form, but typing a message. Fortunately, issue templates help give some hints towards the expected structure of the report, and being able to strip this template can actually give flexibility to developers who might open new issues for concerns that do not qualify as bugs.
\item There is no import mechanism that allows to keep the number and author of preexisting reports. Fortunately, as we will see in Sec.~\ref{migration}, it was still possible to import bug reports while keeping most of the meta-information and renumbering the fewest possible bug reports. This was an important requirement for the switch.
\end{itemize}

\section{Hypotheses}

Given the advantages of the GitHub issue tracker that were listed above, we expected the switch to:
\begin{itemize}
\item increase the number of bug reports, by developers in particular;
\item increase the diversity of bug reporters;
\item improve the interaction between reporters and developers.
\end{itemize}

All of this could lead in turn to a greater openness and transparency of the development and an increased engagement of users toward it, changing (and probably improving) the dynamics of a crucial part of software development.

\section{Migration}
\label{migration}

The only complicated part of the bug tracker switch was the migration of preexisting bug reports. We reused a tool by Andriy Berestovskyy\footnote{Tool available at \url{https://github.com/berestovskyy/bugzilla2github}.} which is designed to import Bugzilla reports (extracted as an XML dump) to GitHub using its REST API. The bugs are imported in an order designed to preserve numbers whenever possible. Bugs whose number is unavailable (e.g. because the number is already taken by a GitHub pull request) are postponed and renumbered. We implemented several improvements to make the tool better fit our needs:
\begin{itemize}
\item \textit{Allowing non-consecutive bug numbers:} Our imported set of bug reports had some holes in the numbering due to deleted bugs. We use postponed bugs to fill the holes. This improvement has now been integrated upstream.
\item \textit{Saving a table of correspondence for renumbered bugs:} This was used later to redirect the old bug URLs to the new ones.
\item \textit{Using the GitHub issue import beta API and overcoming the GitHub rate limits:} Creating a new issue or a new comment through the normal GitHub REST API will trigger notifications (for people who are watching the repository or are mentioned in the issue thread). Such actions are therefore subject to a strict rate limit which prevented using this tool for importing more than a few hundred bugs. Fortunately, GitHub provides a (beta) issue import API which, in addition to not triggering notifications, also allows importing one bug, its comments, and meta information such as closed status and assignee in a single request (thus reducing the duration to import 4900 bugs to just a few hours). Finally, using this API allows to keep the dates of imported bug reports and comments, which is very useful to this study. We didn't manage to import correct closing dates, so we won't study the impact of the bug tracker switch on the time it takes to close a bug.
\end{itemize}

This tool required a correspondence table between Bugzilla and GitHub accounts. Ours was created by manually matching 217 (out of 686) Bugzilla accounts to 175 GitHub accounts,\footnote{The correspondence table can be seen at \url{https://gist.github.com/Zimmi48/d923e52f64fe17c72852d9c148bfcdc6##file-bugzilla2github-L47-L264}.} the difference being due to the high prevalence of duplicate Bugzilla accounts. We gave priority to finding GitHub accounts for people who still had opened bug reports. The most efficient way to do the matching was by searching the name of the reporter + ``github'' on Google. We also sent e-mails to people we hadn't managed to match (and who still had opened bugs).

\paragraph{Urgency.}
Once the switch was approved by the development team on October 3-4\textsuperscript{th}, 2017, the migration had to happen as soon as possible because every new pull request before the migration added to the number of bug reports that would need to be renumbered. It was conducted\footnote{More details on the process are available in a blog post at \url{http://theoz.im/bugzilla}. } on October 18\textsuperscript{th},  the day after the 8.7.0 release (and not before to avoid disturbing the release process). Only 502 bugs (whose numbers were below 1154) had to be renumbered. This number is quite low because these bugs are dating back from the JitterBug era and during the previous switch only the bugs that were still open had been migrated.

\section{Data}

\subsection{Extraction}

All the data for this study was extracted on September 14\textsuperscript{th}, 2018 using the GitHub GraphQL API. Using this API allows to do large requests (100 nodes in a single request) with just the information we need (thus both reducing the bandwidth usage and speeding up the extraction process). We provide the extracted data as CSV files and a Jupyter notebook\footnote{Available at \url{https://github.com/Zimmi48/impact-of-switching-bug-trackers}.} with the code to request this data from GitHub, to load the CSV files, and to run the analyses.

Migrated bugs and comments had their author information stored in the text so we extract it from there. Dates of bug and comment creation have been preserved so we can obtain them transparently for bugs before and after the migration. Comments dating from the JitterBug era do not have valid author information (and closed bug reports from this era have not been migrated) so we won't consider any data from before 2008.

\subsection{Pre-processing}

\paragraph{Excluding specific reporters.}
We have one specific reporter, Jason Gross, who is alone responsible for almost a quarter of all bug reports. To avoid having the behavior of a single individual strongly impact the overall statistics, we exclude his comments, bug reports and pull requests and all the comments they received from our analysis. Similarly, we also exclude the developer who was the initiator and conductor of the bug tracker switch and migration and who is an author of this article.

\paragraph{Merging duplicate Bugzilla accounts.}
This is only relevant to the new reporter analysis which itself is only detailed in the accompanying Jupyter notebook, so refer there for more information.

\paragraph{Removal of migration artifact comments.}
The migration tool created an issue whose body contains only meta-information, and a first comment with the description, for every bug report. These comments are thus migration artifacts and we remove them from our comment analyses. They can be easily identified because they are the comments that were posted at the exact same date and time as the corresponding bug report.

\subsection{Variable definition}

Our outcome variables will be the numbers of bug reports, distinct bug reporters, new bug reporters (who had never reported a bug before), comments and distinct commentators, for a given period. We will use 15-day periods to have sufficiently many bugs by period. Indeed, in our data we have an average of 13 bug reports by 15-day period; 50\% of the periods have between 7 and 18 reports; 90\% of the periods have between 4 and 27. This also allows to have 40 observations centered around the date of the switch for our Regression in Discontinuity analyses.

We will analyze heterogeneous effects by distinguishing between developers and others (others being assimilated to ``users''). We define ``developers'' as the persons who have contributed more than 100 commits since 2008.

When studying the influence of new releases, we define the release dates by the dates of the corresponding news items on the official website (\url{https://coq.inria.fr/news}). We selected beta and major releases and excluded release candidates and patch-level releases. The first release to appear in the news section of the current official website is the 8.2beta release in June, 2008.

\section{Descriptive statistics}

The accompanying Jupyter notebook contains lots of descriptive statistics and figures, such as the distribution of bug reporting in the year, and in the day, graphs of bug reporting for individual developers, etc. We only present a subset of these figures.

\subsection{Cumulative number of bug reports and bug comments}

Fig.~\ref{cumulative_bugs_and_comments} gives a first (cumulative) view of two of our outcome variables: the number of bug reports, and the number of comments. From this figure, we can clearly notice the dominance of developers in the number of bug comments and the dominance of non-developers in the number of bug reports. We can also already notice some slope changes at the bug tracker switch date. The magnitude and significance of this changes will be assessed in the remainder of this article.

\begin{figure}
\includegraphics[width=\textwidth]{cumulative_bugs_and_comments.eps}
\caption{Cumulative number of bug reports and comments (since 2008).} \label{cumulative_bugs_and_comments}
\end{figure}

\subsection{Influence of new releases}
\label{influence_release}

\paragraph{Influence on the number of bug reports.}
We expect release dates to be correlated to peaks of activity on the bug tracker, in particular peaks of new bug reports. Indeed, users will generally test new releases, for instance by trying to port preexisting projects to the new version, which might lead them to find new bugs (either regressions or bugs related to new features). In Fig.~\ref{bug_nb_with_releases}, we can see that while release dates are generally correlated to peaks of new bug reports by non-developers, this isn't so much the case of reports by developers. We can also notice that beta releases are often correlated to higher peaks than final releases which isn't surprising given that the users who test beta versions are more likely to be ready to report bugs, and beta versions are likely to be less polished.

The highest peak is in 2015, after the first beta release of version 8.5. This was the first release in more than two years and it contained five big new features (``the result of five specific long-term projects''\footnote{Quote from \url{https://coq.inria.fr/refman/credits.html##credits-version-8-5}. The next quotes are coming from the following sections of the same chapter of the Coq reference manual.}). Stabilizing these features and their (unplanned) interactions required more than a year of testing and three beta releases. The 8.5 release was both hard on developers, who found the release cycle too long, and on users, who were impacted by many compatibility issues. Consequently, a different approach was taken for the following releases: versions 8.6, 8.7 and 8.8 were ``developed on a time-based development cycle'' and contain ``the result of refinements, stabilization of features and cleanups of the internals of the system along with a few new features''. Attention was given to regression testing by systematically testing changes by building a selection of the largest Coq projects. Therefore, it is not surprising that these frequent releases are correlated to smaller peaks of new bug reports.

\begin{figure}
\includegraphics[width=\textwidth]{bug_nb_with_releases.eps}
\caption{Number of bug reports by 15-day period with release dates (since 2008).} \label{bug_nb_with_releases}
\end{figure}

\paragraph{Influence on the number of comments.}
In Fig.~\ref{comments_with_releases}, we can see that release dates are correlated to peaks of comments by developers (and not so much by other people). This is explained by the fact that bug reports are generally answered by developers, thus any peak in bug reporting is bound to induce a peak in bug commenting by developers.

\begin{figure}
\includegraphics[width=\textwidth]{comments_with_releases.eps}
\caption{Number of bug comments by 15-day period with release dates (since 2008).} \label{comments_with_releases}
\end{figure}

\section{Methodology}

We exploit the fact that the switch to GitHub can be seen as a clear cutoff in time to use a Regression in Discontinuity approach to estimate the effects of the switch on our outcome variables, to determine if these estimated effects are statistically significant, and to interpret these effects in a causal way. 
The idea behind this analysis is that bug reporters did not choose the exact moment in which the switch would take place. In the absence of any other discontinuity, the changes in the behavior of bug reporters just before and just after the migration is likely to depend mainly on the switch even if there was a change in behavioral trends before the switch. It is important to note that estimated effects using this method will inform of the effect near the threshold. Indeed, in the same way that behavior just before and just after the threshold is not likely to be different in the absence of the switch, the behavior several periods before and several periods after is likely to be different even in the absence of the switch. See~\cite{lee2010regression} for more details on Regression on Discontinuity analysis.

We use time as running variable and we estimate the following regression:

\begin{equation*}
\begin{aligned}
\mbox{\emph{Number of bug reports}}_p⁡ =\gamma_0 &+\gamma_1 \mbox{\emph{Relative date}}_p + \gamma_2 \mbox{\emph{After switch}}_p \\
&+ \gamma_3 \mbox{\emph{Relative date}}_p * \mbox{\emph{After switch}}_p + \epsilon_p
\end{aligned}
\end{equation*}

\noindent where $\mbox{\emph{Number of bug reports}}_p⁡$ is the total number of bugs reported during the 15-day period $p$; $\mbox{\emph{Relative date}}_p$ is the number of periods from the date of the switch (zero is the first period after the switch); $\mbox{\emph{After switch}}_p$ is a binary variable equal to one if $p$ is a period after the switch and zero otherwise; $\mbox{\emph{Relative date}}_p * \mbox{\emph{After switch}}_p$ is the interaction term and $\epsilon_p$ is the error. We estimate this regression by Ordinary Least Squares. Coefficients $\gamma_2$ and $\gamma_3$ are the estimates of interest and will tell us the jump of the number of bugs reports just after the migration and the change in trends due to the migration respectively. We estimate this regression for two different sub-samples: the developers and the non-developers.

We replicate this analysis for all our outcome variables by changing the variable on the left hand of the equation.

\section{Results}

\subsection{Impact on the number of bugs}

\begin{table}
\centering
\caption{Estimated impact of the switch on the number of bugs. Coefficients are highlighted if they are statistically significant with $p<0.1$ (*), $p<0.05$ (**) or $p<0.01$ (***). Standard error is in parentheses.}\label{tab:bug_nb}
\begin{tabular}{|r|c|c|c|}
\hline
&  Total & Developers & Others \\
\hline
$\mbox{\emph{After switch}}_p$ & 10.1* & 8.79** & 1.27 \\
& (5.06) & (3.55) & (4.16) \\
\hline
$\mbox{\emph{Relative date}}_p * \mbox{\emph{After switch}}_p$ & 0.717 & -0.335 & 1.05*** \\
& (0.438) & (0.307) & (0.36) \\
\hline
$\mbox{\emph{Relative date}}_p$ & 0.219 & 0.25 & -0.0308 \\
& (0.309) & (0.217) & (0.255) \\
\hline
Constant & 17.6*** & 5.72** & 11.9*** \\
& (3.71) & (2.6) & (3.05) \\
\hline
Observation number & 40 & 40 & 40 \\
\hline
\end{tabular}
\end{table}

Table~\ref{tab:bug_nb} presents the estimated impact of the switch on the number of bugs. Each column shows the estimates for a different sub-sample (i.e. all bug reporters, developers and non-developers). Asterisks represent the level of significance.  Fig.~\ref{bug_nb_rd}  shows the number of reported bugs before and after the switch and the fitting lines corresponding to the regression results. These results give strong evidence that the switch positively affected  the number of bug reports. We can observe different effects for developers and non-developers.

For developers we see a statistically significant positive effect on the jump of bug reports just after the switch and then a stabilization of the reported bugs at a higher level than before.
In particular changing the bug reporting platform is estimated to increase the number of reported bugs by 8 bugs. That is, in average, we observe 8 more bugs during 15-day periods after the switch (but close to the switch) compared to 15-day periods before the switch (but close to the switch). This means that the number of bug reports by developers after the switch is more than twice the number before the switch.

For non-developers, we see a statistically significant increase in the slope but not on the jump in magnitude just after the switch.
In particular, switching the bug reporting platform to GitHub increases the slope of the regression line by 1. This means that after the switch we will, in average, observe an increase of one bug for each 15-day period.

This behavioral difference is not surprising. Developers were well-aware that the migration had happened and this caused a spike in activity. Traditionally, they had used the bug tracker very little (fixing bugs when they saw them or using personal to-do lists) and were using it more all a sudden (as a shared to-do list and a way to increase communication ---note that one of the \emph{explicit} goals of the switch was to trigger this culture shift). On the contrary, the migration was not responsible for any significant leap in the rate of bug reporting by non-developers, but this rate then increased more and more. This could be due in particular to users noticing that their bug reports were treated more promptly, which could be an encouragement for reporting more of them. Another possible influencing factor may be the fact that non-developers may take more time to adapt to this new platform. Indeed, developers were the ones that took the decision to switch the bug tracker, among other reasons because they were already used to GitHub.

\begin{figure}
\includegraphics[width=\textwidth]{bug_nb_rd.eps}
\caption{Number of bug reports by 15-day period before and after the migration (with fitting lines from the regression results).} \label{bug_nb_rd}
\end{figure}

\subsection{Impact on the number of distinct bug reporters}

Table~\ref{tab:reporters} gives the estimated impact of the switch on the number of bug reporters (i.e. number of people that report a bug during a given period). As before, each column shows the estimates for a different sub-sample. Fig.~\ref{reporter_nb_rd}  shows the number of bug reporters before and after the switch and the fitting lines correspond to the regression results. These results demonstrate that the switch to GitHub positively affected not just the number of bug reports but also the number of non-developer bug reporters, suggesting an impact in both the intensive (number of bugs reported per person) and the extensive margin (number of reporters) of bug reporting.

\begin{table}
\centering
\caption{Estimated impact of the switch on the number of reporters by 15-day period. Coefficients are highlighted if they are statistically significant with $p<0.1$ (*), $p<0.05$ (**) or $p<0.01$ (***). Standard error is in parentheses.}\label{tab:reporters}
\begin{tabular}{|r|c|c|c|}
\hline
&  Total & Developers & Others \\
\hline
$\mbox{\emph{After switch}}_p$ & 4.15 & 1.24 & 2.91 \\
& (2.52) & (0.974) & (2.46) \\
\hline
$\mbox{\emph{Relative date}}_p * \mbox{\emph{After switch}}_p$ & 0.289 & -0.156* & 0.444** \\
& (0.218) & (0.0843) & (0.213) \\
\hline
$\mbox{\emph{Relative date}}_p$ & 0.111 & 0.159** & -0.0489 \\
& (0.154) & (0.0596) & (0.151) \\
\hline
Constant & 12*** & 3.97*** & 8.04*** \\
& (1.84) & (0.714) & (1.81) \\
\hline
Observation number & 40 & 40 & 40 \\
\hline
\end{tabular}
\end{table}

\begin{figure}
\includegraphics[width=\textwidth]{reporter_nb_rd.eps}
\caption{Number of distinct reporters by 15-day period before and after the migration (with fitting lines from the regression results).} \label{reporter_nb_rd}
\end{figure}

For non-developers, we see a statistically significant increase in the slope on the trend after the switch.
In particular, switching the bug reporting platform to GitHub increases the slope of the regression line by 0.44. This means that after the switch we will, in average, observe an increase of around one non-developer bug reporter each new month.
So for each period, there will be more non-developers who report a bug. This does not necessarily mean that these are people that never reported bugs before. It could also be that non-developer reporters start reporting more and more frequently. Again, this could be due to the increase of the bug reporting appeal, partly because of the platform, but also partly because they receive more feedback on their bug reports (as will suggest our results on comment numbers).

For developers we see a slight but statistically significant (at 10\% level) negative effect on the slope of the trend after the switch. This means that, compared to the slope of the trend before the switch, the new trend has a lower slope. We wouldn't expect a significant change in the number of developers that report bugs because according to our definition there is a fixed number of developers, and they are already very active. Seeing the figure, we see  that the change in slope is due to a stabilization of a previous positive slope, probably meaning a reach of the maximum level of active developers per period.

\subsection{Impact on the number of comments}

Table~\ref{tab:comment_nb} reports on the estimated impact of the bug tracker switch on the number of comments. We can see a statistically significant jump in the number of comments, which almost doubles. This is almost entirely due to comments by developers. This surge in bug comment activity shows that developers communicate more on the bug reports they receive, instead of just fixing and closing them. This means that the switch was beneficial in terms of transparency of the development, an evolution that is likely to give greater satisfaction to the bug reporters.

\begin{table}
\centering
\caption{Estimated impact of the switch on the number of comments. Coefficients are highlighted if they are statistically significant with $p<0.1$ (*), $p<0.05$ (**) or $p<0.01$ (***). Standard error is in parentheses.}\label{tab:comment_nb}
\begin{tabular}{|r|c|c|c|}
\hline
&  Total & Developers & Others \\
\hline
$\mbox{\emph{After switch}}_p$ & 84.7** & 72.2** & 12.5 \\
& (35.5) & (27.9) & (11.1) \\
\hline
$\mbox{\emph{Relative date}}_p * \mbox{\emph{After switch}}_p$ & -2.1 & -2.79 & 0.698 \\
& (3.07) & (2.41) & (0.957) \\
\hline
$\mbox{\emph{Relative date}}_p$ & 1.95 & 1.65 & 0.3 \\
& (2.17) & (1.71) & (0.677) \\
\hline
Constant & 86.6*** & 59.7*** & 26.9*** \\
& (26) & (20.4) & (8.11) \\
\hline
Observation number & 40 & 40 & 40 \\
\hline
\end{tabular}
\end{table}

\subsection{Impact on the number of distinct commentators}

Table~\ref{tab:commentators} and Fig.~\ref{commentator_nb_rd} show the estimated impact of the bug tracker switch on the number of distinct commentators by 15-day period. We can see a highly statistically significant jump in the number of non-developer commentators. This result is particularly interesting because it highlights a remarkable change in the way non-developers engage in the software developing process following the switch. What aspect of the GitHub platform creates this greater engagement remains to be studied but we hypothesize that this is due in large part to GitHub making it easier to subscribe to the bug report activity, and possibly also to the enhanced search.

\begin{table}
\centering
\caption{Estimated impact of the switch on the number of distinct commentators by 15-day period. Coefficients are highlighted if they are statistically significant with $p<0.1$ (*), $p<0.05$ (**) or $p<0.01$ (***). Standard error is in parentheses.}\label{tab:commentators}
\begin{tabular}{|r|c|c|c|}
\hline
&  Total & Developers & Others \\
\hline
$\mbox{\emph{After switch}}_p$ & 7.48*** & 0.763 & 6.72*** \\
& (2.63) & (0.684) & (2.3) \\
\hline
$\mbox{\emph{Relative date}}_p * \mbox{\emph{After switch}}_p$ & 0.238 & -0.0684 & 0.307 \\
& (0.228) & (0.0592) & (0.199) \\
\hline
$\mbox{\emph{Relative date}}_p$ & -0.00977 & 0.0368 & -0.0466 \\
& (0.161) & (0.0419) & (0.141) \\
\hline
Constant & 13.4*** & 7.04*** & 6.41*** \\
& (1.93) & (0.502) & (1.69) \\
\hline
Observation number & 40 & 40 & 40 \\
\hline
\end{tabular}
\end{table}

\begin{figure}
\includegraphics[width=\textwidth]{commentator_nb_rd.eps}
\caption{Number of distinct commentators by 15-day period before and after the migration (with fitted lines from the regression results).} \label{commentator_nb_rd}
\end{figure}

\section{Threats to validity}

\subsection{Internal threats}

The biggest concern that we have in our analysis is the fact that the date of the switch is close to the release of a new version of Coq. The release of version 8.7.0 took place on October 17\textsuperscript{th}, and the migration on October 18\textsuperscript{th}. Then, the hypothesis that there is no other discontinuity around the threshold of interest does not completely hold in our case. First, we can argue that not all our results are threatened: the jump in developer reports isn't because there is usually no peak in developer reports after new releases, same for the jump in non-developer commentators. Second, we can also argue that there should have been a similar (or higher) peak in the month before the switch since the first beta of 8.7 had been released then (and beta releases generally result in higher peaks). Third, we test for robustness by removing the two 15-day periods following the 8.7.0 release (and the switch) from our Regressions in Discontinuity. With these two periods removed, the regressions still give very close results with the same statistical significance.

We also conduct one standard robustness check for Regression in Discontinuity analysis by changing the time window of the analysis to fifteen and to ten periods before and after the switch. Our main results are robust to these changes.

\subsection{External threats}

This is only a case study on a very particular project (it is a software that can be used as a programming language, but requires expertise, that is developed mostly by French researchers, etc.). The only way to confirm the results we obtained would be to replicate this study on other examples of projects that went through a similar bug tracker switch.

\section{Conclusion and perspectives}

We have analyzed the impact of changing the bug reporting environment using a Regression in Discontinuity analysis with data from the switch of the Coq bug tracker from Bugzilla to GitHub.  We have shown that the switch increased the number of bug reports; started an increase of bug reporters per period (i.e. in a certain period there are more distinct reporters); increased the number of comments by developers and the number of distinct commentators per period among non-developers. These results fully validate our first and third hypotheses (increase of the number of bug reports, by developers in particular; improvement of the interaction between reporters and developers) and partially validate our second hypothesis (increase the diversity of bug reporters). We find particularly interesting the change in the number of comments by developers and in the number of distinct non-developer commentators because it shows a significant change in the dynamics of the bug reporting process that we believe can improve software development by engaging more people in the process, and by increasing openness and transparency.

Given the lack of literature on the effects of bug tracking environments, interested researchers could expand our analysis in many different ways. Analyzing other outcomes (such as pull requests or the rapidity of bug resolution) could give interesting additional insights and help create a more complete picture of the mechanisms at play. It will be also interesting to analyze the effects on a longer period to see if the changes hold in the long run. Replicating our analysis for different contexts and software would increase the external validity of the results.

\scriptsize
\paragraph{}
\noindent \footnotesize{\textbf{Supporting data.}}
We provide our extraction and analysis code as a Jupyter notebook and the data we used as CSV files. We also provide the XML export of Bugzilla that was used for the migration and the migration script. Everything is available from \url{https://github.com/Zimmi48/impact-of-switching-bug-trackers}.
\paragraph{}
\noindent \footnotesize{\textbf{Authors' contributions.}}
TZ prepared and conducted the migration, designed and implemented the data extraction process, implemented the data analyses, and wrote part of the manuscript. AC designed the data analyses and wrote part of the manuscript.
\paragraph{}
\noindent\footnotesize{\textbf{Acknowledgements.}}
We would like to thank Ambre Williams for valuable advice and help on the design and implementation of the data analyses and C\'ecile Rottner for general advice, support, and proof-reading.

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{bib}
%
\end{document}

### 2019-03-05 article/article.tex
Article update.

@@
-\documentclass[runningheads]{llncs}
+%\documentclass[runningheads]{llncs}
+\documentclass[conference]{IEEEtran}
+%\IEEEoverridecommandlockouts
 %
 \usepackage{graphicx}
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
+% If possible, figure files should be included in EPS format.
 %
@@
 %
-\title{Impact of switching bug trackers: a case study}
+\title{Impact of switching bug trackers: a case study on a medium-sized open source project}
 %
-%\titlerunning{}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
+\author{Th\'eo Zimmermann\\
+$\Pi R^2$, Inria and
+IRIF, Paris-Diderot University\\
+Paris, France\\
+Email: theo@irif.fr
+\and
+Annal\'i Casanueva Art\'is\thanks{This work was partly funded by El Conacyt of the Mexican government.}\\
+Paris School of Economics\\
+Paris, France}
 %
-\author{Th\'eo Zimmermann\inst{1,2} \and
-Annal\'i Casanueva Art\'is\inst{3}\thanks{This work was partly funded by El Conacyt of the Mexican government.}}
-%
-\authorrunning{T. Zimmermann and A. Casanueva Art\'is}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{$\Pi R^2$, Inria, Paris, France \and
-IRIF, Paris-Diderot University, Paris, France\\
-\email{theo@irif.fr} \and
-Paris School of Economics, Paris, France}
-%
-\maketitle              % typeset the header of the contribution
+\maketitle
 %
 \begin{abstract}
-Any significant software project requires a bug tracker.
-In open source development, this plays an even bigger role.
-The impact of the bug reporting environment on the bug tracking activity is difficult to assess.
-In this paper, we take advantage of the switch from Bugzilla to GitHub of the bug tracker of Coq, a medium-sized open source project, to evaluate the impact that such a change can have.
-We first report on the switch itself, in particular the migration of 4900 preexisting bug reports using the GitHub issue import API; then we analyze data from before and after the migration.
-We show that the switch induces
-an increase in bug reporting, particularly from principal developers themselves, and more generally an increased engagement with the bug tracking platform, with more comments by developers and also more external commentators.
+For most software projects, the bug tracker is an essential tool.
+In open source development, this tool plays an even more central role as it is generally open to all users, who are encouraged to test the software and report bugs.
+Previous studies have highlighted the act of reporting a bug as a first step leading a user to become an active contributor.
+
+The impact of the bug reporting environment on the bug tracking activity is difficult to assess because of the lack of comparison points.
+In this paper, we take advantage of the switch, from Bugzilla to GitHub, of the bug tracker of Coq, a medium-sized open source project, to evaluate the impact that such a change can have.
+
+We first report on the switch itself, in particular the migration of 4900 preexisting bug reports using the GitHub issue import API.
+Then we analyze data from before and after the switch using a Regression on Discontinuity analysis, a novel methodology in the context of empirical software engineering.
+We show that the switch induces an increase in bug reporting, particularly from principal developers themselves, and more generally an increased engagement with the bug tracking platform, with more comments by developers and also more external commentators.
+\end{abstract}
 
@@
 
-\keywords{bug tracker \and switch \and migration  \and bug report \and issue \and GitHub \and Bugzilla \and open source.}
-\end{abstract}
+\begin{IEEEkeywords}
+bug tracker, switch, migration, bug report, issue, GitHub, Bugzilla, open source, data mining, Regression on Discontinuity
+\end{IEEEkeywords}
 %
@@
 
-Bug reporting is an essential part of software development. However, it involves
-a non-negligible human component, particularly in the context of an open source project where a larger part of bugs are reported by independent users and therefore depend a lot on the good will of individuals.
-Creating an environment that makes bug reporting easier and more appealing may lead to an increase of bug reporting (from both developers and independent users) which should ultimately benefit  software development \cite{Bettenburg2008} (provided that developers are equipped to handle the incoming reports \cite{Anvika,Kanwal2012}).
-However, the impact of such change has rarely been studied.
-Indeed, comparing the number of bug reports in different environments is generally not compelling because it is likely to depend  not just on the reporting environment but also on each particular software and its development dynamics.
-To do a valid comparison one would ideally need to compare two bug reporting environments for the same software, in the same period.
-But, or course, we can't observe this ideal.
+Bug reporting is an essential part of software development.
+In the context of an open source project, the bug reporting and fixing process is generally done on a public bug tracking platform, and, in the absence of paid testers, it depends a lot on the good will of independent users.
+Therefore, it involves a strong social component.
+
+However, having an open bug tracking system is not enough to get the bugs reported.
+In addition to the software having enough users, the process of bug reporting must be easy and appealing.
+Therefore, creating an environment that eases bug reporting and makes it more appealing may lead to an increase in bug reporting activity (from both developers and independent users), which is well associated with an increase in software quality.
+First, if the software has not changed, receiving more bug reports indicates that more defects have been found, and not a sudden apparition of defects (Linus's Law ``given enough eyeballs, all bugs are shallow''~\cite{raymond1999cathedral,van2009shallow}).
+Moreover, assuming that developers are equipped to cope with the increased number of incoming reports~\cite{anvik2005coping, davidson2011coping}, even duplicate reports are not necessarily harmful~\cite{Bettenburg2008}.
+Second, more activity on the bug tracker can also mean users are helping to reproduce bugs, produce traces, etc.
+and thus are working with the developers to get the bugs fixed~\cite{breu2010information}.
+More generally, reporting new bugs and discussing existing bug reports has been shown to be an important step on the path to becoming an active contributor of an open source project~\cite{jensen2007role, nakakoji2002evolution, ye2003toward}.
+
+Despite the importance of the bug tracking environment that we have just highlighted, the impact of a change in this environment has rarely been studied.
+It is generally difficult to compare the number of bug reports in different environments because it is likely to depend  not just on the reporting environment but also on each particular software and its development dynamics, which would make such an analysis not compelling.
+To do a valid comparison, one would ideally need to compare two bug reporting environments for the same software, used by the same users, to report the same bugs, in the same period.
+This is of course impossible.
 A context approaching this ideal scenario would be to compare the number of bug reports of a particular software before and after the switch to a new environment.
 Yet, substantial changes in the environment of bug reporting (e.g.
-changes of bug tracker platform) are rare because once a bug system is in place, any change comes with a significant cost.
+changes of bug tracker platform) are rare because once a bug tracking platform is in place, any change comes with a significant cost.
 
-We use the switch, from Bugzilla to GitHub, of the bug tracker system of the Coq proof assistant, a medium-sized open source software, to analyze the  effects of a change in the bug reporting environment on the number of bugs, reporters, bug comments and commentators.
-We describe the switch event and the associated migration process for preexisting bugs.
-Thanks to the way this migration was conducted, getting data on bug reporting from before and after the switch is easily done using the GitHub GraphQL API.
-Using a Regression in Discontinuity analysis we find strong evidence that the switch of platform increased the number of bug reports; started an increase of bug reporters per period (i.e.
-in a certain period there are more distinct reporters); increased the number of comments by developers and the number of distinct commentators per period among non-developers.
-Taken together our results suggest that switching from Bugzilla to GitHub increased the number of bug reports and bug reporters (probably by making the bug reporting process easier) but also increased the openness of software development and aroused discussion around bugs (notably between developers and non-developers).
+In this paper, we study the case of the Coq proof assistant, a medium-sized open source software project.
+We use the switch, from Bugzilla to GitHub, of the project's bug tracker, to analyze how a change in the bug reporting environment affects the number of bugs, reporters, bug comments and commentators.
+Using a Regression on Discontinuity design (RDD)~\cite{angrist2008mostly,angrist2014mastering,jacob2012practical,lee2010regression}, we find strong evidence that the change of platform increased the number of bug reports and comments by developers, and increased the number of distinct commentators each week among non-developers.
+Our results suggest that switching from Bugzilla to GitHub made the bug reporting process easier and more appealing but also opened the software development to more external contributors and stimulated discussion around bugs (notably between developers and non-developers).
+
+The contributions of this paper are practical, empirical, and methodological.
+First, we improved an existing bug tracker migration tool to handle thousand of bug reports while preserving meta-data, and we helped the Coq development team migrate their bug tracker from Bugzilla to GitHub.
+Second, we analyze the causal impact of this switch on the bug tracking activity.
+Third, we introduce to the field of software engineering a state-of-the-art method for quantitative public policy evaluation: Regression in Discontinuity Design.
+In particular, we explain this method to give the basic intuition behind; we point the reader to the both introductory and state-of-the-art literature and we illustrate one application.
+
+The rest of this paper is organized as follows: In Sec.~\ref{related-work}, we discuss related work.
+In Sec.~\ref{context}, we present the Coq project and the context of the switch.
+In Sec.~\ref{hypotheses}, we formulate hypotheses on the impact of the switch.
+In Sec.~\ref{migration}, we explain the migration process.
+In Sec.~\ref{data}, we describe the data collection and pre-processing method, and we list the variables of interest.
+In Sec.~\ref{descriptive-stats}, we give a preliminary view of the bug tracking activity.
+In Sec.~\ref{methodology}, we explain the Regression on Discontinuity methodology, of which we present the results in Sec.~\ref{results}.
+We interpret the results and relate them to our hypotheses in Sec.~\ref{discussion}.
+We discuss the threats to validity and  present robustness checks in Sec.~\ref{threats}.
+We conclude in Sec.~\ref{conclusion}.
+
+\section{Related work}
+
+\label{related-work}
+
+While there is a very large literature on many aspects of bug reporting (see~\cite{strate2013literature, zhang2016literature} for literature reviews), there is a lack of literature measuring the influence of the bug reporting environment on the bug reporting activity or on any other aspects of software development.
+To our knowledge, there is no previous work measuring the impact of a change in bug reporting environment.
+More generally, there is little literature that compare bug trackers.
+
+\subsection{Comparing and proposing features of bug trackers}
+
+A simple, but quite limited, way to compare bug trackers is to look at the features that each of them proposes.
+
+Karre et al.~\cite{karre2017does} compared 31 bug trackers feature-wise, and gathered these tools in four clusters.
+Bugzilla belongs to the cluster of bug trackers with many features (together with RedMine and Mantis), while GitHub belongs to a cluster of bug trackers attached to source code management systems (together with Savannah and BitBucket).
+%Bugzilla has 20 of the 23 features they listed, and scores the highest number of features (ex.
+aeqo with RedMine).
+They are closely followed by Mantis, and two other bug trackers, with 19 features.
+GitHub only has 15 features out of the 23 listed (overall 11 listed systems have more features).
+
+In the case of the Coq project, the loss of some specific features from Bugzilla was one of the arguments against the switch.
+Nonetheless, GitHub has some features that Bugzilla hasn't, and it is clearly not the number of features that determines which software is better or more adapted to the specific use case.
+
+%We will note that GitHub has very recently added a number of new features to the issue tracker, but these were not present at the time of the switch (just-in-time duplicate retrieval, issue deletion, transfer and pinning, support for multiple issue templates, etc.).
+
+Additionally, Karre et al.
+recommended some missing features, including better support for migrating bug tracking data from a system to another.
+Our work has contributed to improve the situation in that sense.
+Nonetheless, it would be preferable if there existed a standard structured format for storing bug tracking information and all major bug tracking systems supported exporting to and importing from this format.
+
+Abaee and Guru~\cite{abaee2010enhancement} listed the features of four commercial bug trackers, and proposed their own bug tracker with some new features, but without any evaluation.
+
+There is much more literature proposing new features to add to bug trackers, but they are rarely evaluating the impact of adding such features on the bug tracking activity, even when a prototype was presented.
+
+Baysal et al.~\cite{baysal2013situational} identified the need for personalized issue tracking systems after interviewing twenty Mozilla developers.
+They proposed a Bugzilla extension to address this need, and gave a, mostly qualitative, assessment by interviewing developers using their tool in~\cite{baysal2014no}.
+
+Bortis~\cite{bortis2016porchlight} developed a bug triaging tool and evaluated how users interacted with it.
+However, there was no evaluation of its impact in the context of an actual software project.
+
+Just et al.~\cite{just2008towards} surveyed developers from three large open source projects (Apache, Eclipse, and Mozilla; all three projects use Bugzilla) to identify missing features and provided recommendation to design better bug tracking systems.
+%They remark that bug trackers are often mere interfaces to bug databases and ask too much information from end-users without guiding them.
+They recommend providing guidance to bug reporters depending on their level of expertise.
+This issue is indeed something that we have noticed with Bugzilla and other bug trackers of the same category.
+Users are encouraged to self-triage and often do so inaccurately.
+On the contrary, GitHub only provides title and body fields to bug reporters and leaves triaging (labeling) to the contributors with write access.
+Nonetheless, the issue templates can be used to fill the bug report body field with advice on how to present a bug report.
+GitHub has recently added support for multiple issue templates which would allow to let the reporter choose between, for instance "My first bug report", "My first feature request", "Bug report (experienced reporter)", and "Feature request (experienced reporter)", with varied requirements and levels of explanations...
+
+Our article goes beyond a descriptive or normative perspective and quantitatively measures the causal impact of the change of the bug tracking platform (a crucial part of the bug tracking environment) on various aspects of bug reporting activity.
+
+\subsection{Analysis of projects' bug tracking data}
+
+Another way of comparing bug tracking systems could have been to conduct large-scale studies of many software projects using various bug trackers and provide some clues of the differences induced by the use of the different bug tracking systems.
+However, there are no such comparative studies in the literature.
+
+Sowe et al.~\cite{sowe2013multi} noted that most preceding literature had only been comparing few projects at a time, usually using a single bug tracking system.
+In their study, they addressed this in part by studying hundreds of projects, but they did not mention which bug trackers the projects they compared used.\looseness=-1
+
+Bissyandé et al.~\cite{bissyande2013got} published the same year a larger scale investigation using ten thousand projects' bug tracking data, analyzing such things as the correlation between a project's success and its bug tracking activity.
+All of the projects studied were using the same bug tracker (GitHub's).
+
+Francalanci and Merlo~\cite{francalanci2008empirical} analyzed the bug fixing process by studying closed bugs from nine open source projects (four using JIRA, four using SourceForge's bug tracker, and one using another bug tracker).
+They did not, however, try to correlate the bug tracking activity with the bug tracking system used.
+
+%According to the survey in~\cite{karre2017does}, Bugzilla is the most used bug tracker (respondent were 32\% to use Bugzilla and 15\% to use GitHub).
+There is an obvious selection bias due to the choice of open source project contributors this survey was sent to.
+Overall, given the sheer number of projects hosted on GitHub, it does not make any doubt that the GitHub issue tracker is the most used bug tracker in number of projects.
+But this raw fact does not matter much, given that very small projects will use their bug tracker very little.
+\subsection{Switching developer tools}
+
+While, to the best of our knowledge, there is no work evaluating the impact of switching bug trackers, some studies have been conducted to evaluate the impact of switching other developer tools.
+Since bug trackers are important developer tools, our article also relates and contributes to this literature.
+
+De Alwis and Sillito~\cite{de2009software} surveyed developers to understand the perceived impact of moving from centralized to decentralized version control systems.
+As with bug trackers, the move is decided because of anticipated benefits: among them is the openness to external contributors.
+But the switch incurs challenges in the migration process (in particular, developers are very concerned to preserve the full history of the project) and some features are lost in the new tool (in particular, the monotonically increasing revision numbers).
+Muşlu et al.~\cite{mucslu2014transition} performed a similar study based on interviews and a survey.
+They identified expectations and barriers and provided recommendations to developers wanting to conduct such a migration.
+
+Squire~\cite{squire2015should} measured the impact of switching developer support channels from mailing lists and self-hosted forums to Stack Overflow, by first identifying the expectations of 20 projects which had officially decided on such a move, then by studying the impact on developer participation and response time for a selection of four such projects.
+On the other hand, Vasilescu et al.~\cite{vasilescu2014social} compared the behavior of the same users on the R mailing list and the Stack Overflow platform.
+Contrarily to a bug tracker which is expected to be centralized for a given project, support channels can be diverse.
+Project leaders can decide to abandon their mailing list to focus exclusively on Stack Overflow, but they do not decide to ``open'' a support channel on Stack Overflow: it can exist nevertheless and the switch can be organic.
 
@@
 
+\label{context}
+
 \subsection{Coq and GitHub}
-Coq is a (free and open source) proof assistant, that is a software to write and automatically verify mathematical proofs (and programs).
+Coq is a (free and open source) proof assistant, i.e., a software to write and automatically verify mathematical proofs (and programs).
 It originated as a research project in 1984 and the development team is still composed, for the most part, of researchers.
 It gained in popularity along the years, and is now at the center of many verification projects (some very large academic projects, and some industrial projects).
-Its initial developers were awarded  the ACM SIGPLAN Programming Languages Software 2013 Award and the ACM Software System 2013 award.
+Its initial developers were awarded  the ACM SIGPLAN Programming Languages Software 2013 award and the ACM Software System 2013 award.
+
+% XXXX justify more the medium-size qualification
 
 The development is currently centered around the GitHub platform.
-The current repository of Coq the current repository counts more than 25000 commits.
+The Coq repository counts more than 25,000 commits.
 Some of these commits date back from 1999, when the code received major architectural changes.
 The GitHub repository was initially created as a mirror of the official repository (which itself was migrated from SVN to git in 2013) and really started attracting the interest of the development team as a way to open up the development to external contributions, especially starting with the first Coq coding sprint in 2015 (later renamed into the Coq Implementors Workshop and held every year in France).
-From then on, core developers started using pull requests for their own changes more and more often, and it became the norm around the beginning of 2017 when systematic continuous integration tests were introduced on every pull requests.
+From then on, core developers started using pull requests for their own changes more and more often, and it became the norm around the beginning of 2017 when systematic continuous integration tests were introduced on every pull request.
 In the last two years, more than 2,200 pull requests have been opened.
@@
 From 2007 to 2017, the Coq bug tracker platform was a self-hosted Bugzilla instance (before this, from 2001 to 2007 it was using JitterBug, a now discontinued bug tracking system, and before 2001 a simple mailing list).
-In 2017, developers had become quite accustomed to the look and feel of discussion threads on GitHub pull requests, which offered a much nicer experience than the Bugzilla bug tracker.
-Younger developers (in particular, one of the author of this article, who eventually pushed for the switch and conducted the migration) were frustrated by what they perceived as an older and heavier platform.
+In 2017, developers had become quite accustomed to the look and feel of discussion threads on GitHub pull requests, which offered an arguably nicer experience than the Bugzilla bug tracker.
+Younger developers were frustrated by what they perceived as an older and heavier platform.
 The rest of the developers were convinced to migrate towards a non-self-hosted solution in August 2017, when the whole Coq website was shut down for more than a week  because an IT team was frightened by a scam bug report.\footnote{\url{https://sympa.inria.fr/sympa/arc/coq-club/2017-08/msg00040.html}}
@@
 \begin{itemize}
-\item Better formatting, ability to edit comments, and more generally a more pleasant tool to use. We felt that Bugzilla was unpleasant enough to discourage developers to use it to its full potential.
-\item Browsing between code, pull requests and bug reports is easier when everything is in the same place. Furthermore, it means a common notification, mention, and assignment system.
+\item Better formatting, ability to edit comments, and more generally a more pleasant tool to use.
+According to some Coq developers, Bugzilla was unpleasant enough to discourage developers to use it to its full potential.
+\item Version control integration: Browsing between code, pull requests and bug reports is easier when everything is in the same place.
+Furthermore, it means a common notification, mention, and assignment system.
 \item Better indexing by search engines.
@@
 
-The possible drawbacks were mostly that GitHub provides a less-advanced system with very few customization options. In particular:
+The possible drawbacks were mostly that GitHub provides a less advanced system with very few customization options. In particular:
 \begin{itemize}
@@
 That's why the Coq GitHub repository now only has protected branches (and accidentally-pushed topic branches are promptly deleted).
-\item Creating a new bug report doesn't require filling a form, but typing a message.
+\item Creating a new bug report does not require filling a form, but typing a message.
 Fortunately, issue templates help give some hints towards the expected structure of the report, and being able to strip this template can actually give flexibility to developers who might open new issues for concerns that do not qualify as bugs.
@@
 
-Given the advantages of the GitHub issue tracker that were listed above, we expected the switch to:
-\begin{itemize}
-\item increase the number of bug reports, by developers in particular;
-\item increase the diversity of bug reporters;
-\item improve the interaction between reporters and developers.
-\end{itemize}
+\label{hypotheses}
+%ambre: I think this section should be expanded a bit to add motivation for the hypotheses
+
+Given the advantages of the GitHub issue tracker that were listed above, one could expect the switch to:
+\begin{description}
+\item[H1] Increase the bug tracking activity, in terms of the number of reported bugs;
+\item[H2] Extend the pool of bug reporters and commentators;
+\item[H3] Increase the bug tracking activity, in terms of the amount of interaction between users and developers.
+\end{description}
 
@@
 because the number is already taken by a GitHub pull request) are postponed and renumbered.
-We implemented several improvements to make the tool better fit our needs:
+We implemented several improvements to make the tool better fit the needs of the Coq project:
 \begin{itemize}
-\item \textit{Allowing non-consecutive bug numbers:} Our imported set of bug reports had some holes in the numbering due to deleted bugs.
+\item \textit{Allowing non-consecutive bug numbers:} The imported set of bug reports had some holes in the numbering due to deleted bugs.
 We use postponed bugs to fill the holes.
@@
 \item \textit{Using the GitHub issue import beta API and overcoming the GitHub rate limits:} Creating a new issue or a new comment through the normal GitHub REST API will trigger notifications (for people who are watching the repository or are mentioned in the issue thread).
-Such actions are therefore subject to a strict rate limit which prevented using this tool for importing more than a few hundred bugs.
+Therefore, GitHub chooses to impose a strict rate limit on these actions, which prevented using this tool for importing more than a few hundred bugs.
 Fortunately, GitHub provides a (beta) issue import API which, in addition to not triggering notifications, also allows importing one bug, its comments, and meta information such as closed status and assignee in a single request (thus reducing the duration to import 4900 bugs to just a few hours).
-Finally, using this API allows to keep the dates of imported bug reports and comments, which is very useful to this study.
-We didn't manage to import correct closing dates, so we won't study the impact of the bug tracker switch on the time it takes to close a bug.
+Furthermore, using this API allows to keep the dates of imported bug reports and comments, which is very useful to this study.
+We didn't manage to import correct closing dates for every bug report, so we will not study the impact of the bug tracker switch on the time it takes to close a bug.
+
+This improvement has now been integrated upstream as an optional setting.
 \end{itemize}
@@
 This tool required a correspondence table between Bugzilla and GitHub accounts.
-Ours was created by manually matching 217 (out of 686) Bugzilla accounts to 175 GitHub accounts,\footnote{The correspondence table can be seen at \url{https://gist.github.com/Zimmi48/d923e52f64fe17c72852d9c148bfcdc6##file-bugzilla2github-L47-L264}.} the difference being due to the high prevalence of duplicate Bugzilla accounts.
+Ours was created by manually matching 217 (out of 686) Bugzilla accounts to 175 GitHub accounts,\footnote{The correspondence table can be seen at \url{https://frama.link/vRV9etCd}.} the difference being due to the high prevalence of duplicate Bugzilla accounts.
 We gave priority to finding GitHub accounts for people who still had opened bug reports.
-The most efficient way to do the matching was by searching the name of the reporter + ``github'' on Google.
-We also sent e-mails to people we hadn't managed to match (and who still had opened bugs).
+%The most efficient way to do the matching was by searching the name of the reporter + ``github'' on Google. We also sent e-mails to people we hadn't managed to match (and who still had opened bugs).
 
-\paragraph{Urgency.}
-Once the switch was approved by the development team on October 3-4\textsuperscript{th}, 2017, the migration had to happen as soon as possible because every new pull request before the migration added to the number of bug reports that would need to be renumbered.
-It was conducted\footnote{More details on the process are available in a blog post at \url{http://theoz.im/bugzilla}.
-} on October 18\textsuperscript{th},  the day after the 8.7.0 release (and not before to avoid disturbing the release process).
-Only 502 bugs (whose numbers were below 1154) had to be renumbered.
+Once the switch was approved by the development team on October 4\textsuperscript{th}, 2017, the migration had to happen as soon as possible because every new pull request before the migration added to the number of bug reports that would need to be renumbered.
+It was conducted %\footnote{More details on the process are available in a blog post at \url{http://theoz.im/bugzilla}.
+}
+on October 18\textsuperscript{th},  the day after the 8.7.0 release (and not before to avoid disturbing the release process).
+Only 502 out of 4900 bugs (whose numbers were below 1154) had to be renumbered.
 This number is quite low because these bugs are dating back from the JitterBug era and during the previous switch only the bugs that were still open had been migrated.
+Due to the rate of pull request creation, if the switch had been delayed by a year, the number of renumbered bugs would have been closer to 3000.
 
 \section{Data}
+\label{data}
 
@@
 
-All the data for this study was extracted on September 14\textsuperscript{th}, 2018 using the GitHub GraphQL API.
+All the data for this study was extracted on January 17\textsuperscript{th}, 2019 using the GitHub GraphQL API \cite{github_graphql_API}.
 Using this API allows to do large requests (100 nodes in a single request) with just the information we need (thus both reducing the bandwidth usage and speeding up the extraction process).
-We provide the extracted data as CSV files and a Jupyter notebook\footnote{Available at \url{https://github.com/Zimmi48/impact-of-switching-bug-trackers}.} with the code to request this data from GitHub, to load the CSV files, and to run the analyses.
+We provide the extracted data as CSV files and a Jupyter notebook\footnote{Available at \url{https://github.com/Zimmi48/github-migration-analysis}.} with the code to request this data from GitHub, to load the CSV files, to run the pre-processing steps, and the analyses.%, and the robustness checks.
 
-Migrated bugs and comments had their author information stored in the text so we extract it from there.
-Dates of bug and comment creation have been preserved so we can obtain them transparently for bugs before and after the migration.
-Comments dating from the JitterBug era do not have valid author information (and closed bug reports from this era have not been migrated) so we won't consider any data from before 2008.
+Migrated bugs and comments have their author information stored in the text, we extract it from there.
+Dates of bug and comment creation have been preserved which allows us to obtain them transparently for bugs before and after the migration.
+Bug reports and comments from the JitterBug era are missing information and were not fully migrated, therefore we do not consider any data from before 2008.
 
@@
 
-\paragraph{Excluding specific reporters.}
+\paragraph{Excluding specific reporters}
 We have one specific reporter, Jason Gross, who is alone responsible for almost a quarter of all bug reports.
-To avoid having the behavior of a single individual strongly impact the overall statistics, we exclude his comments, bug reports and pull requests and all the comments they received from our analysis.
-Similarly, we also exclude the developer who was the initiator and conductor of the bug tracker switch and migration and who is an author of this article.
+To avoid having the behavior of a single individual strongly impact the overall statistics, we exclude his comments, bug reports and the comments they received from our analysis.
+Similarly, we also exclude the developer who was the main advocate of the bug tracker switch as it could have influenced his behavior before and after the switch.
 
-\paragraph{Merging duplicate Bugzilla accounts.}
-This is only relevant to the new reporter analysis which itself is only detailed in the accompanying Jupyter notebook, so refer there for more information.
+\paragraph{Merging duplicate Bugzilla accounts}
+We found that a significant number of users had created several accounts on the Bugzilla system.
+We merged them to avoid overestimating the rate of new reporters which is analyzed in the companion Jupyter notebook.
 
-\paragraph{Removal of migration artifact comments.}
+\paragraph{Removal of migration artifact comments}
 The migration tool created an issue whose body contains only meta-information, and a first comment with the description, for every bug report.
@@
 
-Our outcome variables will be the numbers of bug reports, distinct bug reporters, new bug reporters (who had never reported a bug before), comments and distinct commentators, for a given period.
-We will use 15-day periods to have sufficiently many bugs by period.
-Indeed, in our data we have an average of 13 bug reports by 15-day period; 50\% of the periods have between 7 and 18 reports; 90\% of the periods have between 4 and 27.
-This also allows to have 40 observations centered around the date of the switch for our Regression in Discontinuity analyses.
+We measure different indicators of bug tracker activity: the numbers of bug reports per day; the number of distinct bug reporters in a week; the number of new bug reporters (who had never reported a bug before) per day; the number of comments per day; the number of distinct commentators in a week; and the number of new commentators per day.
+The number of distinct bug reporters and commentators is intended to allow us to distinguish between having a few prolific contributors and a large base of casual contributors.
+We use an interval of a week instead of a day for measuring the number of distinct bug reporters and commentators because, at the scale of a day, there is less opportunity for repeated contributions by the same contributor, but longer scales would compromise the estimation of effects by removing too much data.
 
-We will analyze heterogeneous effects by distinguishing between developers and others (others being assimilated to ``users'').
+For the visualization of Fig.~\ref{bug_nb_rd} to~\ref{commentator_nb_rd}, we plot a point for the average of each two week period.
+This is just to allow an easier and clearer visual analysis of the discontinuity.
+The regression lines, however, correspond to the definitions detailed in the above paragraph (i.e., analysis with daily periods or weekly periods depending on the variable).
+
+We analyze heterogeneous effects by distinguishing between developers and other contributors (others being assimilated to ``users'').
 We define ``developers'' as the persons who have contributed more than 100 commits since 2008.
+We identified 18 developers.
+These developers are responsible for 91.5\% of all commits since 2008 (this is consistent with standard results on the proportion of commits by the ``core team'' in open source software~\cite{robles2009evolution}).
 
-When studying the influence of new releases, we define the release dates by the dates of the corresponding news items on the official website (\url{https://coq.inria.fr/news}).
+For the dates of the new releases in Section~\ref{descriptive-stats}, we define release dates as the dates of the release announcement on the mailing list or of the corresponding news item on the official website (\url{https://coq.inria.fr/news}), whichever comes first.
 We selected beta and major releases and excluded release candidates and patch-level releases.
-The first release to appear in the news section of the current official website is the 8.2beta release in June, 2008.
+The first release to appear in the news section of the current official website is the 8.2 beta release in June, 2008.
 
 \section{Descriptive statistics}
+\label{descriptive-stats}
 
-The accompanying Jupyter notebook contains lots of descriptive statistics and figures, such as the distribution of bug reporting in the year, and in the day, graphs of bug reporting for individual developers, etc.
-We only present a subset of these figures.
+The companion Jupyter notebook contains an extensive array of descriptive tables and figures, such as the distribution of bug reporting in the year and in the day, graphs of bug reporting for individual developers, etc.
+We only present a subset of these results.
 
@@
 
-Fig.~\ref{cumulative_bugs_and_comments} gives a first (cumulative) view of two of our outcome variables: the number of bug reports, and the number of comments.
-From this figure, we can clearly notice the dominance of developers in the number of bug comments and the dominance of non-developers in the number of bug reports.
+Fig.~\ref{cumulative_bugs} and~\ref{cumulative_comments} give a first (cumulative) view of two of our outcome variables: the number of bug reports, and the number of comments.
+From these figures, we can clearly notice the dominance of developers in the number of bug comments and the dominance of non-developers in the number of bug reports.
 We can also already notice some slope changes at the bug tracker switch date.
-The magnitude and significance of this changes will be assessed in the remainder of this article.
+The magnitude and significance of this changes will be assessed in Sec.
+\ref{results}.
 
 \begin{figure}
-\includegraphics[width=\textwidth]{cumulative_bugs_and_comments.eps}
-\caption{Cumulative number of bug reports and comments (since 2008).} \label{cumulative_bugs_and_comments}
+\includegraphics{cumulative_bugs.eps}
+\caption{Cumulative number of bug reports (since 2008).} \label{cumulative_bugs}
+\end{figure}
+
+\begin{figure}
+\includegraphics{cumulative_comments.eps}
+\caption{Cumulative number of comments (since 2008).} \label{cumulative_comments}
 \end{figure}
@@
 
-\paragraph{Influence on the number of bug reports.}
+\paragraph{Influence on the number of bug reports}
 We expect release dates to be correlated to peaks of activity on the bug tracker, in particular peaks of new bug reports.
 Indeed, users will generally test new releases, for instance by trying to port preexisting projects to the new version, which might lead them to find new bugs (either regressions or bugs related to new features).
-In Fig.~\ref{bug_nb_with_releases}, we can see that while release dates are generally correlated to peaks of new bug reports by non-developers, this isn't so much the case of reports by developers.
-We can also notice that beta releases are often correlated to higher peaks than final releases which isn't surprising given that the users who test beta versions are more likely to be ready to report bugs, and beta versions are likely to be less polished.
+In Fig.~\ref{bug_nb_with_releases}, we can see that while release dates are generally correlated to peaks of new bug reports by non-developers, this is not so much the case of reports by developers.
+We can also notice that beta releases are often correlated to higher peaks than final releases which is not surprising given that the users who test beta versions are more likely to be ready to report bugs, and beta versions are likely to be less polished.
 
@@
 Consequently, a different approach was taken for the following releases: versions 8.6, 8.7 and 8.8 were ``developed on a time-based development cycle'' and contain ``the result of refinements, stabilization of features and cleanups of the internals of the system along with a few new features''.
-Attention was given to regression testing by systematically testing changes by building a selection of the largest Coq projects.
+Attention was given to regression testing: changes were systematically tested by building a selection of the largest Coq projects.
 Therefore, it is not surprising that these frequent releases are correlated to smaller peaks of new bug reports.
 
+%IMPORTANT: metre une phrase dans la partie description des variables qui indique qu'on fait 4 semaines pour celui la.
 \begin{figure}
-\includegraphics[width=\textwidth]{bug_nb_with_releases.eps}
-\caption{Number of bug reports by 15-day period with release dates (since 2008).} \label{bug_nb_with_releases}
+\includegraphics{bug_nb_with_releases.eps}
+\caption{Number of bug reports by 4-week period with release dates (since 2008).} \label{bug_nb_with_releases}
 \end{figure}
 
-\paragraph{Influence on the number of comments.}
+\paragraph{Influence on the number of comments}
 In Fig.~\ref{comments_with_releases}, we can see that release dates are correlated to peaks of comments by developers (and not so much by other people).
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{comments_with_releases.eps}
-\caption{Number of bug comments by 15-day period with release dates (since 2008).} \label{comments_with_releases}
+\includegraphics{comments_with_releases.eps}
+\caption{Number of bug comments by 4-week period with release dates (since 2008).} \label{comments_with_releases}
 \end{figure}
@@
 \section{Methodology}
+\label{methodology}
 
-We exploit the fact that the switch to GitHub can be seen as a clear cutoff in time to use a Regression in Discontinuity approach to estimate the effects of the switch on our outcome variables, to determine if these estimated effects are statistically significant, and to interpret these effects in a causal way.
-The idea behind this analysis is that bug reporters did not choose the exact moment in which the switch would take place.
-In the absence of any other discontinuity, the changes in the behavior of bug reporters just before and just after the migration is likely to depend mainly on the switch even if there was a change in behavioral trends before the switch.
-It is important to note that estimated effects using this method will inform of the effect near the threshold.
+We exploit the fact that the switch from Bugzilla to GitHub can be seen as a clear cutoff in time to use a temporal sharp Regression in Discontinuity design (RDD) to estimate the effects of the switch on our outcome variables; to determine if these estimated effects are statistically significant; and to interpret these effects in a causal way.
+The idea behind RDD is that factors that influence bug reporting activity are fairly similar just before and just after the cut-off point, which makes the changes observed between just after and just before the cut-off a consequence of the change that occurred at the cut-off (i.e., the switch of bug reporting platform).
+
+The assumption behind this analysis is that bug reporters did not choose the exact moment in which the switch would take place.
+In the absence of any other discontinuity, the changes in the behavior of bug reporters just before and just after the migration is likely to depend mainly on the switch: the other factors influencing bug reports evolve slowly, and should be pretty similar just before and just after the switch.
+It is important to note that estimated effects using this method will inform of the effect near the threshold (i.e., it will only give the Local Average Treatment Effect --LATE-- of the switch).
 Indeed, in the same way that behavior just before and just after the threshold is not likely to be different in the absence of the switch, the behavior several periods before and several periods after is likely to be different even in the absence of the switch.
-See~\cite{lee2010regression} for more details on Regression on Discontinuity analysis.
+See~\cite{angrist2014mastering} and~\cite{angrist2008mostly} for an accessible and intuitive introduction to Regression on Discontinuity analysis; and~\cite{lee2010regression} and~\cite{jacob2012practical} for further details and a practitioner's guide to its empirical application.
 
-We use time as running variable and we estimate the following regression:
+We model the evolution of the bug reporting behaviour around the switch by two different functions, one before and one after the switch, fitted using ordinary least squares.
+Their purpose is to accurately estimate the value at the switch in the absence and presence of treatment.
+Since only their value around the cutoff point is relevant, we can choose to estimate them on a restricted bandwidth (number of periods) around the cutoff, or on a larger dataset.
+We also have the choice of different functional forms for these models.
 
+In the context of a Regression in Discontinuity analysis, the choice of the bandwidth of the analysis and the functional form of the regression is always difficult.
+On the one hand, including data points that are far from the cut-off (in our case the switch) will increase precision as it will reduce standard errors (because we will have more data points).
+On the other hand, however, including those points can bias estimation near the cut-off if the functional form of the evolution in the absence of switch is not accurately specified.
+Our main specification takes a conservative approach with a relatively small bandwidth of 180 days before and after the switch to minimize possible bias, and a comparatively simpler linear model.
+However, as a robustness check, we also estimate a quadratic model on a larger time frame (455 days on each side).
+This conservative approach reduces the statistical power of our analysis, increasing the probability of observing false negatives (being unable to detect an effect where such an effect exists).
+
+We use time as rating variable\footnote{The rating variable is also sometimes called the running variable or the forcing variable in the literature.} (with the cutoff centered at zero) and we estimate the following regression:
 \begin{equation*}
-\begin{aligned}
-\mbox{\emph{Number of bug reports}}_p⁡ =\gamma_0 &+\gamma_1 \mbox{\emph{Relative date}}_p + \gamma_2 \mbox{\emph{After switch}}_p \\
-&+ \gamma_3 \mbox{\emph{Relative date}}_p * \mbox{\emph{After switch}}_p + \epsilon_p
-\end{aligned}
+\arraycolsep=0pt
+\begin{array}{l}
+\mbox{\emph{Number of bug reports}}_t = \\
+\quad\quad \gamma_0 +\gamma_1 \times \mbox{\emph{Relative date}}_t
++ \gamma_2 \times \mbox{\emph{After switch}}_t \\
+\quad\quad +~\gamma_3 \times \mbox{\emph{Relative date}}_t \times \mbox{\emph{After switch}}_t
++ \epsilon_t
+\end{array}
 \end{equation*}
+\noindent where $\mbox{\emph{Number of bug reports}}_t⁡$ is the total number of bugs reported during the day $t$; $\mbox{\emph{Relative date}}_t$ is the number of days from the date of the switch (zero is the first period after the switch); $\mbox{\emph{After switch}}_t$ is a binary variable equal to one if $t$ is a period after the switch and zero otherwise; $\mbox{\emph{Relative date}}_t \times \mbox{\emph{After switch}}_t$ is called the interaction term and $\epsilon_t$ is the residual error.
+This is equivalent to estimating the following two regressions, respectively before and after the switch:
+\begin{equation*}
+\begin{array}{l}
+\mbox{\emph{Number of bug reports}}_{t < 0} ⁡=\\
+\quad\quad \gamma_0 + \gamma_1 \times \mbox{\emph{Relative date}}_t + \epsilon_p \medskip\\
+\mbox{\emph{Number of bug reports}}_{t \geq 0} = \\
+\quad\quad (\gamma_0 + \gamma_2) + (\gamma_1 + \gamma_3) \times \mbox{\emph{Relative date}}_t + \epsilon_t
+\end{array}
+\end{equation*}
+We estimate this regression by Ordinary Least Squares~\cite{wooldridge2015introductory} and compute heteroscedasticity robust standard errors.\footnote{The errors $\epsilon_t$ are said to be heteroscedastic if their variances differ.
+The standard calculations for estimation errors assume equal variances~\cite{wooldridge2015introductory}.
+} Coefficient $\gamma_0$ is the estimated value just before the cutoff and $\gamma_1$ the slope before the cutoff.
+Coefficients $\gamma_2$ and $\gamma_3$ are the estimates of interest and will tell us the jump of the number of bugs reports just after the switch and the change in slopes due to the switch respectively.
 
-\noindent where $\mbox{\emph{Number of bug reports}}_p⁡$ is the total number of bugs reported during the 15-day period $p$; $\mbox{\emph{Relative date}}_p$ is the number of periods from the date of the switch (zero is the first period after the switch); $\mbox{\emph{After switch}}_p$ is a binary variable equal to one if $p$ is a period after the switch and zero otherwise; $\mbox{\emph{Relative date}}_p * \mbox{\emph{After switch}}_p$ is the interaction term and $\epsilon_p$ is the error.
-We estimate this regression by Ordinary Least Squares.
-Coefficients $\gamma_2$ and $\gamma_3$ are the estimates of interest and will tell us the jump of the number of bugs reports just after the migration and the change in trends due to the migration respectively.
 We estimate this regression for two different sub-samples: the developers and the non-developers.
-
-We replicate this analysis for all our outcome variables by changing the variable on the left hand of the equation.
+We replicate this analysis for all our outcome variables by changing the variable on the left hand side of the equation.
 
 \section{Results}
-
-\subsection{Impact on the number of bugs}
+\label{results}
+\subsection{Impact on the number of bug reports}
 
@@
 \caption{Estimated impact of the switch on the number of bugs.
-Coefficients are highlighted if they are statistically significant with $p<0.1$ (*), $p<0.05$ (**) or $p<0.01$ (***).
-Standard error is in parentheses.}\label{tab:bug_nb}
-\begin{tabular}{|r|c|c|c|}
-\hline
-&  Total & Developers & Others \\
-\hline
-$\mbox{\emph{After switch}}_p$ & 10.1* & 8.79** & 1.27 \\
-& (5.06) & (3.55) & (4.16) \\
-\hline
-$\mbox{\emph{Relative date}}_p * \mbox{\emph{After switch}}_p$ & 0.717 & -0.335 & 1.05*** \\
-& (0.438) & (0.307) & (0.36) \\
-\hline
-$\mbox{\emph{Relative date}}_p$ & 0.219 & 0.25 & -0.0308 \\
-& (0.309) & (0.217) & (0.255) \\
-\hline
-Constant & 17.6*** & 5.72** & 11.9*** \\
-& (3.71) & (2.6) & (3.05) \\
-\hline
-Observation number & 40 & 40 & 40 \\
-\hline
-\end{tabular}
+Coefficients are highlighted if they are statistically significant with $p<0.1$ ($\dagger$), $p<0.05$ (*), $p<0.01$ (**) or $p<0.001$ (***).
+Standard error is in parentheses.}
+\label{tab:bug_nb}
+\input{bug_nb_rd.tex}
 \end{table}
 
-Table~\ref{tab:bug_nb} presents the estimated impact of the switch on the number of bugs.
-Each column shows the estimates for a different sub-sample (i.e.
-all bug reporters, developers and non-developers).
-Asterisks represent the level of significance.
-Fig.~\ref{bug_nb_rd}  shows the number of reported bugs before and after the switch and the fitting lines corresponding to the regression results.
-These results give strong evidence that the switch positively affected  the number of bug reports.
-We can observe different effects for developers and non-developers.
+\begin{figure}
+\includegraphics{bug_nb_rd.png}
+\caption{Average number of bug reports per day before and after the migration (with fitting lines and confidence intervals from the regression results).} \label{bug_nb_rd}
+\end{figure}
 
-For developers we see a statistically significant positive effect on the jump of bug reports just after the switch and then a stabilization of the reported bugs at a higher level than before.
-In particular changing the bug reporting platform is estimated to increase the number of reported bugs by 8 bugs.
-That is, in average, we observe 8 more bugs during 15-day periods after the switch (but close to the switch) compared to 15-day periods before the switch (but close to the switch).
-This means that the number of bug reports by developers after the switch is more than twice the number before the switch.
+Table~\ref{tab:bug_nb} presents the estimated impact of the switch on the number of reported bugs.
+Each column shows the estimates for a different sub-sample (i.e., all bug reporters, developers and non-developers).
+Asterisks represent different classical levels of statistic significance.
+We interpret results as significant if the p-value is below 0.1.
+Estimates that are not statistically significant cannot be interpreted as an absence of effect but indicate that if such effect occurred, we are unable to discern it with our conservative approach.
+Fig.~\ref{bug_nb_rd}  shows the number of reported bugs before and after the switch and the fitting lines and confidence intervals corresponding to the regression results.
 
-For non-developers, we see a statistically significant increase in the slope but not on the jump in magnitude just after the switch.
-In particular, switching the bug reporting platform to GitHub increases the slope of the regression line by 1.
-This means that after the switch we will, in average, observe an increase of one bug for each 15-day period.
-
-This behavioral difference is not surprising.
-Developers were well-aware that the migration had happened and this caused a spike in activity.
-Traditionally, they had used the bug tracker very little (fixing bugs when they saw them or using personal to-do lists) and were using it more all a sudden (as a shared to-do list and a way to increase communication ---note that one of the \emph{explicit} goals of the switch was to trigger this culture shift).
-On the contrary, the migration was not responsible for any significant leap in the rate of bug reporting by non-developers, but this rate then increased more and more.
-This could be due in particular to users noticing that their bug reports were treated more promptly, which could be an encouragement for reporting more of them.
-Another possible influencing factor may be the fact that non-developers may take more time to adapt to this new platform.
-Indeed, developers were the ones that took the decision to switch the bug tracker, among other reasons because they were already used to GitHub.
-
-\begin{figure}
-\includegraphics[width=\textwidth]{bug_nb_rd.eps}
-\caption{Number of bug reports by 15-day period before and after the migration (with fitting lines from the regression results).} \label{bug_nb_rd}
-\end{figure}
+For developers, we see a statistically significant positive jump in bug reports just after the switch.
+In particular changing the bug reporting platform is estimated to have increased the daily number of reported bugs by 0.7 (representing an increase of around 250\%).
+That is, on average, we observe around one bug reported by developers every day after the switch, %(but close to the switch)
+while before the switch developers reported a bug every three days.
+There is no statistically significant effect on the number of bug reports by other contributors: we cannot reject that such a variation occurred, but we are unable to discern it with our conservative approach.
 
@@
 
-Table~\ref{tab:reporters} gives the estimated impact of the switch on the number of bug reporters (i.e.
-number of people that report a bug during a given period).
-As before, each column shows the estimates for a different sub-sample.
-Fig.~\ref{reporter_nb_rd}  shows the number of bug reporters before and after the switch and the fitting lines correspond to the regression results.
-These results demonstrate that the switch to GitHub positively affected not just the number of bug reports but also the number of non-developer bug reporters, suggesting an impact in both the intensive (number of bugs reported per person) and the extensive margin (number of reporters) of bug reporting.
+Table~\ref{tab:reporters} and Fig.~\ref{reporter_nb_rd} show the estimated impact of the switch on the number of distinct bug reporters each week.
+These results demonstrate that the switch to GitHub positively affected not just the number of bug reports but also the number of distinct developer bug reporters in a given week.
+The number of distinct developers that reported bugs in a given week increased by around 130\% (from a 1.5 reporting developers each week before the switch to 3.42 after)
 
@@
 \centering
-\caption{Estimated impact of the switch on the number of reporters by 15-day period.
-Coefficients are highlighted if they are statistically significant with $p<0.1$ (*), $p<0.05$ (**) or $p<0.01$ (***).
-Standard error is in parentheses.}\label{tab:reporters}
-\begin{tabular}{|r|c|c|c|}
-\hline
-&  Total & Developers & Others \\
-\hline
-$\mbox{\emph{After switch}}_p$ & 4.15 & 1.24 & 2.91 \\
-& (2.52) & (0.974) & (2.46) \\
-\hline
-$\mbox{\emph{Relative date}}_p * \mbox{\emph{After switch}}_p$ & 0.289 & -0.156* & 0.444** \\
-& (0.218) & (0.0843) & (0.213) \\
-\hline
-$\mbox{\emph{Relative date}}_p$ & 0.111 & 0.159** & -0.0489 \\
-& (0.154) & (0.0596) & (0.151) \\
-\hline
-Constant & 12*** & 3.97*** & 8.04*** \\
-& (1.84) & (0.714) & (1.81) \\
-\hline
-Observation number & 40 & 40 & 40 \\
-\hline
-\end{tabular}
+\caption{Estimated impact of the switch on the number of distinct reporters each week.
+Coefficients are highlighted if they are statistically significant with $p<0.1$ ($\dagger$), $p<0.05$ (*), $p<0.01$ (**) or $p<0.001$ (***).
+Standard error is in parentheses.}
+\label{tab:reporters}
+\input{reporter_nb_rd.tex}
 \end{table}
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{reporter_nb_rd.eps}
-\caption{Number of distinct reporters by 15-day period before and after the migration (with fitting lines from the regression results).} \label{reporter_nb_rd}
+\includegraphics{reporter_nb_rd.png}
+\caption{Average number of distinct reporters each week before and after the migration (with fitting lines and confidence intervals from the regression results).} \label{reporter_nb_rd}
 \end{figure}
-
-For non-developers, we see a statistically significant increase in the slope on the trend after the switch.
-In particular, switching the bug reporting platform to GitHub increases the slope of the regression line by 0.44.
-This means that after the switch we will, in average, observe an increase of around one non-developer bug reporter each new month.
-So for each period, there will be more non-developers who report a bug.
-This does not necessarily mean that these are people that never reported bugs before.
-It could also be that non-developer reporters start reporting more and more frequently.
-Again, this could be due to the increase of the bug reporting appeal, partly because of the platform, but also partly because they receive more feedback on their bug reports (as will suggest our results on comment numbers).
-
-For developers we see a slight but statistically significant (at 10\% level) negative effect on the slope of the trend after the switch.
-This means that, compared to the slope of the trend before the switch, the new trend has a lower slope.
-We wouldn't expect a significant change in the number of developers that report bugs because according to our definition there is a fixed number of developers, and they are already very active.
-Seeing the figure, we see  that the change in slope is due to a stabilization of a previous positive slope, probably meaning a reach of the maximum level of active developers per period.
 
@@
 
-Table~\ref{tab:comment_nb} reports on the estimated impact of the bug tracker switch on the number of comments.
-We can see a statistically significant jump in the number of comments, which almost doubles.
-This is almost entirely due to comments by developers.
-This surge in bug comment activity shows that developers communicate more on the bug reports they receive, instead of just fixing and closing them.
-This means that the switch was beneficial in terms of transparency of the development, an evolution that is likely to give greater satisfaction to the bug reporters.
+Table~\ref{tab:comment_nb} and Fig.~\ref{comment_nb_rd} show the estimated impact of the bug tracker switch on the number of comments.
+We see a statistically significant jump in the number of comments, which more than doubles (from 4.8 comments per day before the switch to 10.3 after the switch).
+This appears to be, for the most part, due to comments by developers who increased their average number of comments per day from around 3 before the switch to around 8 after.
 
@@
 \caption{Estimated impact of the switch on the number of comments.
-Coefficients are highlighted if they are statistically significant with $p<0.1$ (*), $p<0.05$ (**) or $p<0.01$ (***).
-Standard error is in parentheses.}\label{tab:comment_nb}
-\begin{tabular}{|r|c|c|c|}
-\hline
-&  Total & Developers & Others \\
-\hline
-$\mbox{\emph{After switch}}_p$ & 84.7** & 72.2** & 12.5 \\
-& (35.5) & (27.9) & (11.1) \\
-\hline
-$\mbox{\emph{Relative date}}_p * \mbox{\emph{After switch}}_p$ & -2.1 & -2.79 & 0.698 \\
-& (3.07) & (2.41) & (0.957) \\
-\hline
-$\mbox{\emph{Relative date}}_p$ & 1.95 & 1.65 & 0.3 \\
-& (2.17) & (1.71) & (0.677) \\
-\hline
-Constant & 86.6*** & 59.7*** & 26.9*** \\
-& (26) & (20.4) & (8.11) \\
-\hline
-Observation number & 40 & 40 & 40 \\
-\hline
-\end{tabular}
+Coefficients are highlighted if they are statistically significant with $p<0.1$ ($\dagger$), $p<0.05$ (*), $p<0.01$ (**) or $p<0.001$ (***).
+Standard error is in parentheses.}
+\label{tab:comment_nb}
+\input{comment_nb_rd.tex}
 \end{table}
+
+\begin{figure}
+\includegraphics{comment_nb_rd.png}
+\caption{Average number of bug comments per day before and after the migration (with fitting lines and confidence intervals from the regression results).} \label{comment_nb_rd}
+\end{figure}
 
@@
 
-Table~\ref{tab:commentators} and Fig.~\ref{commentator_nb_rd} show the estimated impact of the bug tracker switch on the number of distinct commentators by 15-day period.
-We can see a highly statistically significant jump in the number of non-developer commentators.
-This result is particularly interesting because it highlights a remarkable change in the way non-developers engage in the software developing process following the switch.
-What aspect of the GitHub platform creates this greater engagement remains to be studied but we hypothesize that this is due in large part to GitHub making it easier to subscribe to the bug report activity, and possibly also to the enhanced search.
+Table~\ref{tab:commentators} and Fig.~\ref{commentator_nb_rd} show the estimated impact of the bug tracker switch on the number of distinct commentators each week.
+We can see a statistically significant jump in the total number of commentators, in the number of developer commentators, and in the number of non-developer commentators.
+
+In general terms, the number of distinct commentators in a given week changed from an average of 7 to an average of 13 (almost doubling the commentators).
+Analyzing the developers and the non-developers, we can see that this increase is due in larger part to non-developers (whose number of commentators more than doubles) while, among developers, this number increases by 66\%.
 
@@
 \centering
-\caption{Estimated impact of the switch on the number of distinct commentators by 15-day period.
-Coefficients are highlighted if they are statistically significant with $p<0.1$ (*), $p<0.05$ (**) or $p<0.01$ (***).
-Standard error is in parentheses.}\label{tab:commentators}
-\begin{tabular}{|r|c|c|c|}
-\hline
-&  Total & Developers & Others \\
-\hline
-$\mbox{\emph{After switch}}_p$ & 7.48*** & 0.763 & 6.72*** \\
-& (2.63) & (0.684) & (2.3) \\
-\hline
-$\mbox{\emph{Relative date}}_p * \mbox{\emph{After switch}}_p$ & 0.238 & -0.0684 & 0.307 \\
-& (0.228) & (0.0592) & (0.199) \\
-\hline
-$\mbox{\emph{Relative date}}_p$ & -0.00977 & 0.0368 & -0.0466 \\
-& (0.161) & (0.0419) & (0.141) \\
-\hline
-Constant & 13.4*** & 7.04*** & 6.41*** \\
-& (1.93) & (0.502) & (1.69) \\
-\hline
-Observation number & 40 & 40 & 40 \\
-\hline
-\end{tabular}
+\caption{Estimated impact of the switch on the number of distinct commentators each week.
+Coefficients are highlighted if they are statistically significant with $p<0.1$ ($\dagger$), $p<0.05$ (*), $p<0.01$ (**) or $p<0.001$ (***).
+Standard error is in parentheses.}
+\label{tab:commentators}
+\input{commentator_nb_rd.tex}
 \end{table}
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{commentator_nb_rd.eps}
-\caption{Number of distinct commentators by 15-day period before and after the migration (with fitted lines from the regression results).} \label{commentator_nb_rd}
+\includegraphics{commentator_nb_rd.png}
+\caption{Average number of distinct commentators each week before and after the switch (with fitting lines and confidence intervals from the regression results).} \label{commentator_nb_rd}
 \end{figure}
 
+\section{Discussion}
+\label{discussion}
+
+\subsection{H1: impact on the number of reported bugs}
+
+We hypothesized that the switch to a platform that is perceived as more modern and easier to use would increase the number of reported bugs.
+Our data analysis only shows such an effect on the number of bug reported by developers, and not on the bugs reported by non-developers, which only partially validates our first hypothesis.
+
+The developers were using the bug tracker very little before the switch.
+This means that when they found bugs in the course of the development, they stored them in personal task lists, fixed them immediately, or tried to remember them in their heads until they could fix them (likely with some bugs being forgotten in the process).
+Meanwhile, they were discussing a lot through the pull request system on GitHub.
+The move of the bug tracker to GitHub made it closer to the main development platform (the repository where pull requests are submitted and reviewed), both in terms of location and in terms of user interface.
+These reasons are probably what pushed developers to use it more.
+
+The increased use of the bug tracker in place of personal task lists makes the whole development process more transparent and is likely to be beneficial to new contributors.
+
+If it is the case that the switch had no effect on the number of bugs reported by non-developers, this could mean that when people decide to report bugs, they do so before accessing the bug tracker, and nor the requirement of creating an account, nor the heavier and less modern interface, are sufficient to discourage them.
+
+%The second one is that more people intended to report bugs but this increase was blurred out by a simultaneous decrease in the number of duplicate bug reports, thanks to the better search provided by GitHub.
+People intending to report a bug and finding out that it was already reported would post a comment (although if this was frequently the case, there should be an increase in the number of comments by non-developers, which hasn't been observed in a significant way neither) or add a reaction (GitHub feature allowing to up-vote a comment) or leave without taking any action instead.
+To confirm this second interpretation, we would need to analyze the impact of the switch on the number of bugs closed as duplicates.
+
+\subsection{H2: impact on the pool of reporters and commentators}
+
+% how many of the reporters of recent bugs had a GitHub account?
+
+Having a common platform for the pull requests and the bug reports, and using a standard solution where a lot of users would already have an account, was hypothesized to have an impact on the pool of participants in the bug tracker.
+For non-developers, our data analysis only shows this effect in a statistically significant way on the number of commentators, which only partially validates our second hypothesis.
+
+Analyzing deeper, we find that there exists not just a jump in the non-developer commentators in a given week but also a jump in the number of new non-developer commentators (i.e., people that had never before posted a comment).
+We show the results for the new commentators in the companion Jupyter notebook.
+
+While we suggested that reporters would be motivated enough to not be affected by the bug tracking platform, this is less likely to be the case for people willing to comment.
+Furthermore, people who were already following the pull request activity were now subscribed to the bug tracking activity as well and this could have been enough to encourage them to post comments.
+The previous system made it harder and less flexible to subscribe to the bug tracker activity.
+
+This analysis is consistent with previous findings: many people start by \emph{lurking} (i.e., mostly reading discussions) before starting to contribute to an open source project (for instance by participating to discussions)~\cite{von2003community}.
+In the context of GitHub, a significant portion of the new contributors start by \emph{watching} the repository (i.e., subscribing to its activity).
+These contributors contribute for longer periods and achieve a wider variety of contribution types, probably due to the knowledge and confidence they gained during the watching process~\cite{sheoran2014understanding}.
+
+% Commenté pour faire plus court, et ça n'est pas suffisamment essentiel ici.
+%These results are particularly interesting because they highlight an increase in the number of people involved in the development process in the large sense.
+They also show a remarkable change in the way non-developers engage in the software developing process following the switch.
+
+The significant impact on distinct reporters and commentators among developers is more surprising but shows that some developers were disengaged from any kind of activity on the bug tracker before the switch.
+One developer explained this as a difficulty to keep track of activity on two platforms at once (GitHub for the pull requests and Bugzilla for the bug reports).
+Comparing data from pull request commentators and bug report commentators confirmed this: before the switch there were more developers commenting pull requests than bug reports in any given week, whereas after the switch this difference disappeared completely (graph available in the accompanying notebook).
+
+%todo for future... are there more non-developers closer to become developers ?
+
+\subsection{H3: impact on the interaction between users and developers}
+
+The surge in the number of comments posted by developers shows that developers communicate more on the bug reports they receive, instead of just fixing and closing them.
+This result suggests once again that the switch was beneficial in terms of transparency of the development, an evolution that is likely to give greater satisfaction to the bug reporters and probably encourage engagement in the process of bug reporting and development more generally.
+Taken together with the increased number of non-developer commentators, this validates our third hypothesis.
+
+% Commenté pour faire plus court, et ça n'est pas suffisamment essentiel ici.
+Overall, these results suggest that the change of platform reinforced discussion among developers and among developers and non-developers, which may improve efficiency in the developing process, incentivize non-developers to have a more active role (maybe even assuming developing tasks) and in the long run increase the quality of the software (as bugs are probably better understood and fixed in better accordance to users' needs).
+
 \section{Threats to validity}
-
-\subsection{Internal threats}
+\label{threats}
+\subsection{Internal validity and robustness checks}
 
@@
 Then, the hypothesis that there is no other discontinuity around the threshold of interest does not completely hold in our case.
-First, we can argue that not all our results are threatened: the jump in developer reports isn't because there is usually no peak in developer reports after new releases, same for the jump in non-developer commentators.
-Second, we can also argue that there should have been a similar (or higher) peak in the month before the switch since the first beta of 8.7 had been released then (and beta releases generally result in higher peaks).
-Third, we test for robustness by removing the two 15-day periods following the 8.7.0 release (and the switch) from our Regressions in Discontinuity.
-With these two periods removed, the regressions still give very close results with the same statistical significance.
+%First, we can argue that only our result on the number of comments by developers is threatened: the other statistically significant effects that we observed are not threatened because there is usually no peak in the corresponding variables after new releases.
+%Second, we can also argue that there should have been a similar (or higher) peak in the month before the switch since the first beta of 8.7 had been released then (and beta releases generally result in higher peaks).
+To confirm that this is not what is driving our results, we conduct two robustness checks with our data.
+The first consists in removing the two weeks following the 8.7.0 release (and the switch) from our Regressions in Discontinuity.
+With these periods removed, the regressions still give very close results.
+The second consists in including a binary control variable that is equal to one if the period is in the two weeks after a release and zero otherwise.
+This control variable will capture the high peaks due to the release without the need of eliminating periods close to the switch.
+Again, our results remain very similar when we include this control variable.
 
-We also conduct one standard robustness check for Regression in Discontinuity analysis by changing the time window of the analysis to fifteen and to ten periods before and after the switch.
+Our main specification for Regression on Discontinuity uses a small bandwidth and a linear model.
+In this section, we also present the results using all the available data (455 days on each side of the cutoff, corresponding to all the data that we collected after the switch and the same number of days before).
+To reduce the associated bias, we include a quadratic term to allow the regression to better fit the data and "adjust" to far away data points.
+Following Gelman et al.
+~\cite{gelman2018high}, we include only a quadratic polynomial and not a polynomial of a higher order.
+Our results still hold with this specification.
+
+%In order to corroborate that our results are not driven by activity occurring much after or much before the switching period (that is more likely to be affected by other factor beyond the switch), we check if our results are robust to the exclusion of the 1\%, 5\% and 10\% external values (i.e., values that are furthest from the switch) ~\cite{jacob2012practical}.
+
+%In the context of a Regression in Discontinuity analysis, the choice of the bandwidth (number of periods around the switch) of the analysis and the functional form of the regression is always difficult.
+By the one hand, including data points that are far from the cut-off (in our case the switch) will increase precision as it will reduce standard errors (because we will have more data points).
+By the other hand, however, including those points can bias estimation near the cut-off.
+In our preferred specification that we present in the results section, we take a conservative approach and take a relatively small bandwidth to minimize possible bias.
+In this section, we also present the results using all the disposable data.
+To reduce the associated bias, we include a quadratic term to allow the regression to better fit the data and "adjust" for far away data points.
+Following Gelman et al.
+(2018) ~\cite{gelman2018high}, we include only a quadratic polynomial and not a polynomial of a higher order.
+
+%We also conduct one standard robustness check for Regression in Discontinuity analysis by changing the time window of the analysis to fifteen and to ten periods before and after the switch.
 Our main results are robust to these changes.
 
-\subsection{External threats}
+\subsection{External validity}
 
-This is only a case study on a very particular project (it is a software that can be used as a programming language, but requires expertise, that is developed mostly by French researchers, etc.).
-The only way to confirm the results we obtained would be to replicate this study on other examples of projects that went through a similar bug tracker switch.
+This is only a case study on a very specific project (by the target audience, the location and composition of its core team, etc.).
+We would really benefit from conducting similar analyses on other projects that went through a similar bug tracker switch.
+
+We located another project that went through a similar switch from Bugzilla to GitHub and used the Bugzilla import tool with our modifications.
+It would have been a perfect case for replicating this analysis.
+A preliminary visualization of the data shows that, compared to the few years before the switch, there have been more bug reports and comments posted after the switch.
+Unfortunately, we learned that some data was lost during the crash of their previous bug tracker, which makes it really difficult to draw any reliable conclusion.
+%after talking with the developer responsible for the migration, we learned that the switch was the consequence of a crash of the Bugzilla instance that could not be fixed, and that some data (bugs and comments) were lost in the crash.
+Some of it could be recovered and was manually added, which makes it inaccurately appear as activity after the switch.
+
+We nonetheless believe that our conclusions can be generalized to some extent and encourage researchers to replicate this study, for instance using our migration tool and/or our analysis pipeline and methodology.
+
+There is no reason to believe that these results would generalize to proprietary software projects.
+Some projects that are also less likely to benefit as much from a switch to GitHub are the larger projects (like Mozilla, Linux, Gnome, KDE, Apache) which have been using Bugzilla for a long time and have defined rigorous development processes, in particular regarding the use of the bug tracker.
+Coq is an example of a project that is developed in a very Bazaar-like fashion, with no clear rules on how to use the bug tracker, and this lack of rules was hindering the use of a complex bug tracker like Bugzilla.
+We believe that open source projects that share this lack of rules and this medium size are the more likely to benefit from a switch from Bugzilla (or similar bug trackers like Mantis), to GitHub (or the very similar GitLab platform), especially if they are already heavily using the latter for pull requests.
 
 \section{Conclusion and perspectives}
-
+\label{conclusion}
 We have analyzed the impact of changing the bug reporting environment using a Regression in Discontinuity analysis with data from the switch of the Coq bug tracker from Bugzilla to GitHub.
-We have shown that the switch increased the number of bug reports; started an increase of bug reporters per period (i.e.
-in a certain period there are more distinct reporters); increased the number of comments by developers and the number of distinct commentators per period among non-developers.
-These results fully validate our first and third hypotheses (increase of the number of bug reports, by developers in particular; improvement of the interaction between reporters and developers) and partially validate our second hypothesis (increase the diversity of bug reporters).
-We find particularly interesting the change in the number of comments by developers and in the number of distinct non-developer commentators because it shows a significant change in the dynamics of the bug reporting process that we believe can improve software development by engaging more people in the process, and by increasing openness and transparency.
+We have shown that the switch incurred an increase in bug tracking activity in two ways.
+First, the developers are using the bug tracker more, to report their own bugs, and to discuss bug reports.
+Second, the non-developers are more numerous to take an active role by commenting bug reports as well.
+We find particularly interesting the change in the number of comments by developers and in the number of distinct non-developer commentators because it shows a shift in the dynamics of the bug reporting process that we believe can improve software development by engaging more people in the process, and by increasing openness and transparency.
 
@@
 Analyzing other outcomes (such as pull requests or the rapidity of bug resolution) could give interesting additional insights and help create a more complete picture of the mechanisms at play.
-It will be also interesting to analyze the effects on a longer period to see if the changes hold in the long run.
+% je pense qu'on peut se permettre d'enlever cette phrase maintenant
+%It will be also interesting to analyze the effects on a longer period to see if the changes hold in the long run.
 Replicating our analysis for different contexts and software would increase the external validity of the results.
 
+Finally, we introduced and showed an example of the use of a Regression on Discontinuity design, that is novel in the context of empirical software engineering.
+We think that there are many cases where this methodology would be useful to help demonstrate causality from mined software repository data.
+
+% When to switch?
+It would be useful to come up with measures that indicate when a bug tracker is not adapted and a switch to another platform would be an improvement.
+(Or depending on specific objectives.)
+
 \scriptsize
-\paragraph{}
+%\paragraph{}
 \noindent \footnotesize{\textbf{Supporting data.}}
@@
 We also provide the XML export of Bugzilla that was used for the migration and the migration script.
-Everything is available from \url{https://github.com/Zimmi48/impact-of-switching-bug-trackers}.
-\paragraph{}
-\noindent \footnotesize{\textbf{Authors' contributions.}}
-TZ prepared and conducted the migration, designed and implemented the data extraction process, implemented the data analyses, and wrote part of the manuscript.
-AC designed the data analyses and wrote part of the manuscript.
-\paragraph{}
-\noindent\footnotesize{\textbf{Acknowledgements.}}
-We would like to thank Ambre Williams for valuable advice and help on the design and implementation of the data analyses and C\'ecile Rottner for general advice, support, and proof-reading.
+Everything is available from \url{https://github.com/Zimmi48/github-migration-analysis}.
+%\paragraph{}
+
+%\noindent \footnotesize{\textbf{Authors' contributions.}}
+%[masked for double-blind review]
+%TZ found the research question, prepared and conducted the migration, designed and implemented the data extraction process, implemented the data analyses, and wrote part of the manuscript.
+AC designed the data processing and analyses, guided the implementation of the data analyses, presented the econometric method in use and wrote part of the manuscript.
+%\paragraph{}
+
+%\noindent\footnotesize{\textbf{Acknowledgements.}}
+%We would like to thank Ambre Williams for valuable advice and help on the design and implementation of the data analyses and C\'ecile Rottner for general advice, support, and proof-reading.
+%We would like to thank Andriy Berestovskyy for providing the Bugzilla import tool, and Martin Michlmayr for doing the first steps in porting our modifications upstream.
 

### 2019-03-05 article/article.tex
Fix compilation.

@@
 
-The highest peak is in 2015, after the first beta release of version 8.5.
-This was the first release in more than two years and it contained five big new features (``the result of five specific long-term projects''\footnote{Quote from \url{https://coq.inria.fr/refman/credits.html##credits-version-8-5}.
-The next quotes are coming from the following sections of the same chapter of the Coq reference manual.}).
-Stabilizing these features and their (unplanned) interactions required more than a year of testing and three beta releases.
-The 8.5 release was both hard on developers, who found the release cycle too long, and on users, who were impacted by many compatibility issues.
-Consequently, a different approach was taken for the following releases: versions 8.6, 8.7 and 8.8 were ``developed on a time-based development cycle'' and contain ``the result of refinements, stabilization of features and cleanups of the internals of the system along with a few new features''.
-Attention was given to regression testing: changes were systematically tested by building a selection of the largest Coq projects.
-Therefore, it is not surprising that these frequent releases are correlated to smaller peaks of new bug reports.
+The highest peak is in 2015, after the first beta release of version
+8.5. This was the first release in more than two years and it
+contained five big new features (``the result of five specific
+long-term projects''\footnote{Quote from
+  \url{https://coq.inria.fr/distrib/V8.9.0/refman/credits.html\#credits-version-8-5}. The
+  next quotes are coming from the following sections of the same
+  chapter of the Coq reference manual.}). Stabilizing these features
+and their (unplanned) interactions required more than a year of
+testing and three beta releases. The 8.5 release was both hard on
+developers, who found the release cycle too long, and on users, who
+were impacted by many compatibility issues. Consequently, a different
+approach was taken for the following releases: versions 8.6, 8.7 and
+8.8 were ``developed on a time-based development cycle'' and contain
+``the result of refinements, stabilization of features and cleanups of
+the internals of the system along with a few new features''. Attention
+was given to regression testing: changes were systematically tested by
+building a selection of the largest Coq projects. Therefore, it is not
+surprising that these frequent releases are correlated to smaller
+peaks of new bug reports.
 
@@
 
-%\noindent\footnotesize{\textbf{Acknowledgements.}}
-%We would like to thank Ambre Williams for valuable advice and help on the design and implementation of the data analyses and C\'ecile Rottner for general advice, support, and proof-reading.
-%We would like to thank Andriy Berestovskyy for providing the Bugzilla import tool, and Martin Michlmayr for doing the first steps in porting our modifications upstream.
+\noindent\footnotesize{\textbf{Acknowledgements.}}
+We would like to thank Ambre Williams for valuable advice and help on the design and implementation of the data analyses and C\'ecile Rottner for general advice, support, and proof-reading.
+We would like to thank Andriy Berestovskyy for providing the Bugzilla import tool, and Martin Michlmayr for doing the first steps in porting our modifications upstream.
 
@@
 % ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
-\bibliographystyle{splncs04}
 \bibliography{bib}
+\bibliographystyle{IEEEtran}
 %

### 2019-03-05 article/article.tex
Fix URL to GitHub repository.

@@
 Using this API allows to do large requests (100 nodes in a single request) with just the information we need (thus both reducing the bandwidth usage and speeding up the extraction process).
-We provide the extracted data as CSV files and a Jupyter notebook\footnote{Available at \url{https://github.com/Zimmi48/github-migration-analysis}.} with the code to request this data from GitHub, to load the CSV files, to run the pre-processing steps, and the analyses.%, and the robustness checks.
+We provide the extracted data as CSV files and a Jupyter notebook\footnote{\url{https://github.com/Zimmi48/impact-of-switching-bug-trackers}.} with the code to request this data from GitHub, to load the CSV files, to run the pre-processing steps, and the analyses.%, and the robustness checks.
 
@@
 
-The highest peak is in 2015, after the first beta release of version
-8.5. This was the first release in more than two years and it
-contained five big new features (``the result of five specific
-long-term projects''\footnote{Quote from
-  \url{https://coq.inria.fr/distrib/V8.9.0/refman/credits.html\#credits-version-8-5}. The
-  next quotes are coming from the following sections of the same
-  chapter of the Coq reference manual.}). Stabilizing these features
-and their (unplanned) interactions required more than a year of
-testing and three beta releases. The 8.5 release was both hard on
-developers, who found the release cycle too long, and on users, who
-were impacted by many compatibility issues. Consequently, a different
-approach was taken for the following releases: versions 8.6, 8.7 and
-8.8 were ``developed on a time-based development cycle'' and contain
-``the result of refinements, stabilization of features and cleanups of
-the internals of the system along with a few new features''. Attention
-was given to regression testing: changes were systematically tested by
-building a selection of the largest Coq projects. Therefore, it is not
-surprising that these frequent releases are correlated to smaller
-peaks of new bug reports.
+The highest peak is in 2015, after the first beta release of version 8.5.
+This was the first release in more than two years and it contained five big new features (``the result of five specific long-term projects''\footnote{Quote from:\\ \url{https://coq.inria.fr/distrib/V8.9.0/refman/credits.html\#credits-version-8-5}\\
+The next quotes are coming from the following sections of the same chapter of the Coq reference manual.}).
+Stabilizing these features and their (unplanned) interactions required more than a year of testing and three beta releases.
+The 8.5 release was both hard on developers, who found the release cycle too long, and on users, who were impacted by many compatibility issues.
+Consequently, a different approach was taken for the following releases: versions 8.6, 8.7 and 8.8 were ``developed on a time-based development cycle'' and contain ``the result of refinements, stabilization of features and cleanups of the internals of the system along with a few new features''.
+Attention was given to regression testing: changes were systematically tested by building a selection of the largest Coq projects.
+Therefore, it is not surprising that these frequent releases are correlated to smaller peaks of new bug reports.
 
@@
 We also provide the XML export of Bugzilla that was used for the migration and the migration script.
-Everything is available from \url{https://github.com/Zimmi48/github-migration-analysis}.
+Everything is available from \url{https://github.com/Zimmi48/impact-of-switching-bug-trackers}.
 %\paragraph{}

### 2019-03-31 article/article.tex
Functional notebook with OCaml for extended version.

@@
 The impact of the bug reporting environment on the bug tracking activity is difficult to assess because of the lack of comparison points.
-In this paper, we take advantage of the switch, from Bugzilla to GitHub, of the bug tracker of Coq, a medium-sized open source project, to evaluate the impact that such a change can have.
+In this paper, we take advantage of the switch, from Bugzilla to GitHub, of the bug tracker of Coq, a medium-sized open source project, to evaluate and interpret the impact that such a change can have.
 
-We first report on the switch itself, in particular the migration of 4900 preexisting bug reports using the GitHub issue import API.
+We first report on the switch itself, including the migration of preexisting bug reports.
 Then we analyze data from before and after the switch using a Regression on Discontinuity analysis, a novel methodology in the context of empirical software engineering.
+We complete this quantitative analysis with qualitative data from interviews with developers.
 We show that the switch induces an increase in bug reporting, particularly from principal developers themselves, and more generally an increased engagement with the bug tracking platform, with more comments by developers and also more external commentators.
@@
 \begin{IEEEkeywords}
-bug tracker, switch, migration, bug report, issue, GitHub, Bugzilla, open source, data mining, Regression on Discontinuity
+bug tracker, switch, migration, bug report, issue, GitHub, Bugzilla, open source, data mining, Regression on Discontinuity, interviews
 \end{IEEEkeywords}
@@
 
-\subsection{Cumulative number of bug reports and bug comments}
+%\subsection{Cumulative number of bug reports and bug comments}
 
-Fig.~\ref{cumulative_bugs} and~\ref{cumulative_comments} give a first (cumulative) view of two of our outcome variables: the number of bug reports, and the number of comments.
+%Fig.~\ref{cumulative_bugs} and~\ref{cumulative_comments} give a first (cumulative) view of two of our outcome variables: the number of bug reports, and the number of comments.
 From these figures, we can clearly notice the dominance of developers in the number of bug comments and the dominance of non-developers in the number of bug reports.
@@
 
-\begin{figure}
-\includegraphics{cumulative_bugs.eps}
-\caption{Cumulative number of bug reports (since 2008).} \label{cumulative_bugs}
-\end{figure}
+%\begin{figure}
+%\includegraphics{cumulative_bugs.eps}
+%\caption{Cumulative number of bug reports (since 2008).} \label{cumulative_bugs}
+%\end{figure}
 
-\begin{figure}
-\includegraphics{cumulative_comments.eps}
-\caption{Cumulative number of comments (since 2008).} \label{cumulative_comments}
-\end{figure}
+%\begin{figure}
+%\includegraphics{cumulative_comments.eps}
+%\caption{Cumulative number of comments (since 2008).} \label{cumulative_comments}
+%\end{figure}
 
@@
 \caption{Number of bug comments by 4-week period with release dates (since 2008).} \label{comments_with_releases}
+\end{figure}
+
+\begin{figure}
+\includegraphics{reporters_with_releases.eps}
+\caption{Average number of weekly distinct reporters by 4-week period with release dates (since 2008).} \label{reporters_with_releases}
+\end{figure}
+
+\begin{figure}
+\includegraphics{commentators_with_releases.eps}
+\caption{Average number of weekly distinct commentators by 4-week period with release dates (since 2008).} \label{commentators_with_releases}
 \end{figure}

### 2019-04-07 article/article.tex
Snapshot (Jupyter notebook works fine).

@@
 %
-\author{Th\'eo Zimmermann\\
-$\Pi R^2$, Inria and
-IRIF, Paris-Diderot University\\
-Paris, France\\
-Email: theo@irif.fr
-\and
-Annal\'i Casanueva Art\'is\thanks{This work was partly funded by El Conacyt of the Mexican government.}\\
-Paris School of Economics\\
-Paris, France}
+%\author{Th\'eo Zimmermann\\
+%$\Pi R^2$, Inria and
+%IRIF, Paris-Diderot University\\
+%Paris, France\\
+%Email: theo@irif.fr
+%\and
+%Annal\'i Casanueva Art\'is\thanks{This work was partly funded by El Conacyt of the Mexican government.}\\
+%Paris School of Economics\\
+%Paris, France}
 %
@@
 We use the switch, from Bugzilla to GitHub, of the project's bug tracker, to analyze how a change in the bug reporting environment affects the number of bugs, reporters, bug comments and commentators.
-Using a Regression on Discontinuity design (RDD)~\cite{angrist2008mostly,angrist2014mastering,jacob2012practical,lee2010regression}, we find strong evidence that the change of platform increased the number of bug reports and comments by developers, and increased the number of distinct commentators each week among non-developers.
+Using a temporal Regression on Discontinuity design (RDD)~\cite{angrist2008mostly,angrist2014mastering,jacob2012practical,lee2010regression,hausman2018regression}, we find strong evidence that the change of platform increased the number of bug reports and comments by developers, and increased the number of distinct commentators each week among non-developers.
 Our results suggest that switching from Bugzilla to GitHub made the bug reporting process easier and more appealing but also opened the software development to more external contributors and stimulated discussion around bugs (notably between developers and non-developers).
@@
 In Sec.~\ref{context}, we present the Coq project and the context of the switch.
-In Sec.~\ref{hypotheses}, we formulate hypotheses on the impact of the switch.
+In Sec.~\ref{research_questions}, we present our research questions.
 In Sec.~\ref{migration}, we explain the migration process.
@@
 From 2007 to 2017, the Coq bug tracker platform was a self-hosted Bugzilla instance (before this, from 2001 to 2007 it was using JitterBug, a now discontinued bug tracking system, and before 2001 a simple mailing list).
-In 2017, developers had become quite accustomed to the look and feel of discussion threads on GitHub pull requests, which offered an arguably nicer experience than the Bugzilla bug tracker.
-Younger developers were frustrated by what they perceived as an older and heavier platform.
-The rest of the developers were convinced to migrate towards a non-self-hosted solution in August 2017, when the whole Coq website was shut down for more than a week  because an IT team was frightened by a scam bug report.\footnote{\url{https://sympa.inria.fr/sympa/arc/coq-club/2017-08/msg00040.html}}
+In 2017, in a context where all code changes were conducted through GitHub pull requests, the question of switching to GitHub's integrated bug tracker arose.
+After some preliminary testing showing that migrating all preexisting issues was doable, the Coq development team decided to switch to GitHub issues on October 4th, 2017.
 
-The expected advantages of the GitHub issue tracker over Bugzilla were:
+According to the developers we interviewed, the motivations for switching were:
 \begin{itemize}
-\item Better formatting, ability to edit comments, and more generally a more pleasant tool to use.
-According to some Coq developers, Bugzilla was unpleasant enough to discourage developers to use it to its full potential.
-\item Version control integration: Browsing between code, pull requests and bug reports is easier when everything is in the same place.
+\item Consolidating the development tools on a single, integrated, platform: browsing between code, pull requests and bug reports is easier without having to switch websites.
 Furthermore, it means a common notification, mention, and assignment system.
-\item Better indexing by search engines.
-\item Shared management of permissions, milestones, categories (labels), etc.
+\item GitHub's support for formatting and editing comments, cross-referencing and auto-closing issues, and more generally a more pleasant tool to use.
+On the other hand, many Coq developers perceived Bugzilla as unpleasant, old-fashioned, and very slow.
+It was a complex tool with many underused advanced features.
+\item No need to administrate (maintain and update) the bug tracker. The previous system was perceived as failing often by several developers.
 \item Easier to get started for newcomers (especially as many may already know GitHub and already have a GitHub account).
@@
 
-The possible drawbacks were mostly that GitHub provides a less advanced system with very few customization options. In particular:
+Again, according to the developers we interviewed, the principal risks associated with the switch of bug tracker were:
 \begin{itemize}
-\item The permission system is much less fine-grained.
-Giving triaging rights to users requires giving them ``write access'', which includes commit rights on non-protected branches.
-That's why the Coq GitHub repository now only has protected branches (and accidentally-pushed topic branches are promptly deleted).
-\item Creating a new bug report does not require filling a form, but typing a message.
-Fortunately, issue templates help give some hints towards the expected structure of the report, and being able to strip this template can actually give flexibility to developers who might open new issues for concerns that do not qualify as bugs.
-\item There is no import mechanism that allows to keep the number and author of preexisting reports.
-Fortunately, as we will see in Sec.~\ref{migration}, it was still possible to import bug reports while keeping most of the meta-information and renumbering the fewest possible bug reports.
-This was an important requirement for the switch.
+\item Losing control of a critical tool to a private company and its future decisions.
+In particular, GitHub is not an open source platform (so cannot be cloned elsewhere) and backing up data from GitHub becomes even more important but not trivial to do.
+\item Losing or corrupting bug tracking data during the transition from Bugzilla to GitHub.
+Even if, as we will see in Sec.~\ref{migration}, most of the data was correctly migrated, some meta-data had to be converted to text, and external links and documentation about the bug tracker could have become stale or inaccurate.
+%In practice, developers were quite happy about the result of the migration.
+\item A few developers additionally mentioned risks of losing useful features from Bugzilla, or discouraging developers for whom it would be hard to adapt to the new tool.
 \end{itemize}
 
-\section{Hypotheses}
-
-\label{hypotheses}
+\section{Research questions}
+After a consideration of the motivations and keeping the possible risks in mind, we found particularly interesting and compelling to study the following research questions more in detail to give insights into the effect of a bug tracker switch
+\label{research_questions}
 %ambre: I think this section should be expanded a bit to add motivation for the hypotheses
 
-Given the advantages of the GitHub issue tracker that were listed above, one could expect the switch to:
-\begin{description}
-\item[H1] Increase the bug tracking activity, in terms of the number of reported bugs;
-\item[H2] Extend the pool of bug reporters and commentators;
-\item[H3] Increase the bug tracking activity, in terms of the amount of interaction between users and developers.
-\end{description}
+%Given the advantages of the GitHub issue tracker that were listed above, one could expect the switch to:
+%\begin{description}
+%\item[H1] Increase the bug tracking activity, in terms of the number of reported bugs;
+%\item[H2] Extend the pool of bug reporters and commentators;
+%\item[H3] Increase the bug tracking activity, in terms of the amount of interaction between users and developers.
+%\end{description}
+
+\paragraph{RQ1: Impact on level of bug tracking activity.}
+%Do not forget the "why"
+
+\paragraph{RQ2: Impact on quality of bug tracking activity.}
+\paragraph{RQ3: Impact on the audience of the bug tracker.}
 
@@
 Using this API allows to do large requests (100 nodes in a single request) with just the information we need (thus both reducing the bandwidth usage and speeding up the extraction process).
-We provide the extracted data as CSV files and a Jupyter notebook\footnote{\url{https://github.com/Zimmi48/impact-of-switching-bug-trackers}.} with the code to request this data from GitHub, to load the CSV files, to run the pre-processing steps, and the analyses.%, and the robustness checks.
+In the supplementary materials, we provide the extracted data as CSV files and a Jupyter notebook with the code to request this data from GitHub, to load the CSV files, to run the pre-processing steps, and the analyses.%, and the robustness checks.
 
@@
 \paragraph{Excluding specific reporters}
-We have one specific reporter, Jason Gross, who is alone responsible for almost a quarter of all bug reports.
+We have one specific reporter who is alone responsible for almost a quarter of all bug reports.
 To avoid having the behavior of a single individual strongly impact the overall statistics, we exclude his comments, bug reports and the comments they received from our analysis.
@@
 
-The companion Jupyter notebook contains an extensive array of descriptive tables and figures, such as the distribution of bug reporting in the year and in the day, graphs of bug reporting for individual developers, etc.
+Figures~\ref{bug_nb_with_releases}, \ref{comments_with_releases}, \ref{reporters_with_releases} and \ref{commentators_with_releases} show the evolution of our four main outcome variables since January 2016 until March 2019.
+We choose to present the data starting in 2016 because on January 2016 the 8.5 version of Coq was released, and this marks the adoption of a more rapid release cycle (with a new major release every 6 to 10 months)\footnote{See \url{https://coq.inria.fr/distrib/V8.9.0/refman/credits.html}.}.
+This change in the development process could have impacted the activity on the bug tracker and made it not comparable.
+In Fig.~\ref{bug_nb_with_releases} and \ref{comments_with_releases}, each point represents the 4-week average of the number of issues and number of comments per day respectively for both developers and non-developers.
+In Fig.~\ref{reporters_with_releases} and \ref{commentators_with_releases}, each point represents the 4-week average of the number of distinct reporters and distinct commentators per week.
+In all four figures, the vertical red line shows the date of the bug tracker switch; the vertical black lines represent the date of major releases and the discontinuous horizontal lines represent the mean before and after the switch respectively.
+
+Considering the whole period (2016-2019), less issues are reported by developers than by non-developers.
+In a four-day period, on average, two issues will be opened by developers and five by non-developers (Fig.~\ref{bug_nb_with_releases}).
+On the contrary, developers post more comments than non-developers.
+On an average day, developers will post around five comments and non-developers two (Fig.~\ref{comments_with_releases}).
+This is not surprising because some bugs found by developers are resolved directly without opening an issue and because issue reports are generally answered by developers.
+There are more distinct non-developers than developers who open issues in an average week (around six distinct non-developers and two distinct developers, Fig.~\ref{reporters_with_releases}).
+There are slightly more distinct non-developers (per week) that comment than developers.
+More precisely, on an average week, around six distinct non-developers and around 5.5 developers comment an issue (Fig.~\ref{commentators_with_releases}).
+
+%It is important to note that we are close to the theoretical saturation of the possible number of distinct developer commentators (already all distinct developers are commenting on weekly basis).
+For all outcomes, the mean before the switch is statistically significantly lower than the mean after the switch both for developers and non developers.
+For some outcomes, the difference in activity patterns before and after the switch is clear to the naked eye.
+In Fig.~\ref{bug_nb_with_releases} we see an increase in the number of issues reported by developers.
+In Fig.~\ref{comments_with_releases} we see an increase in the number of comments by developers with several data points after the release that are well above the highest point before the release.
+Finally, in Fig.~\ref{commentators_with_releases} we see a clear increase in the number of distinct non-developer commentators with the great majority of points after the switch above the great majority of points before the switch.
+Some other differences may be appreciated but are less clear to the naked eye than those presented.
+
+%The companion Jupyter notebook contains an extensive array of descriptive tables and figures, such as the distribution of bug reporting in the year and in the day, graphs of bug reporting for individual developers, etc.
 We only present a subset of these results.
@@
 
-\subsection{Influence of new releases}
-\label{influence_release}
+%\subsection{Influence of new releases}
+%\label{influence_release}
 
-\paragraph{Influence on the number of bug reports}
-We expect release dates to be correlated to peaks of activity on the bug tracker, in particular peaks of new bug reports.
+%\paragraph{Influence on the number of bug reports}
+%We expect release dates to be correlated to peaks of activity on the bug tracker, in particular peaks of new bug reports.
 Indeed, users will generally test new releases, for instance by trying to port preexisting projects to the new version, which might lead them to find new bugs (either regressions or bugs related to new features).
@@
 
-The highest peak is in 2015, after the first beta release of version 8.5.
+%The highest peak is in 2015, after the first beta release of version 8.5.
 This was the first release in more than two years and it contained five big new features (``the result of five specific long-term projects''\footnote{Quote from:\\ \url{https://coq.inria.fr/distrib/V8.9.0/refman/credits.html\#credits-version-8-5}\\
-The next quotes are coming from the following sections of the same chapter of the Coq reference manual.}).
+%The next quotes are coming from the following sections of the same chapter of the Coq reference manual.}).
 Stabilizing these features and their (unplanned) interactions required more than a year of testing and three beta releases.
@@
 \includegraphics{bug_nb_with_releases.eps}
-\caption{Number of bug reports by 4-week period with release dates (since 2008).} \label{bug_nb_with_releases}
+\caption{Number of issues per day (averaged by 4-week periods) with release dates (since 2016).} \label{bug_nb_with_releases}
 \end{figure}
 
-\paragraph{Influence on the number of comments}
-In Fig.~\ref{comments_with_releases}, we can see that release dates are correlated to peaks of comments by developers (and not so much by other people).
+%\paragraph{Influence on the number of comments}
+%In Fig.~\ref{comments_with_releases}, we can see that release dates are correlated to peaks of comments by developers (and not so much by other people).
 This is explained by the fact that bug reports are generally answered by developers, thus any peak in bug reporting is bound to induce a peak in bug commenting by developers.
@@
 \includegraphics{comments_with_releases.eps}
-\caption{Number of bug comments by 4-week period with release dates (since 2008).} \label{comments_with_releases}
+\caption{Number of comments per day (averaged by 4-week periods) with release dates (since 2016).} \label{comments_with_releases}
 \end{figure}
@@
 \includegraphics{reporters_with_releases.eps}
-\caption{Average number of weekly distinct reporters by 4-week period with release dates (since 2008).} \label{reporters_with_releases}
+\caption{Number of weekly distinct reporters (averaged by 4-week periods) with release dates (since 2016).} \label{reporters_with_releases}
 \end{figure}
@@
 \includegraphics{commentators_with_releases.eps}
-\caption{Average number of weekly distinct commentators by 4-week period with release dates (since 2008).} \label{commentators_with_releases}
+\caption{Number of weekly distinct commentators (averaged by 4-week periods) with release dates (since 2016).} \label{commentators_with_releases}
 \end{figure}
@@
 
+\subsection{Quantitative}
 We exploit the fact that the switch from Bugzilla to GitHub can be seen as a clear cutoff in time to use a temporal sharp Regression in Discontinuity design (RDD) to estimate the effects of the switch on our outcome variables; to determine if these estimated effects are statistically significant; and to interpret these effects in a causal way.
@@
 
+\subsection{Qualitative}
+
+In order to obtain qualitative insights into the effects of the switch that will help us put in context and interpret the quantitative results, we conducted semi-structured interviews of all developers of Coq that were active at the time of the switch (with the exception of the developer that conducted the migration).
+A total of 9 interviews were conducted between 19th March and 4th April 2019.
+Some were done face to face and the others by video-conference with an average length of 33.5 minutes per interview.
+Both authors were present during the interviews and all the interviews were recorded.
+Some basic questions were asked to all of the interviewees and according to the answers some additional questions were also asked.
+The answers were coded by both authors in an attempt to minimize errors.
+Code and questions can be found in the supplementary material.
+
 \section{Results}
@@
 \centering
-\caption{Estimated impact of the switch on the number of bugs.
+\caption{Estimated impact of the switch on the number of issues.
 Coefficients are highlighted if they are statistically significant with $p<0.1$ ($\dagger$), $p<0.05$ (*), $p<0.01$ (**) or $p<0.001$ (***).
@@
 \includegraphics{bug_nb_rd.png}
-\caption{Average number of bug reports per day before and after the migration (with fitting lines and confidence intervals from the regression results).} \label{bug_nb_rd}
+\caption{Number of issues per day before and after the migration (with fitting lines and confidence intervals from the regression results, and points corresponding to average values over two-week periods).} \label{bug_nb_rd}
 \end{figure}
@@
 \includegraphics{reporter_nb_rd.png}
-\caption{Average number of distinct reporters each week before and after the migration (with fitting lines and confidence intervals from the regression results).} \label{reporter_nb_rd}
+\caption{Number of weekly distinct reporters before and after the migration (with fitting lines and confidence intervals from the regression results, and points corresponding to average values over two-week periods).} \label{reporter_nb_rd}
 \end{figure}
@@
 \includegraphics{comment_nb_rd.png}
-\caption{Average number of bug comments per day before and after the migration (with fitting lines and confidence intervals from the regression results).} \label{comment_nb_rd}
+\caption{Number of comments per day before and after the migration (with fitting lines and confidence intervals from the regression results, and points corresponding to average values over two-week periods).} \label{comment_nb_rd}
 \end{figure}
@@
 \includegraphics{commentator_nb_rd.png}
-\caption{Average number of distinct commentators each week before and after the switch (with fitting lines and confidence intervals from the regression results).} \label{commentator_nb_rd}
+\caption{Number of weekly distinct commentators before and after the switch (with fitting lines and confidence intervals from the regression results, and points corresponding to average values over two-week periods).} \label{commentator_nb_rd}
 \end{figure}
@@
 
-The biggest concern that we have in our analysis is the fact that the date of the switch is close to the release of a new version of Coq.
+The biggest concern that we have in our analysis is the fact that the date of the switch is close to the release date of a new version of Coq.
 The release of version 8.7.0 took place on October 17\textsuperscript{th}, and the migration on October 18\textsuperscript{th}.
@@
 %Second, we can also argue that there should have been a similar (or higher) peak in the month before the switch since the first beta of 8.7 had been released then (and beta releases generally result in higher peaks).
-To confirm that this is not what is driving our results, we conduct two robustness checks with our data.
+To confirm that it is not the migration that is driving our results, we do the exact same analysis that we do for the bug tracker switch, replacing the cutoff date by the date of all six major releases since 2009 (once we remove the latest release because we do not have yet enough data points after the release).
+We show that the only significant effect in the bug-tracker activity after a major release is the effect after the major release that match the switch.
+It is also, for the majority of the outcomes, by far the effect with the highest magnitude.
+
+%To provide some evidence that the date of the switch is not associated with an increase in the activity of open source software due to external factors (which will invalidate our results as explained before), we replicate our analysis for a comparable software: another French medium size open-source project: OCaml.
+We do not observe any increase in the bug-tracking activity for this project after the Coq bug-tracker switch date.
+If anything we see a decrease in activity so if there was an external factor that affected french open-source projects decreasing their activity.
+we will be underestimating the effect of the switch.
+
+%two robustness checks with our data.
 The first consists in removing the two weeks following the 8.7.0 release (and the switch) from our Regressions in Discontinuity.
@@
 Our results still hold with this specification.
+
+Temporal RDD have some notable differences with other types of RDD \cite{hausman2018regression}.
+One major difference is the capacity of anticipation and reaction to the treatment that affect directly the outcome.
+To overcome this potential problem, we redo our analysis with a donut RD (i.e.
+we exclude the data points just after and just before the migration, which are potentially problematic), following \cite{hausman2018regression}.
+Our results still hold with a slight decrease in the level of significativity but keeping the significativity at 10\%.
 
@@
 Our main results are robust to these changes.
-
+All results presented in this section can be found in the Jupyter notebook provided as a supplemental material. 
 \subsection{External validity}
@@
 
+% Takeaways
 % When to switch?
@@
 We also provide the XML export of Bugzilla that was used for the migration and the migration script.
-Everything is available from \url{https://github.com/Zimmi48/impact-of-switching-bug-trackers}.
+% XXXXX We also provide the interview questions and coding.
+%Everything is available from \url{XXXXXXX}.
 %\paragraph{}
@@
 
-\noindent\footnotesize{\textbf{Acknowledgements.}}
-We would like to thank Ambre Williams for valuable advice and help on the design and implementation of the data analyses and C\'ecile Rottner for general advice, support, and proof-reading.
-We would like to thank Andriy Berestovskyy for providing the Bugzilla import tool, and Martin Michlmayr for doing the first steps in porting our modifications upstream.
+%\noindent\footnotesize{\textbf{Acknowledgements.}}
+%We would like to thank Ambre Williams for valuable advice and help on the design and implementation of the data analyses and C\'ecile Rottner for general advice, support, and proof-reading.
+%We would like to thank Andriy Berestovskyy for providing the Bugzilla import tool, and Martin Michlmayr for doing the first steps in porting our modifications upstream.
 

### 2019-07-26 article/article.tex
Final version of the ICSME paper.

@@
-%\documentclass[runningheads]{llncs}
 \documentclass[conference]{IEEEtran}
-%\IEEEoverridecommandlockouts
+
+\usepackage[noadjust]{cite}
+\usepackage{balance}
+
+\IEEEoverridecommandlockouts
 %
@@
 %
-%\author{Th\'eo Zimmermann\\
-%$\Pi R^2$, Inria and
-%IRIF, Paris-Diderot University\\
-%Paris, France\\
-%Email: theo@irif.fr
-%\and
-%Annal\'i Casanueva Art\'is\thanks{This work was partly funded by El Conacyt of the Mexican government.}\\
-%Paris School of Economics\\
-%Paris, France}
+\author{Th\'eo Zimmermann (theo@irif.fr)\\
+Universit\'e de Paris, IRIF, CNRS, F-75013 Paris, France\\
+Inria, $\pi.r^2$ project-team
+\and
+Annal\'i Casanueva Art\'is\thanks{This work was partly funded by El Conacyt of the Mexican government.}\\
+Paris School of Economics, F-75014 Paris, France}
 %
@@
 
-We first report on the switch itself, including the migration of preexisting bug reports.
-Then we analyze data from before and after the switch using a Regression on Discontinuity analysis, a novel methodology in the context of empirical software engineering.
+We first report on the switch itself, including the migration of preexisting issues.
+Then we analyze data from before and after the switch using a regression discontinuity design, an econometric methodology imported from quantitative policy analysis.
 We complete this quantitative analysis with qualitative data from interviews with developers.
@@
 \begin{IEEEkeywords}
-bug tracker, switch, migration, bug report, issue, GitHub, Bugzilla, open source, data mining, Regression on Discontinuity, interviews
+bug tracker, switch, migration, bug report, issue, GitHub, Bugzilla, open source, data mining, regression discontinuity design, RDD, interviews
 \end{IEEEkeywords}
@@
 Bug reporting is an essential part of software development.
-In the context of an open source project, the bug reporting and fixing process is generally done on a public bug tracking platform, and, in the absence of paid testers, it depends a lot on the good will of independent users.
+In the context of an open source project, the bug reporting and fixing process is generally done on a public bug tracking platform, and, in the absence of paid testers, it depends a lot on the goodwill of independent users.
 Therefore, it involves a strong social component.
 
-However, having an open bug tracking system is not enough to get the bugs reported.
-In addition to the software having enough users, the process of bug reporting must be easy and appealing.
-Therefore, creating an environment that eases bug reporting and makes it more appealing may lead to an increase in bug reporting activity (from both developers and independent users), which is well associated with an increase in software quality.
-First, if the software has not changed, receiving more bug reports indicates that more defects have been found, and not a sudden apparition of defects (Linus's Law ``given enough eyeballs, all bugs are shallow''~\cite{raymond1999cathedral,van2009shallow}).
-Moreover, assuming that developers are equipped to cope with the increased number of incoming reports~\cite{anvik2005coping, davidson2011coping}, even duplicate reports are not necessarily harmful~\cite{Bettenburg2008}.
-Second, more activity on the bug tracker can also mean users are helping to reproduce bugs, produce traces, etc.
+However, having an open bug tracking system is not enough to attract participants.
+In addition to the software having enough users, the process of opening an issue (be it bug report or feature request) must be easy and appealing.
+Therefore, creating a favorable bug tracking environment may lead to an increase in bug tracking activity (from both developers and independent users), which may result in an increase in software quality.
+First, the participation of more users is helpful to find a larger proportion of bugs~\cite{raymond1999cathedral,van2009shallow}.
+Second, assuming that developers are equipped to cope with the increased number of incoming reports~\cite{anvik2005coping, davidson2011coping}, even duplicate reports are not necessarily harmful~\cite{Bettenburg2008}.
+Third, more activity on the bug tracker can also mean users are helping to reproduce bugs, produce traces, etc.
 and thus are working with the developers to get the bugs fixed~\cite{breu2010information}.
-More generally, reporting new bugs and discussing existing bug reports has been shown to be an important step on the path to becoming an active contributor of an open source project~\cite{jensen2007role, nakakoji2002evolution, ye2003toward}.
+More generally, opening issues and discussing existing ones has been shown to be an important step on the path to becoming an active contributor of an open source project~\cite{jensen2007role, nakakoji2002evolution, ye2003toward}.
 
-Despite the importance of the bug tracking environment that we have just highlighted, the impact of a change in this environment has rarely been studied.
-It is generally difficult to compare the number of bug reports in different environments because it is likely to depend  not just on the reporting environment but also on each particular software and its development dynamics, which would make such an analysis not compelling.
-To do a valid comparison, one would ideally need to compare two bug reporting environments for the same software, used by the same users, to report the same bugs, in the same period.
-This is of course impossible.
-A context approaching this ideal scenario would be to compare the number of bug reports of a particular software before and after the switch to a new environment.
-Yet, substantial changes in the environment of bug reporting (e.g.
-changes of bug tracker platform) are rare because once a bug tracking platform is in place, any change comes with a significant cost.
+Despite the importance of the bug tracking environment that we have just highlighted, the impact of a change in this environment has rarely been studied, whether it is the bug tracking platform in full, a feature, or a policy.
+Furthermore, since the choice of bug tracker is not independent from the characteristics of the project, a comparison across projects using different bug trackers would not be sufficient to draw causal conclusions.
 
 In this paper, we study the case of the Coq proof assistant, a medium-sized open source software project.
-We use the switch, from Bugzilla to GitHub, of the project's bug tracker, to analyze how a change in the bug reporting environment affects the number of bugs, reporters, bug comments and commentators.
-Using a temporal Regression on Discontinuity design (RDD)~\cite{angrist2008mostly,angrist2014mastering,jacob2012practical,lee2010regression,hausman2018regression}, we find strong evidence that the change of platform increased the number of bug reports and comments by developers, and increased the number of distinct commentators each week among non-developers.
-Our results suggest that switching from Bugzilla to GitHub made the bug reporting process easier and more appealing but also opened the software development to more external contributors and stimulated discussion around bugs (notably between developers and non-developers).
-
 The contributions of this paper are practical, empirical, and methodological.
-First, we improved an existing bug tracker migration tool to handle thousand of bug reports while preserving meta-data, and we helped the Coq development team migrate their bug tracker from Bugzilla to GitHub.
-Second, we analyze the causal impact of this switch on the bug tracking activity.
-Third, we introduce to the field of software engineering a state-of-the-art method for quantitative public policy evaluation: Regression in Discontinuity Design.
-In particular, we explain this method to give the basic intuition behind; we point the reader to the both introductory and state-of-the-art literature and we illustrate one application.
+First, we improved an existing bug tracker migration tool to handle thousands of issues while preserving meta-data, and we helped the Coq development team switch their bug tracker from Bugzilla to GitHub.
+Second, we analyze the causal impact of this switch on the bug tracking activity through mining repository data, and we interpret the results through qualitative assessment based on interviews with developers.
+Third, we demonstrate the application of regression discontinuity design (RDD), an econometric method to derive causality, in the context of empirical software engineering; we explain the method to give the basic intuition behind it; we point the reader to introductory, basic, and state-of-the-art literature~\cite{angrist2008mostly,angrist2014mastering,jacob2012practical,lee2010regression,hausman2018regression}.
 
@@
 In Sec.~\ref{migration}, we explain the migration process.
-In Sec.~\ref{data}, we describe the data collection and pre-processing method, and we list the variables of interest.
+In Sec.~\ref{data}, we describe the data collection and pre-processing, and we list the variables of interest.
 In Sec.~\ref{descriptive-stats}, we give a preliminary view of the bug tracking activity.
-In Sec.~\ref{methodology}, we explain the Regression on Discontinuity methodology, of which we present the results in Sec.~\ref{results}.
-We interpret the results and relate them to our hypotheses in Sec.~\ref{discussion}.
-We discuss the threats to validity and  present robustness checks in Sec.~\ref{threats}.
+In Sec.~\ref{methodology}, we explain our quantitative and qualitative methodology.
+We present quantitative results in Sec.~\ref{results}.
+We present qualitative results, interpret the quantitative results, and relate them to our research questions in Sec.~\ref{discussion}.
+We discuss threats to validity and present robustness checks in Sec.~\ref{threats}.
 We conclude in Sec.~\ref{conclusion}.
@@
 
-While there is a very large literature on many aspects of bug reporting (see~\cite{strate2013literature, zhang2016literature} for literature reviews), there is a lack of literature measuring the influence of the bug reporting environment on the bug reporting activity or on any other aspects of software development.
+While there is a very large literature on many aspects of bug reporting (see~\cite{strate2013literature, zhang2016literature} for literature reviews), there is a lack of literature measuring the influence of the bug reporting environment on the bug reporting activity, or on any other aspects of software development.
 To our knowledge, there is no previous work measuring the impact of a change in bug reporting environment.
-More generally, there is little literature that compare bug trackers.
+More generally, there is little literature that compares bug trackers.
 
@@
 A simple, but quite limited, way to compare bug trackers is to look at the features that each of them proposes.
-
 Karre et al.~\cite{karre2017does} compared 31 bug trackers feature-wise, and gathered these tools in four clusters.
 Bugzilla belongs to the cluster of bug trackers with many features (together with RedMine and Mantis), while GitHub belongs to a cluster of bug trackers attached to source code management systems (together with Savannah and BitBucket).
-%Bugzilla has 20 of the 23 features they listed, and scores the highest number of features (ex.
-aeqo with RedMine).
-They are closely followed by Mantis, and two other bug trackers, with 19 features.
-GitHub only has 15 features out of the 23 listed (overall 11 listed systems have more features).
+Abaee and Guru~\cite{abaee2010enhancement} listed the features of four commercial bug trackers, and proposed their own bug tracker with some new features, but with no evaluation.
 
-In the case of the Coq project, the loss of some specific features from Bugzilla was one of the arguments against the switch.
-Nonetheless, GitHub has some features that Bugzilla hasn't, and it is clearly not the number of features that determines which software is better or more adapted to the specific use case.
-
-%We will note that GitHub has very recently added a number of new features to the issue tracker, but these were not present at the time of the switch (just-in-time duplicate retrieval, issue deletion, transfer and pinning, support for multiple issue templates, etc.).
-
-Additionally, Karre et al.
-recommended some missing features, including better support for migrating bug tracking data from a system to another.
-Our work has contributed to improve the situation in that sense.
-Nonetheless, it would be preferable if there existed a standard structured format for storing bug tracking information and all major bug tracking systems supported exporting to and importing from this format.
-
-Abaee and Guru~\cite{abaee2010enhancement} listed the features of four commercial bug trackers, and proposed their own bug tracker with some new features, but without any evaluation.
-
-There is much more literature proposing new features to add to bug trackers, but they are rarely evaluating the impact of adding such features on the bug tracking activity, even when a prototype was presented.
-
+Many papers propose new features to add to bug trackers, but they rarely evaluate the impact of adding such features, even when a prototype was presented.
 Baysal et al.~\cite{baysal2013situational} identified the need for personalized issue tracking systems after interviewing twenty Mozilla developers.
 They proposed a Bugzilla extension to address this need, and gave a, mostly qualitative, assessment by interviewing developers using their tool in~\cite{baysal2014no}.
-
 Bortis~\cite{bortis2016porchlight} developed a bug triaging tool and evaluated how users interacted with it.
 However, there was no evaluation of its impact in the context of an actual software project.
+Just et al.~\cite{just2008towards} surveyed developers from three large open source projects (Apache, Eclipse, and Mozilla; all three projects use Bugzilla) to identify missing features and provided recommendation to design better bug tracking systems.
 
-Just et al.~\cite{just2008towards} surveyed developers from three large open source projects (Apache, Eclipse, and Mozilla; all three projects use Bugzilla) to identify missing features and provided recommendation to design better bug tracking systems.
-%They remark that bug trackers are often mere interfaces to bug databases and ask too much information from end-users without guiding them.
-They recommend providing guidance to bug reporters depending on their level of expertise.
-This issue is indeed something that we have noticed with Bugzilla and other bug trackers of the same category.
-Users are encouraged to self-triage and often do so inaccurately.
-On the contrary, GitHub only provides title and body fields to bug reporters and leaves triaging (labeling) to the contributors with write access.
-Nonetheless, the issue templates can be used to fill the bug report body field with advice on how to present a bug report.
-GitHub has recently added support for multiple issue templates which would allow to let the reporter choose between, for instance "My first bug report", "My first feature request", "Bug report (experienced reporter)", and "Feature request (experienced reporter)", with varied requirements and levels of explanations...
-
-Our article goes beyond a descriptive or normative perspective and quantitatively measures the causal impact of the change of the bug tracking platform (a crucial part of the bug tracking environment) on various aspects of bug reporting activity.
+Our article goes beyond a descriptive or normative perspective and quantitatively measures the causal impact of the change of the bug tracking platform (a crucial part of the bug tracking environment) on various aspects of bug tracking activity, and gives qualitative insights for its interpretation.
 
@@
 
-Another way of comparing bug tracking systems could have been to conduct large-scale studies of many software projects using various bug trackers and provide some clues of the differences induced by the use of the different bug tracking systems.
+Another way of comparing bug tracking systems could have been to conduct large-scale studies of many software projects using various bug trackers, and derive some differences associated with the use of the various systems.
 However, there are no such comparative studies in the literature.
@@
 Sowe et al.~\cite{sowe2013multi} noted that most preceding literature had only been comparing few projects at a time, usually using a single bug tracking system.
-In their study, they addressed this in part by studying hundreds of projects, but they did not mention which bug trackers the projects they compared used.\looseness=-1
+In their study, they addressed this in part by studying hundreds of projects, but they did not mention which bug trackers the projects they compared used.
 
-Bissyandé et al.~\cite{bissyande2013got} published the same year a larger scale investigation using ten thousand projects' bug tracking data, analyzing such things as the correlation between a project's success and its bug tracking activity.
+Bissyand\'e et al.~\cite{bissyande2013got} published the same year a larger scale investigation using ten thousand projects' bug tracking data, analyzing such things as the correlation between a project's success and its bug tracking activity.
 All of the projects studied were using the same bug tracker (GitHub's).
@@
 Francalanci and Merlo~\cite{francalanci2008empirical} analyzed the bug fixing process by studying closed bugs from nine open source projects (four using JIRA, four using SourceForge's bug tracker, and one using another bug tracker).
-They did not, however, try to correlate the bug tracking activity with the bug tracking system used.
+They did not, however, try to correlate the bug tracking activity with the bug tracking system.\looseness=-1
 
-%According to the survey in~\cite{karre2017does}, Bugzilla is the most used bug tracker (respondent were 32\% to use Bugzilla and 15\% to use GitHub).
-There is an obvious selection bias due to the choice of open source project contributors this survey was sent to.
-Overall, given the sheer number of projects hosted on GitHub, it does not make any doubt that the GitHub issue tracker is the most used bug tracker in number of projects.
-But this raw fact does not matter much, given that very small projects will use their bug tracker very little.
-\subsection{Switching developer tools}
+\subsection{Switching developer tools or processes}
 
-While, to the best of our knowledge, there is no work evaluating the impact of switching bug trackers, some studies have been conducted to evaluate the impact of switching other developer tools.
+While, to the best of our knowledge, no previous work has evaluated the impact of switching bug trackers, some studies have evaluated the impact of switching other developer tools or processes.
 Since bug trackers are important developer tools, our article also relates and contributes to this literature.
@@
 As with bug trackers, the move is decided because of anticipated benefits: among them is the openness to external contributors.
-But the switch incurs challenges in the migration process (in particular, developers are very concerned to preserve the full history of the project) and some features are lost in the new tool (in particular, the monotonically increasing revision numbers).
+But the switch incurs challenges in the migration process (in particular, developers are very concerned to preserve the full history of the project) and some features are lost in the new tool (in particular, the chronologically ordered revision numbers).
 Muşlu et al.~\cite{mucslu2014transition} performed a similar study based on interviews and a survey.
@@
 
-Squire~\cite{squire2015should} measured the impact of switching developer support channels from mailing lists and self-hosted forums to Stack Overflow, by first identifying the expectations of 20 projects which had officially decided on such a move, then by studying the impact on developer participation and response time for a selection of four such projects.
-On the other hand, Vasilescu et al.~\cite{vasilescu2014social} compared the behavior of the same users on the R mailing list and the Stack Overflow platform.
-Contrarily to a bug tracker which is expected to be centralized for a given project, support channels can be diverse.
-Project leaders can decide to abandon their mailing list to focus exclusively on Stack Overflow, but they do not decide to ``open'' a support channel on Stack Overflow: it can exist nevertheless and the switch can be organic.
+Squire~\cite{squire2015should} measured the impact of switching developer support channels from mailing lists and self-hosted forums to Stack Overflow, and compared it with the expectations of projects having decided on such a move,
+whereas Vasilescu et al.~\cite{vasilescu2014social} compared the behavior of the same users during the same period on the R mailing list and on Stack Overflow.
+
+Alencar da Costa et al.~\cite{da2018impact} evaluated the impact of switching from traditional (long) release cycles to faster ones on the time it takes for a bug fix to get released.
+They did a quantitative assessment based on a case study on the Firefox project, and completed it by surveying developers.
+
+\subsection{RDD in empirical software engineering}
+
+While RDD is a well-established method in quantitative public policy evaluation, it has been seldom used in empirical software engineering.
+The only two studies we know of \cite{zhao2017impact,trockman2018adding} analyze hundreds of thousands of GitHub repositories, with time as the rating variable.
+In our paper, we show how this methodology can be used to derive causality from a smaller dataset (the impact of a change on a single active project).
+We demonstrate precisely how to use this method, with specific details on the particular case of time series, how to visualize the results in a compelling way, what are the internal threats to validity, and how to mitigate them with appropriate robustness checks.
 
@@
 
-% XXXX justify more the medium-size qualification
-
 The development is currently centered around the GitHub platform.
-The Coq repository counts more than 25,000 commits.
-Some of these commits date back from 1999, when the code received major architectural changes.
-The GitHub repository was initially created as a mirror of the official repository (which itself was migrated from SVN to git in 2013) and really started attracting the interest of the development team as a way to open up the development to external contributions, especially starting with the first Coq coding sprint in 2015 (later renamed into the Coq Implementors Workshop and held every year in France).
-From then on, core developers started using pull requests for their own changes more and more often, and it became the norm around the beginning of 2017 when systematic continuous integration tests were introduced on every pull request.
-In the last two years, more than 2,200 pull requests have been opened.
+The GitHub repository, which counts more than 25,000 commits starting in 1999, was initially created as a mirror of the official SVN repository.
+It attracted the interest of the development team as a way to open up the development to external contributions, especially starting with the first Coq coding sprint in 2015.
+Then, core developers began using pull requests for their own changes more frequently, and it became the norm around the beginning of 2017 when systematic continuous integration testing was introduced.
+Since 2017, more than 3,000 pull requests have been opened.
 
@@
 In 2017, in a context where all code changes were conducted through GitHub pull requests, the question of switching to GitHub's integrated bug tracker arose.
-After some preliminary testing showing that migrating all preexisting issues was doable, the Coq development team decided to switch to GitHub issues on October 4th, 2017.
+After some preliminary testing showing that migrating all preexisting issues was doable, the Coq development team approved the switch to GitHub issues on Oct. 4\textsuperscript{th}, 2017.
 
@@
 \begin{itemize}
-\item Consolidating the development tools on a single, integrated, platform: browsing between code, pull requests and bug reports is easier without having to switch websites.
+\item Consolidating the development tools on a single, integrated, platform: browsing between code, pull requests and issues is easier without having to switch websites.
 Furthermore, it means a common notification, mention, and assignment system.
 \item GitHub's support for formatting and editing comments, cross-referencing and auto-closing issues, and more generally a more pleasant tool to use.
-On the other hand, many Coq developers perceived Bugzilla as unpleasant, old-fashioned, and very slow.
+Many developers perceived Bugzilla as unpleasant, old-fashioned, and very slow.
 It was a complex tool with many underused advanced features.
-\item No need to administrate (maintain and update) the bug tracker. The previous system was perceived as failing often by several developers.
-\item Easier to get started for newcomers (especially as many may already know GitHub and already have a GitHub account).
+\item No need to administrate (maintain and update) the bug tracker. Almost half of the developers complained about the previous bug tracker failing often.
+\item Easier to get started for newcomers (especially as many may already know GitHub and have a GitHub account).
 \end{itemize}
@@
 \item Losing control of a critical tool to a private company and its future decisions.
-In particular, GitHub is not an open source platform (so cannot be cloned elsewhere) and backing up data from GitHub becomes even more important but not trivial to do.
+In particular, GitHub is not an open source platform (so cannot be cloned elsewhere) and backing up data from GitHub becomes even more important, but not trivial to do.
 \item Losing or corrupting bug tracking data during the transition from Bugzilla to GitHub.
-Even if, as we will see in Sec.~\ref{migration}, most of the data was correctly migrated, some meta-data had to be converted to text, and external links and documentation about the bug tracker could have become stale or inaccurate.
-%In practice, developers were quite happy about the result of the migration.
-\item A few developers additionally mentioned risks of losing useful features from Bugzilla, or discouraging developers for whom it would be hard to adapt to the new tool.
+During the migration, some meta-data had to be converted to text, and external links and documentation about the bug tracker could have become stale or inaccurate.
+\item A few developers additionally mentioned risks of losing useful features from Bugzilla, or discouraging some developers for whom it would be hard to adapt to the new tool.
 \end{itemize}
@@
 \section{Research questions}
-After a consideration of the motivations and keeping the possible risks in mind, we found particularly interesting and compelling to study the following research questions more in detail to give insights into the effect of a bug tracker switch
+
 \label{research_questions}
-%ambre: I think this section should be expanded a bit to add motivation for the hypotheses
 
-%Given the advantages of the GitHub issue tracker that were listed above, one could expect the switch to:
-%\begin{description}
-%\item[H1] Increase the bug tracking activity, in terms of the number of reported bugs;
-%\item[H2] Extend the pool of bug reporters and commentators;
-%\item[H3] Increase the bug tracking activity, in terms of the amount of interaction between users and developers.
-%\end{description}
+Our research questions cover the spectrum of the possible impact such a switch could have on the bug tracking activity:
 
-\paragraph{RQ1: Impact on level of bug tracking activity.}
-%Do not forget the "why"
+\paragraph{Impact on level of %bug tracking
+activity (RQ1)}
 
-\paragraph{RQ2: Impact on quality of bug tracking activity.}
-\paragraph{RQ3: Impact on the audience of the bug tracker.}
+Did the switch to the GitHub bug tracker, with the various expected benefits that such a switch would bring, impact the level of activity on the bug tracker?
+If so, who are the participants whose level of activity was impacted, and how can we explain such changes?
 
-All of this could lead in turn to a greater openness and transparency of the development and an increased engagement of users toward it, changing (and probably improving) the dynamics of a crucial part of software development.
+\paragraph{Impact on quality of %bug tracking
+activity (RQ2)}
+
+Did the switch to GitHub impact the quality of opened issues, and did it impact the quality of interactions between developers, and between developers and non-developers?
+
+\paragraph{Impact on the audience of the bug tracker (RQ3)}
+
+Did the switch to GitHub help onboard more new reporters and commentators? Did it increase the number of distinct non-developers that regularly take part to the bug tracking activity?
 
@@
 
-The only complicated part of the bug tracker switch was the migration of preexisting bug reports.
-We reused a tool by Andriy Berestovskyy\footnote{Tool available at \url{https://github.com/berestovskyy/bugzilla2github}.} which is designed to import Bugzilla reports (extracted as an XML dump) to GitHub using its REST API.
-The bugs are imported in an order designed to preserve numbers whenever possible.
-Bugs whose number is unavailable (e.g.
+As explained in Sec.
+\ref{context}, one of the main risks (and deterrents) to a switch of bug tracker was the possible loss or corruption of bug tracking and associated data.
+In particular, the Coq development team wanted preexisting issues to be migrated to the new bug tracker, while keeping the issue numbers as much as possible: indeed, these numbers are important because they are mentioned in many places (changelog, test-suite, other issues and pull requests, commits, external discussion boards, etc.).\looseness=-1
+
+We reused and adapted a tool~\cite{bugzilla2github} which is designed to import Bugzilla reports (extracted as an XML dump) to GitHub using its REST API~\cite{github_REST_API}.
+The issues are imported in an order designed to preserve numbers whenever possible.
+Issues whose number is unavailable (e.g.
 because the number is already taken by a GitHub pull request) are postponed and renumbered.
-We implemented several improvements to make the tool better fit the needs of the Coq project:
-\begin{itemize}
-\item \textit{Allowing non-consecutive bug numbers:} The imported set of bug reports had some holes in the numbering due to deleted bugs.
-We use postponed bugs to fill the holes.
-This improvement has now been integrated upstream.
-\item \textit{Saving a table of correspondence for renumbered bugs:} This was used later to redirect the old bug URLs to the new ones.
-\item \textit{Using the GitHub issue import beta API and overcoming the GitHub rate limits:} Creating a new issue or a new comment through the normal GitHub REST API will trigger notifications (for people who are watching the repository or are mentioned in the issue thread).
-Therefore, GitHub chooses to impose a strict rate limit on these actions, which prevented using this tool for importing more than a few hundred bugs.
-Fortunately, GitHub provides a (beta) issue import API which, in addition to not triggering notifications, also allows importing one bug, its comments, and meta information such as closed status and assignee in a single request (thus reducing the duration to import 4900 bugs to just a few hours).
-Furthermore, using this API allows to keep the dates of imported bug reports and comments, which is very useful to this study.
-We didn't manage to import correct closing dates for every bug report, so we will not study the impact of the bug tracker switch on the time it takes to close a bug.
+We implemented several changes to make the tool better fit the needs of the Coq project (the main improvements having now been integrated upstream):
 
-This improvement has now been integrated upstream as an optional setting.
-\end{itemize}
+\paragraph{Allowing non-consecutive numbers} The imported set of issues had some holes in the numbering due to deleted issues. We use postponed issues to fill the holes.
+\paragraph{Saving a table of correspondence for renumbered issues} This was used later to redirect the old Bugzilla URLs.
+\paragraph{Using GitHub's issue import API and overcoming GitHub's rate limits} Creating a new issue or a new comment through the normal GitHub REST API triggers notifications (for people who are watching the repository or are mentioned in the issue thread).
+Therefore, GitHub chooses to impose a strict rate limit on these actions, which prevented using this tool for importing more than a few hundred issues.
+Fortunately, GitHub provides a (beta) issue import API which, in addition to not triggering notifications, also allows importing one issue, its comments, and meta information such as closed status and assignee in a single request (thus reducing the duration to import 4900 issues to just a few hours).
+Furthermore, using this API allows to keep the dates of imported issues and comments, which is very useful to this study.
 
-This tool required a correspondence table between Bugzilla and GitHub accounts.
-Ours was created by manually matching 217 (out of 686) Bugzilla accounts to 175 GitHub accounts,\footnote{The correspondence table can be seen at \url{https://frama.link/vRV9etCd}.} the difference being due to the high prevalence of duplicate Bugzilla accounts.
-We gave priority to finding GitHub accounts for people who still had opened bug reports.
-%The most efficient way to do the matching was by searching the name of the reporter + ``github'' on Google. We also sent e-mails to people we hadn't managed to match (and who still had opened bugs).
-
-Once the switch was approved by the development team on October 4\textsuperscript{th}, 2017, the migration had to happen as soon as possible because every new pull request before the migration added to the number of bug reports that would need to be renumbered.
-It was conducted %\footnote{More details on the process are available in a blog post at \url{http://theoz.im/bugzilla}.
-}
-on October 18\textsuperscript{th},  the day after the 8.7.0 release (and not before to avoid disturbing the release process).
-Only 502 out of 4900 bugs (whose numbers were below 1154) had to be renumbered.
-This number is quite low because these bugs are dating back from the JitterBug era and during the previous switch only the bugs that were still open had been migrated.
-Due to the rate of pull request creation, if the switch had been delayed by a year, the number of renumbered bugs would have been closer to 3000.
+Once the switch was approved by the development team on Oct.
+4\textsuperscript{th}, 2017, the migration had to happen as soon as possible because every new pull request before the migration added to the number of issues that would need to be renumbered.
+It was conducted
+on Oct.
+18\textsuperscript{th},  the day after the 8.7.0 release (and not before to avoid disturbing the release process).
+Only 502 out of 4900 issues (whose numbers were below 1154) had to be renumbered.
+This number is quite low because a lot of the most ancient issues had been deleted (during the previous bug tracker migration).
+Due to the rate of pull request creation, if the switch had been delayed by a single year, the number of renumbered bugs would have been closer to 3000.
 
@@
 
-All the data for this study was extracted on January 17\textsuperscript{th}, 2019 using the GitHub GraphQL API \cite{github_graphql_API}.
-Using this API allows to do large requests (100 nodes in a single request) with just the information we need (thus both reducing the bandwidth usage and speeding up the extraction process).
-In the supplementary materials, we provide the extracted data as CSV files and a Jupyter notebook with the code to request this data from GitHub, to load the CSV files, to run the pre-processing steps, and the analyses.%, and the robustness checks.
+All the data for this study was extracted on March 31\textsuperscript{st}, 2019 using the GitHub GraphQL API~\cite{github_graphql_API}.
+Using this API allows us to do large requests (100 nodes in a single request) with just the information we need (thus both reducing the bandwidth usage and speeding up the extraction process).
+In the supplementary materials, we provide the extracted data as CSV files and a Jupyter notebook with the code to request this data from GitHub, to load the CSV files, to run the pre-processing steps and the analyses, including the robustness checks presented in Sec.~\ref{threats}.
 
-Migrated bugs and comments have their author information stored in the text, we extract it from there.
-Dates of bug and comment creation have been preserved which allows us to obtain them transparently for bugs before and after the migration.
-Bug reports and comments from the JitterBug era are missing information and were not fully migrated, therefore we do not consider any data from before 2008.
+As was previously mentioned, the migration conserved the dates of issues and comments, which allows us to obtain them transparently for issues before and after the switch.
+On the other hand, author information for migrated issues and comments had to be encoded in the text, and is thus extracted from there.
+Finally, we do not consider any data from before 2008, because this data comes from a previous migration, and was not properly saved.
 
@@
 
+\label{pre-processing}
+
 \paragraph{Excluding specific reporters}
-We have one specific reporter who is alone responsible for almost a quarter of all bug reports.
-To avoid having the behavior of a single individual strongly impact the overall statistics, we exclude his comments, bug reports and the comments they received from our analysis.
+We have one specific reporter who is alone responsible for almost a quarter of all issues.
+To avoid having the behavior of a single individual strongly impact the overall statistics, we exclude his comments, issues, and the comments they received from our analysis.
 Similarly, we also exclude the developer who was the main advocate of the bug tracker switch as it could have influenced his behavior before and after the switch.
 
-\paragraph{Merging duplicate Bugzilla accounts}
-We found that a significant number of users had created several accounts on the Bugzilla system.
-We merged them to avoid overestimating the rate of new reporters which is analyzed in the companion Jupyter notebook.
+\paragraph{Merging duplicate accounts}
+Because of space constraints, we refer the reader to the supplementary Jupyter notebook for details on this step.
 
 \paragraph{Removal of migration artifact comments}
-The migration tool created an issue whose body contains only meta-information, and a first comment with the description, for every bug report.
-These comments are thus migration artifacts and we remove them from our comment analyses.
-They can be easily identified because they are the comments that were posted at the exact same date and time as the corresponding bug report.
+The migration tool created artifact comments, which are easily identified because they are the comments that were posted at the exact same date and time as the corresponding issue.
+We exclude them from our analysis.
 
@@
 
-We measure different indicators of bug tracker activity: the numbers of bug reports per day; the number of distinct bug reporters in a week; the number of new bug reporters (who had never reported a bug before) per day; the number of comments per day; the number of distinct commentators in a week; and the number of new commentators per day.
-The number of distinct bug reporters and commentators is intended to allow us to distinguish between having a few prolific contributors and a large base of casual contributors.
-We use an interval of a week instead of a day for measuring the number of distinct bug reporters and commentators because, at the scale of a day, there is less opportunity for repeated contributions by the same contributor, but longer scales would compromise the estimation of effects by removing too much data.
+We measure different indicators of bug tracker activity: the numbers of issues per day; the number of distinct reporters in a week; the number of new reporters (who had never opened an issue before) per day; the number of comments per day; the number of distinct commentators in a week; and the number of new commentators per day.
+The number of distinct reporters and commentators is intended to allow us to distinguish between having a few prolific contributors and a large base of casual contributors.
+We use an interval of a week instead of a day for measuring the number of distinct issue authors and commentators because, at the scale of a day, there is less opportunity for repeated contributions by the same contributor, but longer scales would compromise the estimation of effects by removing too much data.
 
-For the visualization of Fig.~\ref{bug_nb_rd} to~\ref{commentator_nb_rd}, we plot a point for the average of each two week period.
-This is just to allow an easier and clearer visual analysis of the discontinuity.
-The regression lines, however, correspond to the definitions detailed in the above paragraph (i.e., analysis with daily periods or weekly periods depending on the variable).
+In the figures of Sec.~\ref{descriptive-stats}, each point represents an average on a four-week period.
+This is to reduce variability and allow an easier visual analysis.
+In the figures of Sec.~\ref{results}, each point represents an average on a two-week period.
 
-We analyze heterogeneous effects by distinguishing between developers and other contributors (others being assimilated to ``users'').
+We analyze heterogeneous effects by distinguishing between developers and non-developers.
 We define ``developers'' as the persons who have contributed more than 100 commits since 2008.
@@
 These developers are responsible for 91.5\% of all commits since 2008 (this is consistent with standard results on the proportion of commits by the ``core team'' in open source software~\cite{robles2009evolution}).
-
-For the dates of the new releases in Section~\ref{descriptive-stats}, we define release dates as the dates of the release announcement on the mailing list or of the corresponding news item on the official website (\url{https://coq.inria.fr/news}), whichever comes first.
-We selected beta and major releases and excluded release candidates and patch-level releases.
-The first release to appear in the news section of the current official website is the 8.2 beta release in June, 2008.
+Out of the 18 developers, 11 were active in the two months preceding the bug tracker switch (committed at least once).
+Once we exclude the two reporters mentioned in Sec.~\ref{pre-processing}, there remain 9 developers that were active at the time of the migration and that were included in our analysis.
+We later interviewed all of them.
+This seemingly arbitrary criterion happens to correctly recover what Coq developers viewed as the ``core team'' during the period around the bug tracker switch.
 
@@
 
-Figures~\ref{bug_nb_with_releases}, \ref{comments_with_releases}, \ref{reporters_with_releases} and \ref{commentators_with_releases} show the evolution of our four main outcome variables since January 2016 until March 2019.
+Figures~\ref{bug_nb_with_releases}, \ref{comments_with_releases}, \ref{reporters_with_releases}, \ref{commentators_with_releases} show the evolution of our four main outcomes from January 2016 to March 2019.
 We choose to present the data starting in 2016 because on January 2016 the 8.5 version of Coq was released, and this marks the adoption of a more rapid release cycle (with a new major release every 6 to 10 months)\footnote{See \url{https://coq.inria.fr/distrib/V8.9.0/refman/credits.html}.}.
@@
 
-Considering the whole period (2016-2019), less issues are reported by developers than by non-developers.
-In a four-day period, on average, two issues will be opened by developers and five by non-developers (Fig.~\ref{bug_nb_with_releases}).
-On the contrary, developers post more comments than non-developers.
-On an average day, developers will post around five comments and non-developers two (Fig.~\ref{comments_with_releases}).
-This is not surprising because some bugs found by developers are resolved directly without opening an issue and because issue reports are generally answered by developers.
-There are more distinct non-developers than developers who open issues in an average week (around six distinct non-developers and two distinct developers, Fig.~\ref{reporters_with_releases}).
-There are slightly more distinct non-developers (per week) that comment than developers.
-More precisely, on an average week, around six distinct non-developers and around 5.5 developers comment an issue (Fig.~\ref{commentators_with_releases}).
-
-%It is important to note that we are close to the theoretical saturation of the possible number of distinct developer commentators (already all distinct developers are commenting on weekly basis).
-For all outcomes, the mean before the switch is statistically significantly lower than the mean after the switch both for developers and non developers.
-For some outcomes, the difference in activity patterns before and after the switch is clear to the naked eye.
-In Fig.~\ref{bug_nb_with_releases} we see an increase in the number of issues reported by developers.
-In Fig.~\ref{comments_with_releases} we see an increase in the number of comments by developers with several data points after the release that are well above the highest point before the release.
-Finally, in Fig.~\ref{commentators_with_releases} we see a clear increase in the number of distinct non-developer commentators with the great majority of points after the switch above the great majority of points before the switch.
-Some other differences may be appreciated but are less clear to the naked eye than those presented.
-
-%The companion Jupyter notebook contains an extensive array of descriptive tables and figures, such as the distribution of bug reporting in the year and in the day, graphs of bug reporting for individual developers, etc.
-We only present a subset of these results.
-
-%\subsection{Cumulative number of bug reports and bug comments}
-
-%Fig.~\ref{cumulative_bugs} and~\ref{cumulative_comments} give a first (cumulative) view of two of our outcome variables: the number of bug reports, and the number of comments.
-From these figures, we can clearly notice the dominance of developers in the number of bug comments and the dominance of non-developers in the number of bug reports.
-We can also already notice some slope changes at the bug tracker switch date.
-The magnitude and significance of this changes will be assessed in Sec.
-\ref{results}.
-
-%\begin{figure}
-%\includegraphics{cumulative_bugs.eps}
-%\caption{Cumulative number of bug reports (since 2008).} \label{cumulative_bugs}
-%\end{figure}
-
-%\begin{figure}
-%\includegraphics{cumulative_comments.eps}
-%\caption{Cumulative number of comments (since 2008).} \label{cumulative_comments}
-%\end{figure}
-
-%\subsection{Influence of new releases}
-%\label{influence_release}
-
-%\paragraph{Influence on the number of bug reports}
-%We expect release dates to be correlated to peaks of activity on the bug tracker, in particular peaks of new bug reports.
-Indeed, users will generally test new releases, for instance by trying to port preexisting projects to the new version, which might lead them to find new bugs (either regressions or bugs related to new features).
-In Fig.~\ref{bug_nb_with_releases}, we can see that while release dates are generally correlated to peaks of new bug reports by non-developers, this is not so much the case of reports by developers.
-We can also notice that beta releases are often correlated to higher peaks than final releases which is not surprising given that the users who test beta versions are more likely to be ready to report bugs, and beta versions are likely to be less polished.
-
-%The highest peak is in 2015, after the first beta release of version 8.5.
-This was the first release in more than two years and it contained five big new features (``the result of five specific long-term projects''\footnote{Quote from:\\ \url{https://coq.inria.fr/distrib/V8.9.0/refman/credits.html\#credits-version-8-5}\\
-%The next quotes are coming from the following sections of the same chapter of the Coq reference manual.}).
-Stabilizing these features and their (unplanned) interactions required more than a year of testing and three beta releases.
-The 8.5 release was both hard on developers, who found the release cycle too long, and on users, who were impacted by many compatibility issues.
-Consequently, a different approach was taken for the following releases: versions 8.6, 8.7 and 8.8 were ``developed on a time-based development cycle'' and contain ``the result of refinements, stabilization of features and cleanups of the internals of the system along with a few new features''.
-Attention was given to regression testing: changes were systematically tested by building a selection of the largest Coq projects.
-Therefore, it is not surprising that these frequent releases are correlated to smaller peaks of new bug reports.
-
-%IMPORTANT: metre une phrase dans la partie description des variables qui indique qu'on fait 4 semaines pour celui la.
 \begin{figure}
@@
 \end{figure}
-
-%\paragraph{Influence on the number of comments}
-%In Fig.~\ref{comments_with_releases}, we can see that release dates are correlated to peaks of comments by developers (and not so much by other people).
-This is explained by the fact that bug reports are generally answered by developers, thus any peak in bug reporting is bound to induce a peak in bug commenting by developers.
 
@@
 
+Considering the whole period (2016-2019), less issues are reported by developers than by non-developers.
+In a four-day period, on average, two issues are opened by developers and five by non-developers.
+On the other hand, developers post more comments than non-developers.
+On an average day, developers post around five comments and non-developers two.
+This is not surprising because some bugs found by developers are resolved directly without opening an issue and because issue reports are generally answered by developers.
+There are more distinct non-developers than developers who open issues in an average week (around six distinct non-developers and two distinct developers).
+On average on the whole period, there are slightly more distinct non-developers (per week) that comment than developers.
+Interestingly, this is a case where the switch of bug tracker has changed the ranking: before the switch there were more distinct developers that commented each week, and afterwards, there were more distinct non-developers.
+
+For all outcomes, the mean before the switch is statistically significantly lower than the mean after the switch both for developers and non-developers (mean difference T-tests, $p < 0.001$).
+For some outcomes, the difference in activity patterns before and after the switch is clear to the naked eye.
+In Fig.~\ref{bug_nb_with_releases}, we see an increase in the number of issues reported by developers.
+In Fig.~\ref{comments_with_releases}, we see an increase in the number of comments by developers with many data points after the switch that are well above the highest point before the switch.
+Finally, in Fig.~\ref{commentators_with_releases}, we see a clear increase in the number of distinct non-developer commentators with almost all points after the switch above almost all points before the switch.
+Some other differences may be appreciated but are less clear to the naked eye than those presented.
+
 \section{Methodology}
@@
 \subsection{Quantitative}
-We exploit the fact that the switch from Bugzilla to GitHub can be seen as a clear cutoff in time to use a temporal sharp Regression in Discontinuity design (RDD) to estimate the effects of the switch on our outcome variables; to determine if these estimated effects are statistically significant; and to interpret these effects in a causal way.
-The idea behind RDD is that factors that influence bug reporting activity are fairly similar just before and just after the cut-off point, which makes the changes observed between just after and just before the cut-off a consequence of the change that occurred at the cut-off (i.e., the switch of bug reporting platform).
+We exploit the fact that the switch from Bugzilla to GitHub can be seen as a clear cutoff in time to use a temporal sharp regression discontinuity design (RDD) to estimate the effects of the switch on our outcome variables; to determine if the estimated effects are statistically significant; and to interpret them causally.
+The idea behind the RDD is that factors that influence bug tracking activity are fairly similar just before and just after the cutoff, which makes the differences observed between just before and just after the cutoff a consequence of the change that occurred at the cutoff (i.e., the switch of bug tracker).
 
-The assumption behind this analysis is that bug reporters did not choose the exact moment in which the switch would take place.
-In the absence of any other discontinuity, the changes in the behavior of bug reporters just before and just after the migration is likely to depend mainly on the switch: the other factors influencing bug reports evolve slowly, and should be pretty similar just before and just after the switch.
+The assumption behind this analysis is that reporters did not choose the exact moment in which the switch would take place.
+In the absence of any other discontinuity, the changes in the behavior of reporters just before and just after the migration is likely to depend mainly on the switch: the other factors influencing bug tracking activity evolve slowly, and should be pretty similar just before and just after the switch.
 It is important to note that estimated effects using this method will inform of the effect near the threshold (i.e., it will only give the Local Average Treatment Effect --LATE-- of the switch).
 Indeed, in the same way that behavior just before and just after the threshold is not likely to be different in the absence of the switch, the behavior several periods before and several periods after is likely to be different even in the absence of the switch.
-See~\cite{angrist2014mastering} and~\cite{angrist2008mostly} for an accessible and intuitive introduction to Regression on Discontinuity analysis; and~\cite{lee2010regression} and~\cite{jacob2012practical} for further details and a practitioner's guide to its empirical application.
+See~\cite{angrist2014mastering,angrist2008mostly} for an accessible and intuitive introduction to RDD; and~\cite{lee2010regression,jacob2012practical} for further details and a practitioner's guide to its empirical application.
 
-We model the evolution of the bug reporting behaviour around the switch by two different functions, one before and one after the switch, fitted using ordinary least squares.
+We model the evolution of the bug tracking activity around the switch by two different functions, one before and one after the switch, fitted using Ordinary Least Squares~\cite{wooldridge2015introductory}.
 Their purpose is to accurately estimate the value at the switch in the absence and presence of treatment.
-Since only their value around the cutoff point is relevant, we can choose to estimate them on a restricted bandwidth (number of periods) around the cutoff, or on a larger dataset.
+Since only their value around the cutoff is relevant, we can choose to estimate them on a restricted bandwidth (number of periods) around the cutoff, or on a larger dataset.
 We also have the choice of different functional forms for these models.
 
-In the context of a Regression in Discontinuity analysis, the choice of the bandwidth of the analysis and the functional form of the regression is always difficult.
-On the one hand, including data points that are far from the cut-off (in our case the switch) will increase precision as it will reduce standard errors (because we will have more data points).
-On the other hand, however, including those points can bias estimation near the cut-off if the functional form of the evolution in the absence of switch is not accurately specified.
-Our main specification takes a conservative approach with a relatively small bandwidth of 180 days before and after the switch to minimize possible bias, and a comparatively simpler linear model.
-However, as a robustness check, we also estimate a quadratic model on a larger time frame (455 days on each side).
+In the context of an RDD, the choice of the bandwidth of the analysis and the functional form of the regression is always difficult.
+On the one hand, including data points that are far from the cutoff (in our case the bug tracker switch) will increase precision as it will reduce standard errors (because we will have more data points).
+On the other hand, including those points can bias estimation near the cutoff if the functional form of the evolution in the absence of switch is not accurately specified.
+Our main specification takes a conservative approach with a relatively small bandwidth of 175 days before and after the switch to minimize possible bias, and a simple linear model to avoid overfitting.
+However, as a robustness check, we also estimate a quadratic model on a larger time frame (511 days on each side).
 This conservative approach reduces the statistical power of our analysis, increasing the probability of observing false negatives (being unable to detect an effect where such an effect exists).
 
-We use time as rating variable\footnote{The rating variable is also sometimes called the running variable or the forcing variable in the literature.} (with the cutoff centered at zero) and we estimate the following regression:
+We use time as rating variable\footnote{The rating variable, also called the running variable or the forcing variable, is the variable on which we observe the cutoff.} and we estimate the following regression:
 \begin{equation*}
@@
 \begin{array}{l}
-\mbox{\emph{Number of bug reports}}_t = \\
+\mbox{\emph{Number of issues}}_t = \\
 \quad\quad \gamma_0 +\gamma_1 \times \mbox{\emph{Relative date}}_t
@@
 \end{equation*}
-\noindent where $\mbox{\emph{Number of bug reports}}_t⁡$ is the total number of bugs reported during the day $t$; $\mbox{\emph{Relative date}}_t$ is the number of days from the date of the switch (zero is the first period after the switch); $\mbox{\emph{After switch}}_t$ is a binary variable equal to one if $t$ is a period after the switch and zero otherwise; $\mbox{\emph{Relative date}}_t \times \mbox{\emph{After switch}}_t$ is called the interaction term and $\epsilon_t$ is the residual error.
+\noindent where $\mbox{\emph{Number of issues}}_t$ is the total number of issues opened during the day $t$; $\mbox{\emph{Relative date}}_t$ is the number of days from the date of the switch (zero is the first period after the switch); $\mbox{\emph{After switch}}_t$ is a binary variable equal to one if $t$ is a period after the switch and zero otherwise; $\mbox{\emph{Relative date}}_t \times \mbox{\emph{After switch}}_t$ is called the interaction term and $\epsilon_t$ is the residual error.
 This is equivalent to estimating the following two regressions, respectively before and after the switch:
@@
 \begin{array}{l}
-\mbox{\emph{Number of bug reports}}_{t < 0} ⁡=\\
+\mbox{\emph{Number of issues}}_{t < 0} =\\
 \quad\quad \gamma_0 + \gamma_1 \times \mbox{\emph{Relative date}}_t + \epsilon_p \medskip\\
-\mbox{\emph{Number of bug reports}}_{t \geq 0} = \\
+\mbox{\emph{Number of issues}}_{t \geq 0} = \\
 \quad\quad (\gamma_0 + \gamma_2) + (\gamma_1 + \gamma_3) \times \mbox{\emph{Relative date}}_t + \epsilon_t
@@
 } Coefficient $\gamma_0$ is the estimated value just before the cutoff and $\gamma_1$ the slope before the cutoff.
-Coefficients $\gamma_2$ and $\gamma_3$ are the estimates of interest and will tell us the jump of the number of bugs reports just after the switch and the change in slopes due to the switch respectively.
+Coefficients $\gamma_2$ and $\gamma_3$ are the estimates of interest and will tell us the jump of the number of issues just after the switch and the change in slope due to the switch respectively.
 
@@
 
-In order to obtain qualitative insights into the effects of the switch that will help us put in context and interpret the quantitative results, we conducted semi-structured interviews of all developers of Coq that were active at the time of the switch (with the exception of the developer that conducted the migration).
-A total of 9 interviews were conducted between 19th March and 4th April 2019.
-Some were done face to face and the others by video-conference with an average length of 33.5 minutes per interview.
-Both authors were present during the interviews and all the interviews were recorded.
-Some basic questions were asked to all of the interviewees and according to the answers some additional questions were also asked.
+In order to obtain qualitative insights into the effects of the switch that will help us interpret the quantitative results, we conducted semi-structured interviews of all developers of Coq that were active at the time of the switch (as defined in section \ref{data}) and that we included in our quantitative analysis.
+A total of 9 interviews were conducted between Mar., 19\textsuperscript{th} and Apr., 4\textsuperscript{th} 2019.
+The two co-authors were present during the interviews and all the interviews were recorded.
 The answers were coded by both authors in an attempt to minimize errors.
-Code and questions can be found in the supplementary material.
+The interview outline with questions asked to all developers, and the encoding of the answers can be found in the supplementary material.
 
-\section{Results}
+\section{Quantitative results}
 \label{results}
-\subsection{Impact on the number of bug reports}
+\subsection{Impact on the number of issues}
 
@@
 \centering
-\caption{Estimated impact of the switch on the number of issues.
-Coefficients are highlighted if they are statistically significant with $p<0.1$ ($\dagger$), $p<0.05$ (*), $p<0.01$ (**) or $p<0.001$ (***).
+\caption{Estimated impact on the number of issues.
+Statistically significant results are in boldface (\textbf{*} means $p<0.05$, \textbf{**} means $p<0.01$, \textbf{***} means $p<0.001$).
 Standard error is in parentheses.}
@@
 \includegraphics{bug_nb_rd.png}
-\caption{Number of issues per day before and after the migration (with fitting lines and confidence intervals from the regression results, and points corresponding to average values over two-week periods).} \label{bug_nb_rd}
+\caption{Number of issues per day before and after the switch (with fitting lines and confidence intervals from the regression results, and points corresponding to average values over two-week periods).} \label{bug_nb_rd}
 \end{figure}
 
-Table~\ref{tab:bug_nb} presents the estimated impact of the switch on the number of reported bugs.
-Each column shows the estimates for a different sub-sample (i.e., all bug reporters, developers and non-developers).
-Asterisks represent different classical levels of statistic significance.
-We interpret results as significant if the p-value is below 0.1.
-Estimates that are not statistically significant cannot be interpreted as an absence of effect but indicate that if such effect occurred, we are unable to discern it with our conservative approach.
-Fig.~\ref{bug_nb_rd}  shows the number of reported bugs before and after the switch and the fitting lines and confidence intervals corresponding to the regression results.
+Table~\ref{tab:bug_nb} presents the estimated impact of the switch on the number of issues.
+Each column shows the estimates for a different sub-sample (all reporters, developers and non-developers).
+Asterisks represent levels of statistic significance.
+We interpret results as significant if the p-value is below 0.05.
+Estimates that are not statistically significant cannot be interpreted as an absence of effect but indicate that if such effect exists, we are unable to discern it with our conservative approach.
+Fig.~\ref{bug_nb_rd}  shows the number of issues before and after the switch and the fitting lines and confidence intervals corresponding to the regression results.
 
-For developers, we see a statistically significant positive jump in bug reports just after the switch.
-In particular changing the bug reporting platform is estimated to have increased the daily number of reported bugs by 0.7 (representing an increase of around 250\%).
-That is, on average, we observe around one bug reported by developers every day after the switch, %(but close to the switch)
-while before the switch developers reported a bug every three days.
-There is no statistically significant effect on the number of bug reports by other contributors: we cannot reject that such a variation occurred, but we are unable to discern it with our conservative approach.
+For developers, we see a statistically significant positive jump in issues just after the switch.
+The switch is estimated to have increased the daily number of issues by 0.7 (representing an increase of around 270\%).
+That is, on average, we observe around one issue opened by developers every day after the switch,
+while, before the switch, developers opened an issue every four days. There is no statistically significant effect on the number of issues by non-developers.
 
-\subsection{Impact on the number of distinct bug reporters}
+\subsection{Impact on the number of comments}
 
-Table~\ref{tab:reporters} and Fig.~\ref{reporter_nb_rd} show the estimated impact of the switch on the number of distinct bug reporters each week.
-These results demonstrate that the switch to GitHub positively affected not just the number of bug reports but also the number of distinct developer bug reporters in a given week.
-The number of distinct developers that reported bugs in a given week increased by around 130\% (from a 1.5 reporting developers each week before the switch to 3.42 after)
+Table~\ref{tab:comment_nb} and Fig.~\ref{comment_nb_rd} show the estimated impact of the bug tracker switch on the number of comments.
+We see a statistically significant positive jump.
+The number of comments just after the switch more than doubles (from around 5 comments per day before the switch to around 10 after).
+This appears to be, for the most part, due to comments by developers who increased their average number of comments per day from around 3 to around 8.
 
@@
 \centering
-\caption{Estimated impact of the switch on the number of distinct reporters each week.
-Coefficients are highlighted if they are statistically significant with $p<0.1$ ($\dagger$), $p<0.05$ (*), $p<0.01$ (**) or $p<0.001$ (***).
+\caption{Estimated impact on the number of comments.
+Statistically significant results are in boldface (\textbf{*} means $p<0.05$, \textbf{**} means $p<0.01$, \textbf{***} means $p<0.001$).
+Standard error is in parentheses.}
+\label{tab:comment_nb}
+\input{comment_nb_rd.tex}
+\end{table}
+
+\begin{figure}
+\includegraphics{comment_nb_rd.png}
+\caption{Number of comments per day before and after the switch (with fitting lines and confidence intervals from the regression results, and points corresponding to average values over two-week periods).} \label{comment_nb_rd}
+\end{figure}
+
+\subsection{Impact on the number of distinct reporters}
+
+Table~\ref{tab:reporters} and Fig.~\ref{reporter_nb_rd} show the estimated impact of the switch on the number of distinct issue reporters each week.
+These results show that the switch had also a positive effect on the number of distinct developer-reporters in a given week.
+The number of distinct developers that opened an issue in a given week increased, on average, by around 130\% (from 1.5 developer-reporters each week before the switch to 3.42 after).
+
+\begin{table}
+\centering
+\caption{Estimated impact on the number of weekly distinct reporters.
+Statistically significant results are in boldface (\textbf{*} means $p<0.05$, \textbf{**} means $p<0.01$, \textbf{***} means $p<0.001$).
 Standard error is in parentheses.}
@@
 \includegraphics{reporter_nb_rd.png}
-\caption{Number of weekly distinct reporters before and after the migration (with fitting lines and confidence intervals from the regression results, and points corresponding to average values over two-week periods).} \label{reporter_nb_rd}
-\end{figure}
-
-\subsection{Impact on the number of comments}
-
-Table~\ref{tab:comment_nb} and Fig.~\ref{comment_nb_rd} show the estimated impact of the bug tracker switch on the number of comments.
-We see a statistically significant jump in the number of comments, which more than doubles (from 4.8 comments per day before the switch to 10.3 after the switch).
-This appears to be, for the most part, due to comments by developers who increased their average number of comments per day from around 3 before the switch to around 8 after.
-
-\begin{table}
-\centering
-\caption{Estimated impact of the switch on the number of comments.
-Coefficients are highlighted if they are statistically significant with $p<0.1$ ($\dagger$), $p<0.05$ (*), $p<0.01$ (**) or $p<0.001$ (***).
-Standard error is in parentheses.}
-\label{tab:comment_nb}
-\input{comment_nb_rd.tex}
-\end{table}
-
-\begin{figure}
-\includegraphics{comment_nb_rd.png}
-\caption{Number of comments per day before and after the migration (with fitting lines and confidence intervals from the regression results, and points corresponding to average values over two-week periods).} \label{comment_nb_rd}
+\caption{Number of weekly distinct reporters before and after the switch (with fitting lines and confidence intervals from the regression results, and points corresponding to average values over two-week periods).} \label{reporter_nb_rd}
 \end{figure}
@@
 
-Table~\ref{tab:commentators} and Fig.~\ref{commentator_nb_rd} show the estimated impact of the bug tracker switch on the number of distinct commentators each week.
-We can see a statistically significant jump in the total number of commentators, in the number of developer commentators, and in the number of non-developer commentators.
+Table~\ref{tab:commentators} and Fig.~\ref{commentator_nb_rd} show the estimated impact of the switch on the number of distinct commentators each week.
+We observe a statistically significant jump in the number of both developer and non-developer commentators.
 
 In general terms, the number of distinct commentators in a given week changed from an average of 7 to an average of 13 (almost doubling the commentators).
-Analyzing the developers and the non-developers, we can see that this increase is due in larger part to non-developers (whose number of commentators more than doubles) while, among developers, this number increases by 66\%.
+This increase is due in larger part to the number of non-developers commentators, which more than doubles (from around 3 to around 7 on an average week), while, among developers, the number of commentators  increases by 66\% (from around 4 to around 6).
 
@@
 \centering
-\caption{Estimated impact of the switch on the number of distinct commentators each week.
-Coefficients are highlighted if they are statistically significant with $p<0.1$ ($\dagger$), $p<0.05$ (*), $p<0.01$ (**) or $p<0.001$ (***).
+\caption{Estimated impact on the number of weekly distinct commentators.
+Statistically significant results are in boldface (\textbf{*} means $p<0.05$, \textbf{**} means $p<0.01$, \textbf{***} means $p<0.001$).
 Standard error is in parentheses.}
@@
 
-\section{Discussion}
+\subsection{Impact on the number of new reporters and commentators}
+
+We refer the reader to the companion Jupyter notebook for the results on new reporters and new commentators.
+The estimated coefficients also show a positive jump, but they are not statistically significant.
+
+\section{Research questions discussion}
 \label{discussion}
 
-\subsection{H1: impact on the number of reported bugs}
+\subsection{RQ1: Impact on level of bug tracking activity}
 
-We hypothesized that the switch to a platform that is perceived as more modern and easier to use would increase the number of reported bugs.
-Our data analysis only shows such an effect on the number of bug reported by developers, and not on the bugs reported by non-developers, which only partially validates our first hypothesis.
+In Sec.~\ref{descriptive-stats}, we observed an increase in the mean number of issues and comments after the switch, both for developers and non-developers.
+However, as shown in Sec.~\ref{results}, we are able to find a causal effect of the switch on these two variables only for developers.
+In particular, the estimated impact of the switch was to more than quadruple the number of opened issues, and to more than double the number of comments, by developers.
+This significant increase in the bug tracking activity is all the more remarkable given that two (out of nine) developers highlighted that it was hard to adapt to the new bug tracker, and thus they reduced their activity after the switch.
 
-The developers were using the bug tracker very little before the switch.
-This means that when they found bugs in the course of the development, they stored them in personal task lists, fixed them immediately, or tried to remember them in their heads until they could fix them (likely with some bugs being forgotten in the process).
-Meanwhile, they were discussing a lot through the pull request system on GitHub.
-The move of the bug tracker to GitHub made it closer to the main development platform (the repository where pull requests are submitted and reviewed), both in terms of location and in terms of user interface.
-These reasons are probably what pushed developers to use it more.
+We not only observed an increase in activity by developers but also an increase in the number of distinct developer reporters and commentators. This shows that developers are active more often.
 
-The increased use of the bug tracker in place of personal task lists makes the whole development process more transparent and is likely to be beneficial to new contributors.
+When we asked developers how they would interpret this effect on the level of activity of developers, they almost all (7 out of 9) mentioned the more comfortable experience of using GitHub's integrated bug tracking system, reminding many of the motivations they had previously listed for switching (better interface, less slow, no switching between issues and pull requests, etc.).
+Many developers told us that they personally did not always opened issues on the Bugzilla bug tracker when they found small bugs, and they now did open issues more often because it was easier to do so.
+One developer in particular admitted he had not been very active on the Bugzilla bug tracker since the main development activity had moved to GitHub's pull requests.
 
-If it is the case that the switch had no effect on the number of bugs reported by non-developers, this could mean that when people decide to report bugs, they do so before accessing the bug tracker, and nor the requirement of creating an account, nor the heavier and less modern interface, are sufficient to discourage them.
+In addition, the GitHub notification system (including the mention system which allows attracting a developer's attention on a specific issue) was identified as being responsible for part of the additional activity.
+When a user is subscribed to a repository (which is the case of most Coq developers), notifications (by e-mail or in the web interface) are sent whenever a new issue is opened or a comment is posted.
+This can attract the attention of developers more quickly and lead to more comments in response.
+It was mentioned how some developers had trouble adjusting their settings to make it easier to manage this large flow of information (GitHub's notification default settings are quite unsuited to repositories with large levels of activity).
 
-%The second one is that more people intended to report bugs but this increase was blurred out by a simultaneous decrease in the number of duplicate bug reports, thanks to the better search provided by GitHub.
-People intending to report a bug and finding out that it was already reported would post a comment (although if this was frequently the case, there should be an increase in the number of comments by non-developers, which hasn't been observed in a significant way neither) or add a reaction (GitHub feature allowing to up-vote a comment) or leave without taking any action instead.
-To confirm this second interpretation, we would need to analyze the impact of the switch on the number of bugs closed as duplicates.
+\subsection{RQ2: Impact on quality of bug tracking activity}
 
-\subsection{H2: impact on the pool of reporters and commentators}
+Measuring quality of bug tracking activity is difficult.
+That is why we could only answer this second research question qualitatively, by asking developers if their perceived a change in quality of issue reports and interactions on the bug tracker.
+A majority (5 out of 9) claimed that GitHub helped at least some reporters produce better issue reports (thanks in particular to formatting and the issue template\footnote{GitHub allows setting up a template to help users opening issues know what information they should provide~\cite{github_issue_template}.}).
+However, three developers added that the switch had also attracted less experienced users whose reports were sometimes of a lower quality.
 
-% how many of the reporters of recent bugs had a GitHub account?
+Regarding the quality of interactions on the bug tracker, again a majority of developers (6 out of 9) claimed that it improved after the switch.
+In particular, it was mentioned that users are more reactive and answer more frequently when they are asked additional information.
+Several developers also emphasized that it was easier to discuss on the new platform, especially with respect to bugs that cross project boundaries.
+Indeed, GitHub makes it easy to link between issues, not only in the same repository, but also across the repositories of multiple projects.
+Examples of related projects include projects that Coq depends on (the OCaml compiler, the camlp5 pre-processing library, the Sphinx documentation tool, etc.) and projects that depend on Coq (a variety of Coq plugins and libraries, editor support packages, etc.).
 
-Having a common platform for the pull requests and the bug reports, and using a standard solution where a lot of users would already have an account, was hypothesized to have an impact on the pool of participants in the bug tracker.
-For non-developers, our data analysis only shows this effect in a statistically significant way on the number of commentators, which only partially validates our second hypothesis.
+Other perceived quality improvements include an easier navigation among issues, which helps both developers do a more efficient triaging, and users discover existing issues more easily, on which they can subsequently leave comments adding information.
+This participation of users is useful to improve the overall quality of some preexisting reports.
 
-Analyzing deeper, we find that there exists not just a jump in the non-developer commentators in a given week but also a jump in the number of new non-developer commentators (i.e., people that had never before posted a comment).
-We show the results for the new commentators in the companion Jupyter notebook.
+\subsection{RQ3: Impact on the audience of the bug tracker}
 
-While we suggested that reporters would be motivated enough to not be affected by the bug tracking platform, this is less likely to be the case for people willing to comment.
-Furthermore, people who were already following the pull request activity were now subscribed to the bug tracking activity as well and this could have been enough to encourage them to post comments.
-The previous system made it harder and less flexible to subscribe to the bug tracker activity.
+In Sec.~\ref{descriptive-stats}, the clearest change after the bug tracker switch was the increase in the number of distinct weekly non-developer commentators.
+This observation was further confirmed in Sec.~\ref{results}: the estimated impact of the switch was to more than double the number of distinct non-developer commentators in an average week.
 
-This analysis is consistent with previous findings: many people start by \emph{lurking} (i.e., mostly reading discussions) before starting to contribute to an open source project (for instance by participating to discussions)~\cite{von2003community}.
-In the context of GitHub, a significant portion of the new contributors start by \emph{watching} the repository (i.e., subscribing to its activity).
-These contributors contribute for longer periods and achieve a wider variety of contribution types, probably due to the knowledge and confidence they gained during the watching process~\cite{sheoran2014understanding}.
+When asked to interpret
+this surge,
+all the developers said that GitHub was more accessible, either because it is a very popular platform on which many people already have an account (especially computer science students who represent a large number of Coq users), or because it is quite simple to use, especially in comparison with Bugzilla.
+These explanations apply to a general increased participation of non-developers, but do not explain specifically an increase in the number of distinct commentators (as opposed to reporters).
 
-% Commenté pour faire plus court, et ça n'est pas suffisamment essentiel ici.
-%These results are particularly interesting because they highlight an increase in the number of people involved in the development process in the large sense.
-They also show a remarkable change in the way non-developers engage in the software developing process following the switch.
+Some factors that were mentioned by a few developers and could have specifically influenced the number of commentators are that some external contributors who already contributed through pull requests now also comment issues more systematically, that more attention is given to newcomers, that there is generally more discussion between developers and non-developers on the bug tracker, and that it is easier for interested users to subscribe to just part of the bug tracking activity (and thus participate to the discussion without following the activity on the whole bug tracker).
 
-The significant impact on distinct reporters and commentators among developers is more surprising but shows that some developers were disengaged from any kind of activity on the bug tracker before the switch.
-One developer explained this as a difficulty to keep track of activity on two platforms at once (GitHub for the pull requests and Bugzilla for the bug reports).
-Comparing data from pull request commentators and bug report commentators confirmed this: before the switch there were more developers commenting pull requests than bug reports in any given week, whereas after the switch this difference disappeared completely (graph available in the accompanying notebook).
+\subsection{General discussion}
 
-%todo for future... are there more non-developers closer to become developers ?
+Beyond the scope of our research questions, we gathered other perceived effects of the bug tracking switch that affect the quality of software development in a broader way.
+Most of the developers we interviewed noted that the bug tracker switch to GitHub contributed to a more general culture switch towards a more recorded communication and transparent development.
+In particular, it was mentioned that GitHub makes written communication less costly, that the mention system can be used in place of personal e-mails, which avoids losing technical content in private threads, and that more types of discussion take place in the bug tracker (included related to long-term plans or development processes).
 
-\subsection{H3: impact on the interaction between users and developers}
+In addition, some developers found that GitHub provides a more global view of the state and history of development.
+They also perceived an increased visibility of developers' own work, and of the Coq project overall.
 
-The surge in the number of comments posted by developers shows that developers communicate more on the bug reports they receive, instead of just fixing and closing them.
-This result suggests once again that the switch was beneficial in terms of transparency of the development, an evolution that is likely to give greater satisfaction to the bug reporters and probably encourage engagement in the process of bug reporting and development more generally.
-Taken together with the increased number of non-developer commentators, this validates our third hypothesis.
+While our RDD methodology can only be used to measure the immediate causal impact, an important question concerns the long-term effect of the switch.
+Our descriptive statistics (Sec.~\ref{descriptive-stats}) and robustness checks on longer time horizons (Sec.~\ref{threats}) seem to validate that the observed effect of the switch was durable.
+This can be explained in part because GitHub evolves faster, according to the needs of ever-evolving open source developing processes.
+During our interviews, we learned that some of the recently added features of GitHub are particularly appreciated.
 
-% Commenté pour faire plus court, et ça n'est pas suffisamment essentiel ici.
-Overall, these results suggest that the change of platform reinforced discussion among developers and among developers and non-developers, which may improve efficiency in the developing process, incentivize non-developers to have a more active role (maybe even assuming developing tasks) and in the long run increase the quality of the software (as bugs are probably better understood and fixed in better accordance to users' needs).
-
-\section{Threats to validity}
+\section{Robustness checks and threats to validity}
 \label{threats}
@@
 
-The biggest concern that we have in our analysis is the fact that the date of the switch is close to the release date of a new version of Coq.
-The release of version 8.7.0 took place on October 17\textsuperscript{th}, and the migration on October 18\textsuperscript{th}.
-Then, the hypothesis that there is no other discontinuity around the threshold of interest does not completely hold in our case.
-%First, we can argue that only our result on the number of comments by developers is threatened: the other statistically significant effects that we observed are not threatened because there is usually no peak in the corresponding variables after new releases.
-%Second, we can also argue that there should have been a similar (or higher) peak in the month before the switch since the first beta of 8.7 had been released then (and beta releases generally result in higher peaks).
-To confirm that it is not the migration that is driving our results, we do the exact same analysis that we do for the bug tracker switch, replacing the cutoff date by the date of all six major releases since 2009 (once we remove the latest release because we do not have yet enough data points after the release).
-We show that the only significant effect in the bug-tracker activity after a major release is the effect after the major release that match the switch.
-It is also, for the majority of the outcomes, by far the effect with the highest magnitude.
+The most important threat when performing RDDs, especially on time-series, is to have another event that occurred near the cutoff, and which could have influenced the outcomes.
+In our case, there were no such events that we could identify, except the 8.7.0 release, that took place on Oct.
+17\textsuperscript{th}, just one day before the switch.
+To confirm that it is not the release that is driving our results, we perform some placebo robustness checks, i.e.
+we do the exact same analysis that we do for the bug tracker switch, replacing the cutoff date by the date of six other releases (all the other major releases since 2009, except the latest one because we do not have yet enough data points after the release).
+We show that the only positive and significant effect in the bug tracker activity after any proposed cutoff is the effect after the switch.
+For some outcomes, we observe a significant effect after some major releases but this effect is negative and of a much lower absolute magnitude.
+It is thus very unlikely that our results are driven by the release close to the switch and not by the switch itself.
 
-%To provide some evidence that the date of the switch is not associated with an increase in the activity of open source software due to external factors (which will invalidate our results as explained before), we replicate our analysis for a comparable software: another French medium size open-source project: OCaml.
-We do not observe any increase in the bug-tracking activity for this project after the Coq bug-tracker switch date.
-If anything we see a decrease in activity so if there was an external factor that affected french open-source projects decreasing their activity.
-we will be underestimating the effect of the switch.
-
-%two robustness checks with our data.
-The first consists in removing the two weeks following the 8.7.0 release (and the switch) from our Regressions in Discontinuity.
-With these periods removed, the regressions still give very close results.
-The second consists in including a binary control variable that is equal to one if the period is in the two weeks after a release and zero otherwise.
-This control variable will capture the high peaks due to the release without the need of eliminating periods close to the switch.
-Again, our results remain very similar when we include this control variable.
-
-Our main specification for Regression on Discontinuity uses a small bandwidth and a linear model.
-In this section, we also present the results using all the available data (455 days on each side of the cutoff, corresponding to all the data that we collected after the switch and the same number of days before).
-To reduce the associated bias, we include a quadratic term to allow the regression to better fit the data and "adjust" to far away data points.
+Our main specification for RDD uses a small bandwidth and a linear model.
+As a robustness check, we estimate our model using larger time horizons (511 days, i.e.
+73 weeks, on each side of the cutoff, corresponding to all the data that we collected after the switch and the same number of days before).
+To reduce the associated bias, we include a quadratic term to allow the regression to better fit the data and ``adjust'' to far away data points.
 Following Gelman et al.
 ~\cite{gelman2018high}, we include only a quadratic polynomial and not a polynomial of a higher order.
-Our results still hold with this specification.
+Virtually all our results still hold with this specification, with comparable estimated magnitude, and with statistical significance (with only the number of distinct developer-reporters going just above the arbitrary $0.05$ p-value threshold).
 
-Temporal RDD have some notable differences with other types of RDD \cite{hausman2018regression}.
-One major difference is the capacity of anticipation and reaction to the treatment that affect directly the outcome.
-To overcome this potential problem, we redo our analysis with a donut RD (i.e.
-we exclude the data points just after and just before the migration, which are potentially problematic), following \cite{hausman2018regression}.
-Our results still hold with a slight decrease in the level of significativity but keeping the significativity at 10\%.
+Temporal RDDs have some notable differences from other types of RDDs, as noted in a recent paper by Hausman and Rapson \cite{hausman2018regression}.
+One major difference is the capacity of anticipation and adaptation to the treatment, that affects directly the outcome.
+To ensure that our results are robust to these effects, we additionally estimate donut RDDs (i.e.
+we exclude the data points just after and just before the switch, which are potentially problematic), as recommended in \cite{hausman2018regression}.
+Our results still hold with comparable estimated magnitudes, and a few of them going just above the arbitrary $0.05$ p-value threshold (this is not unexpected, given that we have removed data points, and this automatically increases the standard errors).
+Time-series are also subject to potential serial auto-correlation of errors (errors at time $t+1$ are not independent from errors at time $t$).
+Failing to take such auto-correlation into account can lead to underestimate standard errors.
+To check that there is no serial auto-correlation in our data, we plot residuals\footnote{Residuals are the differences between measured and estimated values.}~\cite{wagner2002segmented}, and notice the absence of any specific pattern beyond those which show the presence of heteroscedasticity, which we already control for (see Jupyter notebook).
 
-%In order to corroborate that our results are not driven by activity occurring much after or much before the switching period (that is more likely to be affected by other factor beyond the switch), we check if our results are robust to the exclusion of the 1\%, 5\% and 10\% external values (i.e., values that are furthest from the switch) ~\cite{jacob2012practical}.
-
-%In the context of a Regression in Discontinuity analysis, the choice of the bandwidth (number of periods around the switch) of the analysis and the functional form of the regression is always difficult.
-By the one hand, including data points that are far from the cut-off (in our case the switch) will increase precision as it will reduce standard errors (because we will have more data points).
-By the other hand, however, including those points can bias estimation near the cut-off.
-In our preferred specification that we present in the results section, we take a conservative approach and take a relatively small bandwidth to minimize possible bias.
-In this section, we also present the results using all the disposable data.
-To reduce the associated bias, we include a quadratic term to allow the regression to better fit the data and "adjust" for far away data points.
-Following Gelman et al.
-(2018) ~\cite{gelman2018high}, we include only a quadratic polynomial and not a polynomial of a higher order.
-
-%We also conduct one standard robustness check for Regression in Discontinuity analysis by changing the time window of the analysis to fifteen and to ten periods before and after the switch.
-Our main results are robust to these changes.
-All results presented in this section can be found in the Jupyter notebook provided as a supplemental material. 
 \subsection{External validity}
 
-This is only a case study on a very specific project (by the target audience, the location and composition of its core team, etc.).
-We would really benefit from conducting similar analyses on other projects that went through a similar bug tracker switch.
-
-We located another project that went through a similar switch from Bugzilla to GitHub and used the Bugzilla import tool with our modifications.
-It would have been a perfect case for replicating this analysis.
-A preliminary visualization of the data shows that, compared to the few years before the switch, there have been more bug reports and comments posted after the switch.
-Unfortunately, we learned that some data was lost during the crash of their previous bug tracker, which makes it really difficult to draw any reliable conclusion.
-%after talking with the developer responsible for the migration, we learned that the switch was the consequence of a crash of the Bugzilla instance that could not be fixed, and that some data (bugs and comments) were lost in the crash.
-Some of it could be recovered and was manually added, which makes it inaccurately appear as activity after the switch.
-
-We nonetheless believe that our conclusions can be generalized to some extent and encourage researchers to replicate this study, for instance using our migration tool and/or our analysis pipeline and methodology.
-
-There is no reason to believe that these results would generalize to proprietary software projects.
-Some projects that are also less likely to benefit as much from a switch to GitHub are the larger projects (like Mozilla, Linux, Gnome, KDE, Apache) which have been using Bugzilla for a long time and have defined rigorous development processes, in particular regarding the use of the bug tracker.
-Coq is an example of a project that is developed in a very Bazaar-like fashion, with no clear rules on how to use the bug tracker, and this lack of rules was hindering the use of a complex bug tracker like Bugzilla.
-We believe that open source projects that share this lack of rules and this medium size are the more likely to benefit from a switch from Bugzilla (or similar bug trackers like Mantis), to GitHub (or the very similar GitLab platform), especially if they are already heavily using the latter for pull requests.
+This is only a case study on a very specific project with a relatively small core team, and therefore it is very important that this study be replicated before we can draw general conclusions that can serve to formulate recommendations for open source developers.
+Our reusable assets, such as the migration tool and the analysis pipeline should be helpful to perform such replications.
 
@@
 \label{conclusion}
-We have analyzed the impact of changing the bug reporting environment using a Regression in Discontinuity analysis with data from the switch of the Coq bug tracker from Bugzilla to GitHub.
-We have shown that the switch incurred an increase in bug tracking activity in two ways.
-First, the developers are using the bug tracker more, to report their own bugs, and to discuss bug reports.
-Second, the non-developers are more numerous to take an active role by commenting bug reports as well.
-We find particularly interesting the change in the number of comments by developers and in the number of distinct non-developer commentators because it shows a shift in the dynamics of the bug reporting process that we believe can improve software development by engaging more people in the process, and by increasing openness and transparency.
+We have conducted the migration of the Coq bug tracker from Bugzilla to GitHub and we have improved existing tooling in the process.
+Other open source projects have already benefited from this enhanced tooling (Ledger, migrating from Bugzilla to GitHub in 2018), or the experience that we drew from it (OCaml, migrating from Mantis to GitHub in 2019).
+
+By analyzing mined repository data, we have shown that this bug tracker switch resulted in an increase in bug tracking activity in two ways.
+First, the developers are using the bug tracker more, by opening more issues and communicating more.
+Second, the non-developers are more numerous to take an active role by commenting issues as well.
+
+We complemented this quantitative analysis by interviewing the developers.
+Almost all of them were very happy with the bug tracker switch.
+They provided us with insights to explain our estimated causal effects in bug tracking activity, and on the more general impact on the development process and its transparency.
+
+Finally, we presented and explained an example of the use of RDD, an econometric method to derive causality, that is seldom used in empirical software engineering.
+We believe that there are many cases where this methodology would be useful to help demonstrate causality from mined software repository data.
 
 Given the lack of literature on the effects of bug tracking environments, interested researchers could expand our analysis in many different ways.
-Analyzing other outcomes (such as pull requests or the rapidity of bug resolution) could give interesting additional insights and help create a more complete picture of the mechanisms at play.
-% je pense qu'on peut se permettre d'enlever cette phrase maintenant
-%It will be also interesting to analyze the effects on a longer period to see if the changes hold in the long run.
-Replicating our analysis for different contexts and software would increase the external validity of the results.
-
-Finally, we introduced and showed an example of the use of a Regression on Discontinuity design, that is novel in the context of empirical software engineering.
-We think that there are many cases where this methodology would be useful to help demonstrate causality from mined software repository data.
-
-% Takeaways
-% When to switch?
-It would be useful to come up with measures that indicate when a bug tracker is not adapted and a switch to another platform would be an improvement.
-(Or depending on specific objectives.)
+Analyzing other outcomes (such as pull requests or the rapidity of bug resolution) could give interesting additional insights, and help create a more complete picture of the mechanisms at play.
+Replicating our analysis on more projects would increase the external validity of the results, and allow researchers to offer clear takeaways to projects interested in switching their bug trackers.
+Based on this first study, it seems that moving to a more popular, and integrated platform can help increase the level, the quality and the audience of bug tracking activity.
+\\
 
 \scriptsize
-%\paragraph{}
 \noindent \footnotesize{\textbf{Supporting data.}}
-We provide our extraction and analysis code as a Jupyter notebook and the data we used as CSV files.
-We also provide the XML export of Bugzilla that was used for the migration and the migration script.
-% XXXXX We also provide the interview questions and coding.
-%Everything is available from \url{XXXXXXX}.
-%\paragraph{}
+Supplementary materials can be found at \url{https://github.com/Zimmi48/impact-of-switching-bug-trackers/tree/icsme-final}.
+They include our extraction and analysis code in a Jupyter notebook, the bug tracking data gathered from GitHub, the interview questions and coding of the answers.
 
-%\noindent \footnotesize{\textbf{Authors' contributions.}}
-%[masked for double-blind review]
-%TZ found the research question, prepared and conducted the migration, designed and implemented the data extraction process, implemented the data analyses, and wrote part of the manuscript.
-AC designed the data processing and analyses, guided the implementation of the data analyses, presented the econometric method in use and wrote part of the manuscript.
-%\paragraph{}
-
-%\noindent\footnotesize{\textbf{Acknowledgements.}}
-%We would like to thank Ambre Williams for valuable advice and help on the design and implementation of the data analyses and C\'ecile Rottner for general advice, support, and proof-reading.
-%We would like to thank Andriy Berestovskyy for providing the Bugzilla import tool, and Martin Michlmayr for doing the first steps in porting our modifications upstream.
-
+\noindent\footnotesize{\textbf{Acknowledgements.}}
+We would like to thank Ambre Williams for valuable advice and help on the design and implementation of the data analyses, C\'ecile Rottner for general advice, support, and proof-reading, and Yann R\'egis-Gianas for encouragements and guidance.
+We would also like to thank the anonymous reviewers from MSR 2019 and ICSME 2019 for valuable feedback.
+Finally, we would like to thank Andriy Berestovskyy for providing the Bugzilla import tool, and Martin Michlmayr for performing the first steps in porting our modifications upstream.
 %
 % ---- Bibliography ----
+\balance
 \bibliography{bib}
